### 2023-07-17 fig/convergence.tex
init

\begin{tikzpicture}
    \begin{semilogyaxis}[
        xlabel={\# samples},
        ylabel={relative error},
        legend entries={$\varphi_3$, $\varphi_2$, $i_\mathrm{L}$},
        % legend entries={$\varphi_3$, $\varphi_2$},
        legend pos=north east
        ]

        % index 1
        \addplot[color=TUDa-1a, mark=*, mark size=3pt] table {do1_convergence_phi_3.dat};
        \addplot[color=TUDa-9a, mark=triangle*, mark size=3pt] table {do1_convergence_phi_2.dat};
        \addplot[color=TUDa-6a, mark=square*, mark size=3pt] table {do1_convergence_i_L.dat};

        % index 2
        % \addplot[color=TUDa-1a, mark=*, mark size=3pt] table {do2_convergence_phi_3.dat};
        % \addplot[color=TUDa-9a, mark=triangle*, mark size=3pt] table {do2_convergence_phi_2.dat};
    \end{semilogyaxis}
\end{tikzpicture}

### 2023-07-17 create_do1_doe_data.m

@@ -1,0 +1,35 @@
+N_t = 1000;
+
+T_0 = 0;
+T   = 1e-2;
+
+% all available data
+L = [1e-3:0.1e-3:3e-3];
+C = [100e-9:10e-9:300e-9];
+t = [linspace(T_0, T, N_t)]';
+
+x_t   = [];
+phi_3 = [];
+i_L   = [];
+phi_2 = [];
+x_L   = [];
+x_C   = [];
+for iL=1:length(L)
+    for iC=1:length(C)
+        filename = ['do1_L=' num2str(L(iL)) '_C=' num2str(C(iC)) '.dat'];
+        data     = dlmread(filename);
+        data     = data(2:end,:);
+        t_data   = data(:,1);
+        x_t      = [x_t; t];
+
+        phi_3 = [phi_3; interp1(t_data, data(:,2), t)];
+        i_L   = [i_L; interp1(t_data, data(:,3), t)];
+        phi_2 = [phi_2; interp1(t_data, data(:,5), t)];
+
+        x_L = [x_L; L(iL)*ones(N_t, 1)];
+        x_C = [x_C; C(iC)*ones(N_t, 1)];
+    end
+end
+
+filename = ['do1_doe_N_t=' num2str(N_t) '.dat'];
+write_dat (filename, x_t', [x_L'; x_C'; phi_3'; i_L'; phi_2']);

### 2023-07-17 create_do2_doe_data.m

@@ -1,0 +1,33 @@
+N_t = 1000;
+
+T_0 = 0;
+T   = 1e-2;
+
+% all available data
+L = [1e-3:0.1e-3:3e-3];
+C = [100e-9:10e-9:300e-9];
+t = [linspace(T_0, T, N_t)]';
+
+x_t   = [];
+phi_3 = [];
+phi_2 = [];
+x_L   = [];
+x_C   = [];
+for iL=1:length(L)
+    for iC=1:length(C)
+        filename = ['do2_L=' num2str(L(iL)) '_C=' num2str(C(iC)) '.dat'];
+        data     = dlmread(filename);
+        data     = data(2:end,:);
+        t_data   = data(:,1);
+        x_t      = [x_t; t];
+
+        phi_3 = [phi_3; interp1(t_data, data(:,2), t)];
+        phi_2 = [phi_2; interp1(t_data, data(:,4), t)];
+
+        x_L = [x_L; L(iL)*ones(N_t, 1)];
+        x_C = [x_C; C(iC)*ones(N_t, 1)];
+    end
+end
+
+filename = ['do2_doe_N_t=' num2str(N_t) '.dat'];
+write_dat (filename, x_t', [x_L'; x_C'; phi_3'; phi_2']);

### 2023-07-17 diode_oscillator.m

@@ -1,0 +1,121 @@
+problem = '2.2'; % 1.1, 1.2, 2.1, 2.2
+
+T_0    = 0;
+T      = 1e-2;
+Deltat = 1e-6;
+
+R = 500;
+G = 1/R;
+L = 2e-3;
+C = 100e-9;
+
+g_D  = @(v_D) 1e-14*(exp(v_D/26e-3) - 1);
+
+v_s  = @(t) sin(2*pi*300*t); % f = 300
+i_s  = @(t) 1e-4*sin(2*pi*200*t); % f = 200
+i_sp = @(t) 1e-4*2*pi*200*cos(2*pi*200*t);
+
+% index 1 problem: M x' + K(x) x = f(t)
+if (strcmp(problem, '1.1'))
+    M = [zeros(2,5); zeros(1,2) C zeros(1,2); zeros(1,3) L 0; zeros(1,5)];
+    K = @(v_D) [G -G 0 0 1; -G G 0 1 0; 0 0 g_D(v_D) -1 0; 0 -1 1 0 0; -1 0 0 0 0];
+    f = @(t) [0; 0; 0; 0; -v_s(t)];
+
+    % phi_3 = 0, i_L = 0 => phi_1 = v_s, phi_2 = v_s, i_v = 0
+    x_0 = [v_s(T_0); v_s(T_0); 0; 0; 0];
+
+    filename = 'do1.dat';
+    % if (~exist(filename))
+        % v_D = phi_3 = x_3
+        % Deltat         = 1e-7;
+        [x_ref, t_ref] = implicit_euler (T_0, T, Deltat, M, @(x) K(x(3)), @(t) f(t), x_0);
+        % write_dat (filename, t, x);
+
+        % Deltat  = [1e-4 1e-5 1e-6];
+        % e_phi_3 = NaN(1, 3);
+        % e_i_L   = NaN(1, 3);
+        % for it=1:length(Deltat)
+        %     [x, t]      = implicit_euler (T_0, T, Deltat(it), M, @(x) K(x(3)), @(t) f(t), x_0);
+        %     e_phi_3(it) = norm(x_ref(3,:) - interp1(t, x(3,:), t_ref)) / norm(x_ref(3,:));
+        %     e_i_L(it)   = norm(x_ref(4,:) - interp1(t, x(4,:), t_ref)) / norm(x_ref(4,:));
+        % end
+    % end
+end
+
[…]

### 2023-07-17 diode_oscillator_doe.m

@@ -1,0 +1,157 @@
+problem  = 1; % 1, 2
+variable = 'phi_3'; % phi_3, i_L, phi_2, phi_1
+tol      = [1e-2 5e-3 1e-3]; % 1e-2, 5e-3, 1e-3
+
+N_t = 1000;
+T_0 = 0;
+T   = 1e-2;
+
+% all available data
+L = [1e-3:0.1e-3:3e-3];
+C = [100e-9:10e-9:300e-9];
+t = linspace(T_0, T, N_t);
+
+% initial points
+t_i = [t(1); t(end); t(1); t(end); t(1); t(end); t(1); t(end)];
+L_i = [L(1); L(1); L(end); L(end); L(1); L(1); L(end); L(end)];
+C_i = [C(1); C(1); C(1); C(1); C(end); C(end); C(end); C(end)];
+
+if (problem == 1)
+    filename = ['do1_doe_N_t=' num2str(N_t) '.dat'];
+elseif (problem == 2)
+    filename = ['do2_doe_N_t=' num2str(N_t) '.dat'];
+end
+data = dlmread(filename);
+data = data(2:end,:);
+
+t_d = data(:,1);
+L_d = data(:,2);
+C_d = data(:,3);
+X   = [t_d L_d C_d];
+
+phi_3_d = data(:,4);
+
+R   = 500;
+i_s = @(t) 1e-4*sin(2*pi*200*t);
+if (problem == 1)
+    i_L_d   = data(:,5);
+    phi_2_d = data(:,6);
+elseif (problem == 2)
+    phi_2_d = data(:,5);
+    phi_1_d = phi_2_d + R*i_s(t_d);
+end
+
+if (strcmp(variable, 'phi_3'))
+    y = phi_3_d;
+elseif (strcmp(variable, 'i_L'))
+    y = i_L_d;
+elseif (strcmp(variable, 'phi_2'))
+    y = phi_2_d;
+elseif (strcmp(variable, 'phi_1'))
+    y = phi_1_d;
+end
+
+i_i = NaN(length(t_i), 1);
+for i=1:length(i_i)
+    i_i(i) = find((t_i(i) == t_d) & (L_i(i) == L_d) & (C_i(i) == C_d));
+end
+
+f = stk_model('stk_gausscov_aniso', 3);
+
+f.lognoisevariance = NaN;
+% f.lognoisevariance = -Inf;
+tic;
[…]

### 2023-07-17 diode_oscillator_doe_av.m

@@ -1,0 +1,107 @@
+problem  = 2; % 2
+variable = 'i_L'; % i_L
+tol      = [1e-3]; % 1e-3
+
+N_t = 1000;
+T_0 = 0;
+T   = 1e-2;
+
+% all available data
+t = linspace(T_0, T, N_t);
+
+% initial points (need more than boundary)
+t_i = [t(1); t(500); t(end)];
+
+filename = ['do2_doe_N_t=' num2str(N_t) '.dat'];
+data     = dlmread(filename);
+data     = data(2:end,:);
+
+t_d = data(1:N_t,1);
+X   = [t_d];
+
+R   = 500;
+i_s = @(t) 1e-4*sin(2*pi*200*t);
+if (strcmp(variable, 'i_L'))
+    y = i_s(t_d);
+end
+
+i_i = NaN(length(t_i), 1);
+for i=1:length(i_i)
+    i_i(i) = find((t_i(i) == t_d));
+end
+
+f = stk_model('stk_gausscov_aniso', 1);
+
+f.lognoisevariance = NaN;
+% f.lognoisevariance = -Inf;
+tic;
+[theta, sigma] = stk_param_init(f, X(i_i,:), y(i_i));
+fprintf('\nstk_param_init: %d min\n', toc/60);
+
+tic;
+[f.param, f.lognoisevariance] = stk_param_estim(f, X(i_i,:), y(i_i), theta, sigma);
+% [f.param] = stk_param_estim(f, X, y, theta);
+fprintf('\nstk_param_estim: %d min\n', toc/60);
+
+tic;
+y_p = stk_predict(f, X(i_i,:), y(i_i), X);
+fprintf('\nstk_predict: %d min\n', toc/60);
+
+i_t = i_i;
+e_p = 1;
+for i_tol=1:length(tol)
+    fprintf('\ntol: %g\n', tol(i_tol));
+    if (problem == 2)
+        filename = ['do2_doe_i_t_' variable '_tol=' num2str(tol(i_tol)) '_N_t=' num2str(N_t) '.dat'];
+        if (exist(filename))
+            i_t = dlmread(filename);
+
+            filename = ['do2_doe_param_' variable '_tol=' num2str(tol(i_tol)) '_N_t=' num2str(N_t) '.dat'];
+            f.param = dlmread(filename);
+
+            filename = ['do2_doe_lognoisevariance_' variable '_tol=' num2str(tol(i_tol)) '_N_t=' num2str(N_t) '.dat'];
[…]

### 2023-07-17 diode_oscillator_results.m

@@ -1,0 +1,236 @@
+problem  = 2; % 1, 2
+
+plot_input       = 0; % 0, 1
+plot_convergence = 0; % 0, 1
+plot_solution    = 0; % 0, 1
+plot_consistency = 1; % 0, 1
+
+% only relevant for input/convergence
+variable = 'phi_3'; % phi_3, i_L, phi_2
+
+% only relevant for solution
+tol = 1e-3; % 1e-2, 5e-3, 1e-3
+
+if (plot_input == 1)
+    TOL = [1e-2, 5e-3, 1e-3];
+    N_t = 1000;
+
+    for tol=TOL
+        if (problem == 1)
+            filename = ['do1_doe_i_t_' variable '_tol=' num2str(tol) '_N_t=' num2str(N_t) '.dat'];
+            i_t      = dlmread(filename);
+
+            filename = ['do1_doe_N_t=' num2str(N_t) '.dat'];
+        elseif (problem == 2)
+            filename = ['do2_doe_i_t_' variable '_tol=' num2str(tol) '_N_t=' num2str(N_t) '.dat'];
+            i_t      = dlmread(filename);
+
+            filename = ['do2_doe_N_t=' num2str(N_t) '.dat'];
+        end
+
+        data = dlmread(filename);
+        data = data(2:end,:);
+
+        t = data(:,1);
+        L = data(:,2);
+        C = data(:,3);
+        X = [t L C];
+
+        if (problem == 1)
+            filename = ['do1_input_' variable '_tol=' num2str(tol) '.dat'];
+        elseif (problem == 2)
+            filename = ['do2_input_' variable '_tol=' num2str(tol) '.dat'];
+        end
+
+        fid = fopen(filename, 'w');
+        fprintf(fid, 'x  y\n');
+        dlmwrite(fid, X(i_t,2:3), 'delimiter', '  ', 'append', 'on');
+        fclose(fid);
+    end
+end
+
+if (plot_convergence == 1)
+    tol = [1e-2, 5e-3, 1e-3];
+    N_t = 1000;
+
+    N_s = NaN(length(tol), 1);
+    for it=1:length(tol)
+       if (problem == 1)
+            filename = ['do1_doe_i_t_' variable '_tol=' num2str(tol(it)) '_N_t=' num2str(N_t) '.dat'];
[…]

### 2023-07-17 do1_sc.m

@@ -1,0 +1,23 @@
+% syms, sym, simplify, jacobian, inv, null, rank, +, -, *, /, ^, .'
+pkg load symbolic
+
+syms phi_1 phi_2 phi_3 i_L i_V G C L g_D v_s
+
+M = [[0, 0, 0, 0, 0]; [0, 0, 0, 0, 0]; [0, 0, C, 0, 0]; [0, 0, 0, L, 0]; [0, 0, 0, 0, 0]];
+
+K = [[G, -G, 0, 0, 1]; [-G, G, 0, 1, 0]; [0, 0, g_D, -1, 0]; [0, -1, 1, 0, 0]; [-1, 0, 0, 0, 0]];
+
+f = [[0]; [0]; [0]; [0]; [-v_s]];
+
+% first basis functions
+Q = W = sym([[1, 0, 0]; [0, 1, 0]; [0, 0, 0]; [0, 0, 0]; [0, 0, 1]]);
+P = V = sym([[0, 0]; [0, 0]; [1, 0]; [0, 1]; [0, 0]]);
+
+% first stage
+Mt   = V.' * M * P;
+Kt_P = V.' * K * P;
+Kt_Q = V.' * K * Q;
+ft   = V.' * f;
+Kb_P = W.' * K * P;
+Kb_Q = W.' * K * Q;
+fb   = W.' * f;

### 2023-07-17 do2_sc.m

@@ -1,0 +1,49 @@
+% syms, sym, simplify, jacobian, inv, null, rank, +, -, *, /, ^, .'
+pkg load symbolic
+
+syms phi_1 phi_2 phi_3 i_L G C L g_D i_s i_sp
+
+M = [[0, 0, 0, 0]; [0, 0, 0, 0]; [0, 0, C, 0]; [0, 0, 0, L]];
+
+K = [[G, -G, 0, 0]; [-G, G, 0, 1]; [0, 0, g_D, -1]; [0, -1, 1, 0]];
+
+f = [[i_s]; [0]; [0]; [0]];
+
+% first basis functions
+Q = W = sym([[1, 0]; [0, 1]; [0, 0]; [0, 0]]);
+P = V = sym([[0, 0]; [0, 0]; [1, 0]; [0, 1]]);
+
+% first stage
+Mt   = V.' * M * P;
+Kt_P = V.' * K * P;
+Kt_Q = V.' * K * Q;
+ft   = V.' * f;
+Kb_P = W.' * K * P;
+Kb_Q = W.' * K * Q;
+fb   = W.' * f;
+
+% second basis functions
+Qb = Wb = sym([[1]; [1]])
+
+% Wb.' * Kb_P
+Qt  = sym([[1]; [0]])
+Qtp = sym([[0]; [0]]);
+Pt  = sym([[0]; [1]])
+Ptp = sym([[0]; [0]]);
+
+% xt_P = inv(Wb.' * Kb_P * Pt) * Wb.' * fb;
+xt_P  = i_s;
+xt_Pp = i_sp;
+
+% (Mt * Qt).';
+Wt = sym([[0]; [1]])
+Vt = sym([[1]; [0]])
+
+% second stage
+Mt_2  = Vt.' * Mt * Qt
+Kt_P2 = Vt.' * (Mt * Qtp + Kt_P * Qt)
+Kt_Q2 = Vt.' * Kt_Q * Qb
+ft_2  = Vt.' * (ft - Mt * Pt * xt_Pp - (Mt * Ptp + Kt_P * Pt) * xt_P)
+Kb_P2 = Wt.' * (Mt * Qtp + Kt_P * Qt)
+Kb_Q2 = Wt.' * Kt_Q * Qb
+fb_2  = Wt.' * (ft - Mt * Pt * xt_Pp - (Mt * Ptp + Kt_P * Pt) * xt_P)

### 2023-07-17 implicit_euler.m

@@ -1,0 +1,35 @@
+%   [x, t] = implicit_euler (T_0, T, Deltat, M, K, f, x_0)
+%
+% Computes the numerical solution of an initial value problem of the form
+%      M x' + K(x) x = f(t) on [T_0,T]
+%               x(0) = x_0
+% using the implicit Euler method.
+%
+% INPUT:
+%       - T_0, T, Deltat: time interval [T_0,T] and timestep
+%       - M: constant matrix
+%       - K: @(x) nonlinear matrix
+%       - f: @(t) time dependent rhs
+%       - x_0: initial condition
+%
+% OUTPUT:
+%       - x: [N_x, N_t] solution vector at the discrete points in time
+%       - t: [1, N_t] vector containing the discrete points in time
+
+function [x, t] = implicit_euler (T_0, T, Deltat, M, K, f, x_0)
+    N_x = length(x_0);
+    N_t = round((T - T_0) / Deltat);
+    x   = NaN(N_x, N_t+1);
+    x(:,1) = x_0;
+
+    t = T_0:Deltat:(T_0 + N_t*Deltat);
+    for n=1:N_t
+        if (~mod(n, 1000))
+            fprintf('\niteration no.: %i/%i\n', n, N_t);
+        end
+
+        % M (x_npo - x_n) + Deltat K(x_npo) x_npo = Deltat f(t_npo)
+        g = @(x_npo) M*(x_npo - x(:,n)) + Deltat*K(x_npo)*x_npo - Deltat*f(t(n+1));
+        x(:,n+1) = fsolve(g, x(:,n));
+    end
+end

### 2023-11-30 doc/article/article.tex
index-aware-learning

\documentclass[AMA,STIX1COL]{WileyNJD-v2}

\usepackage{mathtools}
\usepackage{tuda-pgfplots}
\usepackage{siunitx}
\usepackage{subcaption}

\articletype{Original Article}

\received{}
\revised{}
\accepted{}

\raggedbottom
\newcommand{\mb}[1]{\mathbf{#1}}
\newcommand{\mbt}[1]{\tilde{\mathbf{#1}}}
\newcommand{\mbb}[1]{\bar{\mathbf{#1}}}
\newcommand{\mbh}[1]{\hat{\mathbf{#1}}}
\newcommand{\mr}[1]{\mathrm{#1}}
\newcommand{\T}{{\!\top}}
\newcommand{\ddt}{\frac{\mathrm{d}}{\mathrm{d}t}}
\newcommand{\A}[1]{\mb{A}_\mr{#1}}
\newcommand{\AT}[1]{\mb{A}_\mr{#1}^{\T}}
\newcommand{\qC}{\mb{q}_\mr{C}}
\newcommand{\gR}{\mb{g}_\mr{R}}
\newcommand{\phiL}{\boldsymbol{\phi}_\mr{L}}
\newcommand{\vphi}{\boldsymbol{\varphi}}
\renewcommand{\i}[1]{\mb{i}_\mr{#1}}
\renewcommand{\v}[1]{\mb{v}_\mr{#1}}

\begin{document}

\title{Index-aware learning of circuits}

\author[1]{Idoia Cortes Garcia}
\author[1,2]{Peter Förster*}
\author[2]{Lennart Jansen}
\author[1]{Wil Schilders}
\author[2]{Sebastian Schöps}
% \authormark{}
\address[1]{\orgname{Eindhoven University of Technology}, \orgaddress{\country{Netherlands}}}
\address[2]{\orgname{Technical University of Darmstadt}, \orgaddress{\country{Germany}}}
\corres{*Peter Förster, Eindhoven University of Technology. \email{p.f.forster@tue.nl}}
% \presentaddress{}

\abstract[Summary]{Electrical circuits are present in a variety of technologies, making their design an important part of computer aided engineering. The growing number of tunable parameters that affect the final design leads to a need for new approaches of quantifying their impact. Machine learning may play a key role in this regard, however current approaches often make suboptimal use of existing knowledge about the system at hand. In terms of circuits, their description via modified nodal analysis is well-understood. This particular formulation leads to systems of differential-algebraic equations (DAEs) which bring with them a number of peculiarities, e.g. hidden constraints that the solution needs to fulfill. We aim to use the recently introduced dissection concept for DAEs that can decouple a given system into ordinary differential equations, only depending on differential variables, and purely algebraic equations that describe the relations between differential and algebraic variables. The idea then is to only learn the differential variables and reconstruct the algebraic ones using the relations from the decoupling. This approach guarantees that the algebraic constraints are fulfilled up to the accuracy of the nonlinear system solver, which represents the main benefit highlighted in this article.}

\keywords{electrical circuits, differential-algebraic equations, modified nodal analysis, dissection concept, machine learning}

\maketitle

\section{Introduction}
\label{sec:int}
Design optimization and uncertainty quantification are ubiquitous tools of modern computer aided  engineering. Due to the ever-increasing complexity of engineering systems, machine learning techniques have gained popularity for constructing surrogate models when the objective function becomes expensive to evaluate and a large number of (design or uncertainty) parameters are present. In these cases, classical model order reduction techniques\cite{schilders2008} or function approximation approaches\cite{xiu2010} suffer from the curse of dimensionality: the number of operations to construct and the memory required to store the model grow exponentially with respect to the number of parameters. Experimental and in some cases even theoretical evidence\cite{beneventano2020} shows that machine learning approaches may be able to overcome the curse of dimensionality and provide surrogates that are fast to evaluate, while requiring comparably little data.

In the particular case of electrical circuit design, neural networks have been used for design optimization for over 20 years\cite{zaabab1995,fang2000}. More recently, Gaussian process regression has been employed for both uncertainty quantification and design optimization during analog integrated circuit design\cite{wang2018,sanabria2020}. The commonality between these approaches is that they focus on the learning part: they aim to provide a computationally efficient and accurate surrogate, given data produced by some circuit simulation software. Thus, they all treat the circuit simulator as a black box that simply provides the data which is then used for constructing the surrogate. In contrast, we want to exploit the known structure that underlies the equations describing electrical circuits.

More specifically, we consider the modified nodal analysis\cite{ho1975} (MNA). MNA is one of the most popular circuit descriptions and lies at the center of SPICE-like simulation software such as LTspice\cite{ltspice}, Xyce\cite{xyce} and PSpice\cite{pspice}. Applying MNA to a given circuit leads to systems of differential-algebraic equations (DAEs), which can generally be written as systems of implicit differential equations \cite{denk1991}
\begin{align*}
    \mb{f}(\mb{x}', \mb{x}, t, \mb{p}) = \mb{0}, \quad \mb{x}(0) = \mb{x}_0,
\end{align*}
where the Jacobian $\mb{J}_{\mb{x}'}(\mb{f})$, of $\mb{f}$ w.r.t.~$\mb{x}'$, is singular and $\mb{p}$ are the design or uncertainty parameters. To give a more intuitive description, one can think of DAEs as ordinary differential equations (ODEs) that are constrained to a manifold determined by (hidden) constraints on the solution variables. Multiple so-called index concepts were introduced in order to characterize the difficulties that appear in the context of DAEs. A reference by Mehrmann\cite{mehrmann2015} outlines some of the most prominent, including the differentiation, tractability, strangeness and perturbation indices. For all of these, the index $\nu \in \mathbb{N}$ depends on structural properties of the DAE and (informally) measures the degree of difficulty associated with solving it. In this sense, higher index DAEs usually lead to more problems, both analytically and numerically.

Various relationships between the indices, including situations when they coincide, are also known\cite{mehrmann2015} and the structure and index of MNA are well understood for both independent sources\cite{tischendorf1999} and controlled sources\cite{estevez2000ijcta}. Rather recently, the dissection concept was introduced\cite{jansen2014}. It combines features of the strangeness and tractability indices to decouple a DAE into an ODE and a set of purely algebraic equations. In particular, the entire dynamics of the solution may then be found only using the ODE. We aim to use this dissection concept to propose an approach for learning electrical circuits more accurately and efficiently.

In the following, \autoref{sec:mna} introduces MNA in more detail and states some well-known related results. Afterwards, \autoref{sec:dc} outlines the dissection concept and showcases its properties using small example circuits. The same examples are then used to present the new approach on an abstract level in \autoref{sec:ial} and on a numerical level in \autoref{sec:ne}, where we also showcase an alternative method that trades some of the accuracy of our approach for an easier implementation. Preliminary conclusions about the effectiveness of our approach and future research directions are given in \autoref{sec:cfr}.

\section{Modified nodal analysis}
\label{sec:mna}
We begin by looking at the system of DAEs that results from MNA. From here onward we will disregard controlled sources. As these are of crucial importance for engineering applications however, we will state whether extensions including controlled sources are available or missing at the appropriate times. Borrowing the notation from Tischendorf\cite{tischendorf1999}, and only explicitly expressing the time dependence of independent sources for brevity, the system of MNA reads
\begin{subequations}
    \label{eq:mnai}
    \begin{align}
            \A{C} \ddt \qC \big( \AT{C} \vphi \big) + \A{R} \gR \big( \AT{R} \vphi \big) + \A{L} \i{L} + \A{V} \i{V} + \A{I} \i{s}(t) &= \mb{0}\\ 
            \ddt \phiL \big( \i{L} \big) - \AT{L} \vphi &= \mb{0}\\
            \AT{V} \vphi - \v{s}(t)&= \mb{0},
    \end{align}
\end{subequations}
where $\gR$, $\qC$ and $\phiL$ represent functions modeling resistive, capacitive or inductive devices respectively. The terms for independent current and voltage sources are given by $\i{s}(t)$ and $\v{s}(t)$, while $\vphi$ denotes the vector of nodal potentials and $\i{L} $, $\i{V}$ are the currents flowing through branches containing inductors or voltage sources respectively. The last ingredient is given by the incidence matrices $\A{*}$, where $*$ indicates the device type. These collect the branch to node relations of the underlying electrical network when considering the branches and nodes as edges and vertices of a graph. In the context of design optimization or uncertainty quantification, the device functions $\gR$, $\qC$ or $\phiL$ could e.g.~depend on (subsets of) the parameters $\mb{p}$.

In order to obtain a version of \eqref{eq:mnai} that is better suited to analysis and implementation, we consider the device function Jacobians
\begin{align}
    \mb{L}(\i{L}) \coloneqq \mb{J}_{\i{L}}(\phiL), \quad \mb{C} \big( \AT{C} \vphi \big) \coloneqq \mb{J}_{\AT{C} \vphi}(\qC) \label{eq:dfj},
\end{align}
where we use the same notation for the Jacobians as in \autoref{sec:int}. We also assume that the device function for resistors may be expressed as
\begin{align}
    \gR \big( \AT{R} \vphi \big) = \mb{G} \big( \AT{R} \vphi \big) \AT{R} \vphi \label{eq:dfr}.
\end{align}
In practice, this is accomplished by either considering the chordal (static) conductance $\frac{[\gR]_i}{[\AT{R} \vphi]_i}$ for each resistor $\mr{R}_i$ separately or by using the Jacobian $\mathbf{J}_{ \AT{R} \vphi}(\gR)$. Inserting \eqref{eq:dfj} and \eqref{eq:dfr} into the original system and writing it in matrix form yields
\begin{align}
    \begin{bmatrix}
        \A{C}^{\phantom{\T}} \mb{C} \big( \AT{C} \vphi \big) \AT{C} & &\\
        & \mb{L}(\i{L}) &\\
        & & \mb{0}
    \end{bmatrix} \ddt \begin{bmatrix}
        \vphi\\
        \i{L}\\
        \i{V}
    \end{bmatrix} + \begin{bmatrix}
        \A{R}^{\phantom{\T}} \mb{G} \big( \AT{R} \vphi \big) \AT{R} & \A{L} & \A{V}\\
        -\AT{L} & &\\
        -\AT{V} & &
    \end{bmatrix} \begin{bmatrix}
        \vphi\\
        \i{L}\\
        \i{V}
    \end{bmatrix} + \begin{bmatrix}
        \A{I} \i{s}(t)\\
        \mb{0}\\
        \v{s}(t)
    \end{bmatrix} = \mb{0} \label{eq:mnaii},
\end{align}
where we assume that $\qC$ and $\phiL$ do not explicitly depend on time. MNA is usually presented in the form of \eqref{eq:mnaii} and for it, there exists the following well-known topological index result.
\begin{quote}
    Assuming the matrix-valued functions $\mathbf{G}$, $\mathbf{L}$, $\mathbf{C}$ from \eqref{eq:dfj} and \eqref{eq:dfr} positive definite:
    \begin{enumerate}
        \item The index of \eqref{eq:mnaii} is $\nu \leq 2$.
        \item The index of \eqref{eq:mnaii} is $\nu \leq 1$, if and only if there are no loops consisting only of capacitors and voltage sources containing at least one voltage source and no cutsets consisting only of inductors and current sources.
    \end{enumerate}
    The result can be found in terms of the tractability index\cite{tischendorf1999}, the differentiation index \cite{estevez2000ijcta} and the dissection concept\cite{jansen2014}. % \cite[Theorem 4.28]{jansen2014}
\end{quote}
We remark that there also exist extensive results about when the index of MNA including controlled sources does not exceed $\nu = 2$\cite{estevez2000ijcta}.

\section{Dissection concept}
\label{sec:dc}
This section serves as a general introduction to the dissection concept and more specifically its application to MNA as given in \eqref{eq:mnaii}. Consider a linear DAE in standard form\cite{jansen2014},
\begin{align}
    \mb{M} \ddt \mb{x} + \mb{K} \mb{x} + \mb{f}(t) = \mb{0}, \quad \ker \mb{M} \supsetneq \{ \mb{0} \} \label{eq:dae},
\end{align}
where we again keep the time dependence of $\mb{x}$ implicit. We now demonstrate the first two steps of the dissection concept when applied to systems of the form of \eqref{eq:dae} and remark that although \eqref{eq:mnaii} is a nonlinear DAE, the decoupling still takes the same form in this case due to the special structure of MNA\cite{jansen2014}.

\subsection{Index one case}
Assuming $\mb{M} \in \mathbb{R}^{n \times m}$, we define four basis functions $\mb{P}$, $\mb{Q}$, $\mb{V}$, $\mb{W}$ such that
\begin{align*}
    \mr{im\, } \mb{Q} = \ker \mb{M}, \quad \mr{im\, } \mb{W} = \ker \mb{M}^{\T}
\end{align*}
and the columns of $\mb{P}$ and $\mb{Q}$ together form a basis of $\mathbb{R}^n$, while the columns of $\mb{V}$ and $\mb{W}$ together form a basis of $\mathbb{R}^m$.

\paragraph{Splitting the solution}
The key idea is to use the basis functions to split the solution $\mb{x}$ into two parts
\begin{align}
    \mb{x} = \mb{P} \mbt{x} + \mb{Q} \mbb{x} \label{eq:vs},
\end{align}
where $\tilde{\cdot}$ is used to indicate differential (dynamic) variables and $\bar{\cdot}$ signifies algebraic (fixed) variables. When inserting the splitting \eqref{eq:vs} into the general system \eqref{eq:dae} we observe the following
\begin{align}
    \mb{M} \mb{P} \ddt \mbt{x} + \mb{K} \mb{P} \mbt{x} + \mb{K} \mb{Q} \mbb{x} + \mb{f}(t) = \mb{0} \label{eq:rs}
\end{align}
and this motivates the notation, as only $\mbt{x}$ appears differentiated in time.

\paragraph{Splitting the system}
The procedure then continues by multiplying \eqref{eq:rs} with $\mb{V}^{\T}$ and $\mb{W}^{\T}$ from the left. This yields
\begin{subequations}
    \label{eq:io}
    \begin{align}
            \underbrace{\mb{V}^{\T} \mb{M} \mb{P}}_{\eqqcolon \mbt{M}} \ddt \mbt{x} + \underbrace{\mb{V}^{\T} \mb{K} \mb{P}}_{\eqqcolon \mbt{K}_\mb{P}} \mbt{x} + \underbrace{\mb{V}^{\T} \mb{K} \mb{Q}}_{\eqqcolon \mbt{K}_\mb{Q}} \mbb{x} + \underbrace{\mb{V}^{\T} \mb{f}(t)}_{\eqqcolon \mbt{f}(t)} &= \mb{0} \label{eq:ioa}\\
            \underbrace{\mb{W}^{\T} \mb{K} \mb{P}}_{\eqqcolon \mbb{K}_\mb{P}} \mbt{x} + \underbrace{\mb{W}^{\T} \mb{K} \mb{Q}}_{\eqqcolon \mbb{K}_\mb{Q}} \mbb{x} + \underbrace{\mb{W}^{\T} \mb{f}(t)}_{\eqqcolon \mbb{f}(t)} &= \mb{0} \label{eq:iob},
    \end{align}
\end{subequations}
where we also introduce shorthands for the arising products. The key observations regarding \eqref{eq:io} are the following: $\mbt{M}$ is regular by construction, thus if $\mbb{K}_\mb{Q}$ is regular, \eqref{eq:iob} gives a purely algebraic equation for $\mbb{x}$ in terms of $\mbt{x}$. Solving this equation for $\mbb{x}$ and inserting the solution into \eqref{eq:ioa} then yields an ODE in the differential variables $\mbt{x}$. In this case, the \textbf{(dissection) index} of the DAE is defined to be $\nu = 1$.

\subsection{Index two case}
There are of course many DAEs, including those described by MNA, which have a higher index. For these the dissection concept proceeds by introducing additional basis functions and continuing the splitting process in a similar fashion.

\paragraph{Splitting the algebraic variables}
We begin by considering basis functions $\mbb{P}, \mbb{Q}, \mbb{V}, \mbb{W}$ of $\mbb{K}_\mb{Q}$ and split the algebraic variables $\mbb{x}$ further as follows
\begin{align}
    \mbb{x} = \mbb{P} \mbb{x}_\mb{P} + \mbb{Q} \mbb{x}_\mb{Q} \label{eq:xb}.
\end{align}
Inserting this decomposition into \eqref{eq:iob} and multiplying once by $\mbb{V}^{\T}$ and $\mbb{W}^{\T}$ from the left yields
\begin{subequations}
    \label{eq:iti}
    \begin{align}
        \mbb{V}^{\T} \mbb{K}_\mb{P} \mbt{x} + \mbb{V}^{\T} \mbb{K}_\mb{Q} \mbb{P} \mbb{x}_\mb{P} + \mbb{V}^{\T} \mbb{f}(t) &= \mb{0} \label{eq:itia}\\
        \mbb{W}^{\T} \mbb{K}_\mb{P} \mbt{x} + \mbb{W}^{\T} \mbb{f}(t) &= \mb{0} \label{eq:itib}.
    \end{align}
\end{subequations}

\paragraph{Splitting the differential variables}
We now need basis functions $\mbt{P}, \mbt{Q}$ of $\mbb{W}^{\T} \mbb{K}_\mb{P}$ to split the differential variables $\mbt{x}$ as
\begin{align}
    \mbt{x} = \mbt{P} \mbt{x}_\mb{P} +
    \mbt{Q} \mbt{x}_\mb{Q} \label{eq:xt}
\end{align}
and inserting this splitting into \eqref{eq:itib} then allows for an explicit description of $\mbt{x}_\mb{P}$
\begin{align*}
    \mbt{x}_\mb{P} = -\big( \mbb{W}^{\T} \mbb{K}_\mb{P} \mbt{P} \big)^{-1} \mbb{W}^{\T} \mbb{f}(t) \eqqcolon \mbt{f}_\mb{P}(t),
\end{align*}
since $\mbb{W}^{\T} \mbb{K}_\mb{P} \mbt{P}$ is assumed to be regular (otherwise the DAE is underdetermined\cite{jansen2014}).

\paragraph{Splitting the solution}
Similarly, inserting \eqref{eq:xt} into \eqref{eq:itia} gives a description of $\mbb{x}_\mb{P}$ in terms of $\mbt{x}_\mb{Q}$ as $\mbb{V}^{\T} \mbb{K}_\mb{Q} \mbb{P}$ is guaranteed to be nonsingular
\begin{align*}
    \mbb{x}_\mb{P} = -\big( \mbb{V}^{\T} \mbb{K}_\mb{Q} \mbb{P} \big)^{-1} \big( \mbb{V}^{\T} \mbb{K}_\mb{P} \mbt{Q} \mbt{x}_\mb{Q} + \mbb{V}^{\T} \mbb{K}_\mb{P} \mbt{P} \mbt{x}_\mb{P} + \mbb{V}^{\T} \mbb{f}(t) \big) \eqqcolon \mbb{K} \mbt{x}_\mb{Q} + \mbb{f}_\mb{P}(t).
\end{align*}
Expanding $\mbt{x}$ and $\mbb{x}$ in \eqref{eq:ioa}, according to \eqref{eq:xb} and \eqref{eq:xt} respectively, and using the above expressions for $\mbt{x}_\mb{P}$ and $\mbb{x}_\mb{P}$ gives
\begin{align}
    \mbt{M} \mbt{Q} \ddt \mbt{x}_\mb{Q} + \big( \mbt{K}_\mb{P} \mbt{Q} + \mbt{K}_\mb{Q} \mbb{P} \mbb{K} \big) \mbt{x}_\mb{Q} + \mbt{K}_\mb{Q} \mbb{Q} \mbb{x}_\mb{Q} + \mbh{f}(t) = \mb{0} \label{eq:itii},
\end{align}
where we define
\begin{align*}
    \mbh{f}(t) \coloneqq \mbt{M} \mbt{P} \ddt \mbt{f}_\mb{P}(t) + \mbt{K}_\mb{P} \mbt{P} \mbt{f}_\mb{P}(t) + \mbt{K}_\mb{Q} \mbb{P} \mbb{f}_\mb{P}(t) + \mbt{f}(t).
\end{align*}

\paragraph{Splitting the system}
Finally, we require basis functions $\mbt{V}$ and $\mbt{W}$ of $\mbt{M} \mbt{Q}$. Multiplying \eqref{eq:itii} from the left by $\mbt{V}^{\T}$ and $\mbt{W}^{\T}$ and redefining $\mbt{x}_2 \coloneqq \mbt{x}_\mb{Q}$ and $\mbb{x}_2 \coloneqq \mbb{x}_\mb{Q}$ yields
\begin{subequations}
    \label{eq:itiii}
    \begin{align}
            \underbrace{\mbt{V}^{\T} \mbt{M} \mbt{Q}}_{\eqqcolon \mbt{M}_2} \ddt \mbt{x}_2 + \underbrace{\mbt{V}^{\T} \big( \mbt{K}_\mb{P} \mbt{Q} + \mbt{K}_\mb{Q} \mbb{P} \mbb{K} \big)}_{\eqqcolon \mbt{K}_{\mb{P}_2}} \mbt{x}_2 + \underbrace{\mbt{V}^{\T} \mbt{K}_\mb{Q} \mbb{Q}}_{\eqqcolon \mbt{K}_{\mb{Q}_2}} \mbb{x}_2 + \underbrace{\mbt{V}^{\T} \mbh{f}(t)}_{\eqqcolon \mbt{f}_2(t)} &= \mb{0} \label{eq:itiiia}\\
            \underbrace{\mbt{W}^{\T} \big( \mbt{K}_\mb{P} \mbt{Q} + \mbt{K}_\mb{Q} \mbb{P} \mbb{K} \big)}_{\eqqcolon \mbb{K}_{\mb{P}_2}} \mbt{x}_2 + \underbrace{\mbt{W}^{\T} \mbt{K}_\mb{Q} \mbb{Q}}_{\eqqcolon \mbb{K}_{\mb{Q}_2}} \mbb{x}_2 + \underbrace{\mbt{W}^{\T} \mbh{f}(t)}_{\eqqcolon \mbb{f}_2(t)} &= \mb{0}. \label{eq:itiiib}
    \end{align}
\end{subequations}
We observe that \eqref{eq:itiii} is of the same form as \eqref{eq:io} and that $\mbt{M}_2$ is again regular by construction. Therefore, if $\mbb{K}_{\mb{Q}_2}$ is also regular we can extract an ODE analogous to the previous case by solving \eqref{eq:itiiib} for $\mbb{x}_2$ (in terms of $\mbt{x}_2$) and inserting that solution into \eqref{eq:itiiia}. The DAE is then said to have \textbf{dissection index} $\nu = 2$ and we call $\mbt{x}_2$ the differential variables and $\mbb{x}_2$ the algebraic variables. Furthermore, since \eqref{eq:itiii} and \eqref{eq:io} are structurally identical, we may simply apply the second step of the decoupling procedure to \eqref{eq:itiii} in order to deal with even higher index DAEs.

To clarify the naming convention for the following sections, we call $\mbt{x}_2$ and $\mbb{x}_2$ the purely differential and algebraic variables of an index two DAE respectively, whereas $\mbb{x}$ denotes the explicit (obvious) algebraic variables and $\mbt{x}$ still contains hidden (implicit) algebraic variables that are identified in the second step of the dissection concept.

\subsection{Dissection concept and MNA}
\label{subsec:dcmna}
Before we consider example circuits, we take a brief look at the first step of the dissection concept for MNA. For this, we recall the system of MNA \eqref{eq:mnaii}
\begin{align*}
    \begin{bmatrix}
        \A{C}^{\phantom{\T}} \mb{C} \big( \AT{C} \vphi \big) \AT{C} & &\\
        & \mb{L}(\i{L}) &\\
        & & \mb{0}
    \end{bmatrix} \ddt \begin{bmatrix}
        \vphi\\
        \i{L}\\
        \i{V}
    \end{bmatrix} + \begin{bmatrix}
        \A{R}^{\phantom{\T}} \mb{G} \big( \AT{R} \vphi \big) \AT{R} & \A{L} & \A{V}\\
        -\AT{L} & &\\
        -\AT{V} & &
    \end{bmatrix} \begin{bmatrix}
        \vphi\\
        \i{L}\\
        \i{V}
    \end{bmatrix} + \begin{bmatrix}
        \A{I} \i{s}(t)\\
        \mb{0}\\
        \v{s}(t)
    \end{bmatrix} = \mb{0}
\end{align*}
and determine the basis functions $\mb{P}$ and $\mb{Q}$. Taking into account the standard assumption, compare \autoref{sec:mna}, that $\mb{L}(\i{L})$ and $\mb{C} \big( \AT{C} \vphi \big)$ are positive definite, we find\cite{jansen2014}
\begin{align*}
    \mb{Q} = \begin{bmatrix}
        \mb{Q}_\mr{C} &\\
        & \mb{0}\\
        & \mb{I}
    \end{bmatrix}, \quad \mb{P} = \begin{bmatrix}
        \mb{P}_\mr{C} &\\
        & \mb{I}\\
        & \mb{0}
    \end{bmatrix},
\end{align*}
where $\mr{im\, } \mb{Q}_\mr{C} = \ker \AT{C}$, $\mb{0}$ and $\mb{I}$ are zero and identity matrices of appropriate dimension and the columns of $\mb{Q}_\mr{C}$ and $\mb{P}_\mr{C}$ together form a basis of $\mathbb{R}^{n_\varphi}$ with $n_\varphi$ the dimension of $\vphi$. The description of $\mb{Q}_\mr{C}$ results from the fact that (with $\mb{y} \coloneqq \AT{C} \mb{x}$)
\begin{align*}
    \mb{x}^{\T} \A{C}^{\phantom{\T}} \mb{C} \big( \AT{C} \vphi \big) \AT{C} \mb{x} = \mb{y}^{\T} \mb{C} \big( \AT{C} \vphi \big) \mb{y} > 0, \quad \forall \mb{y} \in \mathbb{R}^{n_\varphi} \setminus \{ \mb{0} \},
\end{align*}
since $\mb{C} \big( \AT{C} \vphi \big)$ is positive definite and hence the kernel of $\A{C}^{\phantom{\T}} \mb{C} \big( \AT{C} \vphi \big) \AT{C}$ is determined by the kernel of $\AT{C}$. As $\A{C}$ is an incidence (constant) matrix, we find that $\mb{Q}$ and $\mb{P}$ are also constant.

\paragraph{First example circuit}
\begin{figure}
    \begin{center}
        \includegraphics[width=0.5\textwidth]{do1}
    \end{center}
    \caption{First example circuit: simple diode oscillator.}
    \label{fig:do1}
\end{figure}
We now demonstrate the dissection concept by applying it to the example circuit given in \autoref{fig:do1}. The circuit contains a voltage source $v_\mathrm{s}(t)$, a linear resistor with conductance $G$, a linear capacitor with capacitance $C$, a linear inductor with inductance $L$, as well as a diode $\mathrm{D}$ that is modeled by a nonlinear resistance $g_\mathrm{D}(\varphi_3)$. Comparing the conditions of the topological index result from \autoref{sec:mna} with the example circuit shows that the circuit has index $\nu = 1$, thus we only have to perform the first step of the dissection concept.

We begin by writing out the system derived from applying MNA to the example
\begin{align*}
    \begin{bmatrix}
        0 & & & &\\
        & 0 & & &\\
        & & C & &\\
        & & & L &\\
        & & & & 0
    \end{bmatrix} \ddt \begin{bmatrix}
        \varphi_1\\
        \varphi_2\\
        \varphi_3\\
        i_\mr{L}\\
        i_\mr{V}
    \end{bmatrix} + \begin{bmatrix}
        G & -G & &  & 1\\
        -G & G & & 1 &\\
        & & g_\mr{D}(\varphi_3) & -1 &\\
        & -1 & 1 & &\\
        -1 & & & &
    \end{bmatrix} \begin{bmatrix}
        \varphi_1\\
        \varphi_2\\
        \varphi_3\\
        i_\mr{L}\\
        i_\mr{V}
    \end{bmatrix} + \begin{bmatrix}
        0\\
        0\\
        0\\
        0\\
        v_\mr{s}(t)
    \end{bmatrix} = \mb{0}.
\end{align*}
For the first basis functions $\mb{Q}$ and $\mb{P}$ we find
\begin{align*}
    \mb{Q} = \begin{bmatrix}
        1 & &\\
        & 1 &\\
        & 0 &\\
        & & 0\\
        & & 1
    \end{bmatrix}, \quad \mb{P} = \begin{bmatrix}
        0 &\\
        0 &\\
        1 &\\
        & 1\\
        & 0
    \end{bmatrix}
\end{align*}
and as the matrix equivalent to $\mb{M}$ in the context of \eqref{eq:dae} is symmetric, we have $\mb{W} = \mb{Q}$ and $\mb{V} = \mb{P}$ for the remaining two basis functions. Using these, we find the new unknowns $\mbt{x} = [\varphi_3\ i_\mr{L}]^{\T}$ and $\mbb{x} = [\varphi_1\ \varphi_2\ i_\mr{V}]^{\T}$, which allow us to obtain the systems corresponding to \eqref{eq:ioa} and \eqref{eq:iob}
\begin{subequations}
    \label{eq:ei}
    \begin{align}
        \begin{bmatrix}
            C &\\
            & L
        \end{bmatrix} \ddt \begin{bmatrix}
            \varphi_3\\
            i_\mr{L}
        \end{bmatrix} + \begin{bmatrix}
            g_\mr{D}(\varphi_3) & -1\\
            1 &
        \end{bmatrix} \begin{bmatrix}
            \varphi_3\\
            i_\mr{L}
        \end{bmatrix} + \begin{bmatrix}
            0 & &\\
            & -1 & 0
        \end{bmatrix} \begin{bmatrix}
            \varphi_1\\
            \varphi_2\\
            i_\mr{V}
        \end{bmatrix} &= \mb{0} \label{eq:eia}\\
        \begin{bmatrix}
            0 &\\
            & 1\\
            & 0
        \end{bmatrix} \begin{bmatrix}
            \varphi_3\\
            i_\mr{L}
        \end{bmatrix} + \begin{bmatrix}
            G & -G & 1\\
            -G & G &\\
            -1 & &
        \end{bmatrix} \begin{bmatrix}
            \varphi_1\\
            \varphi_2\\
            i_\mr{V}
        \end{bmatrix} + \begin{bmatrix}
            0\\
            0\\
            v_\mr{s}
        \end{bmatrix} &= \mb{0} \label{eq:eib}.
    \end{align}
\end{subequations}
Solving \eqref{eq:eib} for $\mbb{x}$ and subsequently inserting the solution into \eqref{eq:eia} then gives an ODE and a purely algebraic system as promised
\begin{align}
    \begin{bmatrix}
        C &\\
        & L
    \end{bmatrix} \ddt \begin{bmatrix}
        \varphi_3\\
        i_L
    \end{bmatrix} + \begin{bmatrix}
        g_\mr{D}(\varphi_3) & -1\\
        1 & R
    \end{bmatrix} \begin{bmatrix}
        \varphi_3\\
        i_\mr{L}
    \end{bmatrix} + \begin{bmatrix}
        0\\
        -v_\mr{s}
    \end{bmatrix} = \mb{0} \quad \text{and} \quad \begin{bmatrix}
        \varphi_1\\
        \varphi_2\\
        i_\mr{V}
    \end{bmatrix} = \begin{bmatrix}
        v_\mr{s}\\
        v_\mr{s} - R i_\mr{L}\\
        -i_\mr{L}
    \end{bmatrix}, \label{eq:do1}
\end{align}
where $R = 1/G$ is the inverse of the conductance.

\paragraph{Second example circuit}
\begin{figure}
    \begin{center}
        \includegraphics[width=0.5\textwidth]{do2}
    \end{center}
    \caption{Second example circuit: simple diode oscillator with current instead of voltage source.}
    \label{fig:do2}
\end{figure}
We derive a second example from the first by substituting a current source $i_\mr{s}(t)$ for the voltage source, compare \autoref{fig:do2}. Looking at the index criteria from \autoref{sec:mna}, we observe that this circuit has index $\nu = 2$, as there now is a cutset consisting of the inductor and current source. Therefore, we need to execute two steps of the dissection concept in order to split the equations into purely differential and algebraic parts respectively.

The corresponding MNA system is given by
\begin{align*}
    \begin{bmatrix}
        0 & & &\\
        & 0 & &\\
        & & C &\\
        & & & L
    \end{bmatrix} \ddt \begin{bmatrix}
        \varphi_1\\
        \varphi_2\\
        \varphi_3\\
        i_\mr{L}
    \end{bmatrix} + \begin{bmatrix}
        G & -G & &\\
        -G & G & & 1\\
        & & g_\mr{D}(\varphi_3) & -1\\
        & -1 & 1 &
    \end{bmatrix} \begin{bmatrix}
        \varphi_1\\
        \varphi_2\\
        \varphi_3\\
        i_\mr{L}
    \end{bmatrix} + \begin{bmatrix}
        -i_\mr{s}(t)\\
        0\\
        0\\
        0
    \end{bmatrix} = \mb{0},
\end{align*}
where we also note that the dimension is smaller compared to the previous example. The first two basis functions $\mb{Q}$ and $\mb{P}$ are
\begin{align*}
    \mb{Q} = \begin{bmatrix}
        1 &\\
        & 1\\
        & 0\\
        & 0
    \end{bmatrix}, \quad \mb{P} = \begin{bmatrix}
        0 &\\
        0 &\\
        1 &\\
        & 1
    \end{bmatrix},
\end{align*}
where it holds again that $\mb{W} = \mb{Q}$ and $\mb{V} = \mb{P}$ due to symmetry. We now list the remaining basis functions, omitting the intermediate steps for brevity,
\begin{align*}
    \mbb{Q} = \mbb{W} = \begin{bmatrix}
        1\\
        1
    \end{bmatrix}, \quad \mbb{P} = \bar{\mathbf{V}} = \begin{bmatrix}
        1\\
        0
    \end{bmatrix}, \quad \mbt{Q} = \begin{bmatrix}
        1\\
        0
    \end{bmatrix}, \quad \mbt{P} = \begin{bmatrix}
        0\\
        1
    \end{bmatrix}, \quad \mbt{W} = \begin{bmatrix}
        0\\
        1
    \end{bmatrix}, \quad \mbt{V} = \begin{bmatrix}
        1\\
        0
    \end{bmatrix}
\end{align*}
and finally obtain an ODE in one variable only and a second purely algebraic system that recovers the remaining unknowns
\begin{align}
    C \ddt \varphi_3 + g_\mr{D}(\varphi_3) \varphi_3 - i_\mr{s} = 0 \quad
    \text{and} \quad \begin{bmatrix}
        \varphi_1\\
        \varphi_2\\
        i_\mr{L}
    \end{bmatrix} = \begin{bmatrix}
        \varphi_2 + R i_\mr{s}\\
        \varphi_3 + L \ddt i_\mr{s}\\
        i_\mr{s}
    \end{bmatrix} \label{eq:do2}.
\end{align}
A key observation at this point is that all basis functions of the second step are also constant, as was the case for the first step. Although we did not continue the abstract decoupling process for MNA in \autoref{subsec:dcmna}, this is indeed a general phenomenon. The decoupling can be organized based on topological conditions only\cite{jansen2014}, which agrees with the intuition given by the index result from \autoref{sec:mna}. % \cite[Section 7.1]{jansen2014}
We remark that a similar result for circuits including controlled sources exists, however only for the first stage of the decoupling, as it is framed in the context of semi-explicit methods for which one only requires a DAE in semi-explicit form\cite{jansen2014}. An extension of this result to circuits of higher index is within the scope of further research.

\section{Index-aware learning}
\label{sec:ial}
In the following, we outline how to use the dissection concept in the context of machine learning. The workflow is illustrated in \autoref{fig:ial} and directly follows the structure of the dissection concept. We give a general description in a first step, followed by examples using the two circuits from \autoref{sec:dc}.
\begin{enumerate}
    \item Our approach begins by decoupling a given DAE into an ODE and a purely algebraic equation (AE) following the main steps of the dissection concept.

    \item Afterwards, only the differential degrees of freedom (DOFs) of the ODE are learned. For a DAE of index $\nu$, these would be the entries of $\tilde{\mathbf{x}}_\nu$ using the notation of \autoref{sec:dc}.

    \item The remaining algebraic DOFs $\bar{\mathbf{x}}_\nu$ may then be reconstructed using the algebraic equation.
\end{enumerate}
We remark that the identification of the differential DOFs and thus the advantages of the second point are in principle available for any Spice based simulator via the topological decoupling\cite{jansen2014}, whereas the reconstruction of the algebraic DOFs requires an additional implementation. However, it is possible to trade some of the accuracy of our approach for an easier implementation, see \autoref{subsec:aa}.
\begin{figure}
    \begin{center}
        \includegraphics[width=.7\textwidth]{ial}
    \end{center}
    \caption{Schematic workflow of index-aware learning.}
    \label{fig:ial}
\end{figure}

In terms of the example circuits from \autoref{subsec:dcmna} the workflow amounts to the following: for the first example we only consider $\varphi_3$ and $i_\mr{L}$ as variables that need to be learned and for the second example solely $\varphi_3$ is left. All remaining degrees of freedom may be recovered using \eqref{eq:do1} or \eqref{eq:do2} respectively. Thus the learning effort is already reduced quite significantly in these two examples; from five to two variables in the first and from four down to one variable in the second example. Another key benefit comes from the fact that the reconstructed algebraic variables exactly fulfill the inherent constraints of the DAE. This means that even though the learned solution variables (think of $\varphi_3$ for example) are only approximations, the reconstructions (e.g. $\varphi_1$ or $\varphi_2$) will still be consistent. This may be of great importance for systems where the physical interpretability of the solution depends on it satisfying the constraints.

There is yet another, maybe less expected, benefit that might occur. Looking at the decoupled system from \eqref{eq:do2} we find that the resistance parameter $R$ and the inductance parameter $L$ only appear in the algebraic equation. In terms of our original goal of speeding up design optimization or uncertainty quantification, where a lot of solutions for varying parameter values are required, this means the following: instead of needing to solve the full system for a given combination of $R$ and $L$, we can instead merely solve the algebraic equation to obtain the full solution. While the algebraic equation might be more complicated than the simple linear relations of \eqref{eq:do1} and \eqref{eq:do2} in general, solving it is almost certainly much faster than having to integrate the entire system in time. This becomes an even bigger advantage when knowledge about the solution is only required for specific points in time, since the algebraic equation may be solved pointwise.

Lastly, we emphasize that the approach is independent of the particular machine learning technique that is used for predicting the differential DOFs. Thus methods developed especially for ODEs may be employed and exchanged depending on the problem at hand.

\section{Numerical examples}
\label{sec:ne}
Before we present numerical results regarding the two example circuits introduced in \autoref{subsec:dcmna}, we want to provide some background on our machine learning method of choice, Gaussian processes (GPs), and the particular learning strategy we employ.

\paragraph{Gaussian processes}
The following brief introduction is based on the textbook of Rasmussen and Williams\cite{rasmussen2006} and we refer to the book itself for more details. We consider the problem of approximating one component $x(t)$ of the DAE solution $\mb{x}(t)$, compare \eqref{eq:dae}, based on observations
\begin{align*}
    O = \big\{ (t_i, x_i): 1 \leq i \leq N \big\}.
\end{align*}
We will focus on the one-dimensional case for clarity of exposition, however we remark that the ideas extend to the case where the solution component $x(t, \mb{p})$ also depends on the parameters and thus more than one variable, compare also the textbook\cite{rasmussen2006}. A GP suitable for this problem is defined by a (prior) mean function $m: \mathbb{R} \to \mathbb{R}$ and covariance function $k: \mathbb{R} \times \mathbb{R} \to \mathbb{R}$. The learning problem is then tackled using Bayesian inference, such that one aims to obtain the posterior distribution $\hat{x}(t)$, given the observations $O$ and a point $t$, where $x$ is to be predicted. A particular feature of GPs is that this posterior process, under suitable assumptions, turns out to be another GP with posterior mean and covariance\cite{basak2021}
\begin{subequations}
    \label{eq:gp}
    \begin{align}
        \hat{m}(t) &= m(t) + \mb{w}(t)^\T (\mb{x} - \mb{m}) \label{eq:gpm}\\
        \hat{k}(t) &= k(t, t) - \mb{w}(t)^\T \mb{k}(t) \label{eq:gpv},
    \end{align}
\end{subequations}
where $\mb{m} \coloneqq [m(t_1), \dotsc, m(t_N)]^\T$ denotes the prior mean function evaluated at the observations, $\mb{k}(t) \coloneqq [k(t, t_1), \dotsc, k(t, t_N)]^\T$ similarly denotes the pairwise evaluation of the covariance function using the prediction point $t$ and the observations, and $\mb{x} \coloneqq [x_1, \dotsc, x_N]^\T$ are the observed function values. The weights $\mb{w}(t)$ are given by the solution of
\begin{align*}
    \big( \mb{K} + \sigma^2 \mb{I} \big) \mb{w}(t) = \mb{k}(t),
\end{align*}
where $[\mb{K}]_{i,j} \coloneqq k(t_i, t_j)$ and $\sigma^2$ models i.i.d.~additive Gaussian noise on the observations. For a discussion on modeling choices where the assumptions leading to \eqref{eq:gp} are not fulfilled, see the review article by Swiler et al.\cite{swiler2020}.

The key model component influencing the learning process is the covariance (or kernel) function $k$, since it determines the approximation properties of the GP. It encodes prior knowledge about the function $x(t)$ that is to be learned, such as differentiability or characteristic length scales. We opt for a radial basis function kernel, given by
\begin{align*}
    k(t_i, t_j, \sigma_k, \ell) = \sigma_k^2 \exp \biggl(- \Big( \frac{t_i - t_j}{\ell} \Big)^2 \biggr),
\end{align*}
where $\sigma_k$ and the length scale $\ell$ are hyperparameters, which allow for better approximation capabilities of the GP. The kernel is selected to match the differentiability of the solution.

In practice, the mean function $m$ is often taken to be zero as the data is assumed standardized and the hyperparameters are then determined by minimizing the negative log likelihood\cite{basak2021}
\begin{align*}
    -\log \bigl( p(\mb{x} | \sigma, \sigma_k, \ell) \bigr) = \frac{1}{2} \big( \mb{x}^\T \mb{K}(\sigma, \sigma_k, \ell)^{-1} \mb{x} + \log \bigl( \det \mb{K}(\sigma, \sigma_k, \ell) \bigr) + N \log(2\pi) \big),
\end{align*}
where $[\mb{K}(\sigma, \sigma_k, \ell)]_{i,j} \coloneqq k(t_i, t_j, \sigma_k, \ell) + \sigma^2$.

\paragraph{Learning strategy}
The learning strategy aims to exploit one of the key features of GPs: that they provide both a mean prediction $\hat{m}$ and an associated variance estimate $\hat{k}$, as detailed in \eqref{eq:gp}. We use these properties in conjunction with a greedy sample selection strategy, that starts out with a small number of training data points and adds additional samples based on the variance estimate of the GP. This idea is not new, see again the textbook of Rasmussen and Williams\cite{rasmussen2006} for more references and details, however we still want to outline our particular approach to make the results better interpretable and reproducible.

Our implementation proceeds as follows:
\begin{enumerate}
    \item We select a grid of time points $T = \{ t_i: 1 \leq i \leq N_T \}$ and parameter values $P = \{ \mb{p}_i: 1 \leq i \leq N_P \}$, where each $\mb{p}_i$ represents a specific combination of parameter values, for which we want to model the solution of the DAE using a GP.
    \item We select a subset $D \subset T \times P$ and use the corresponding solutions for the initial training of a GP.
    \item As a termination criterion, we compute the posterior mean $\hat{m}(t, \mb{p})$ using \eqref{eq:gpm} in all grid points $(t, \mb{p}) \in T \times P$ and check whether the relative prediction error
    \begin{align*}
        e \coloneqq \frac{\| \mbh{m} - \mb{x} \|_2}{ \| \mb{x} \|_2},
    \end{align*}
    is below a desired tolerance, and if not, continue with 4. Here, $\mbh{m} \coloneqq [\hat{m}(t_1, \mb{p}_1), \dotsc, \hat{m}(t_{N_T}, \mb{p}_{N_P})]$ collects the mean predictions for all $(t, \mb{p}) \in T \times P$ and $\mb{x}$ contains the corresponding simulated solution values.
    \item We compute the variance prediction $\hat{k}(t, \mb{p})$ using \eqref{eq:gpv} for all $(t, \mb{p}) \in (T \times P) \setminus D$ and add the point of maximum variance to the training data set $D$.
    \item Finally, we retrain the model and continue with 3.
\end{enumerate}

We note that several improvements may be made to this strategy, such as properly maximizing the variance estimate in 4, instead of sampling on a discrete grid. The implementation is based on the STK toolbox\cite{stk}.

\paragraph{First example circuit}
We again consider the example of \autoref{fig:do1}, where we choose $R = \SI{500}{\ohm}$, $v_\mr{s}(t) = \sin(600 \pi t)\, \si{\volt}$ and the diode is modeled by $g_\mr{D}(\varphi_3) = 10^{-14} \big( e^{(\varphi_3 / \SI{26}{\mV})} - 1 \big)\, \si{\siemens}$. The remaining parameter values are chosen according to the greedy sample selection strategy with $\SI{1}{\milli\henry} \leq L \leq \SI{3}{\milli\henry}$ and $\SI{100}{\nano\farad} \leq C \leq \SI{300}{\nano\farad}$, i.e.~$\mb{p} = [L\ C]$ in the context of \autoref{sec:int}. The starting points for the strategy are given by all combinations of the boundary values for $L$ and $C$ together with $t = \SI{0}{\ms}$ and $t = \SI{10}{\ms}$ once for each combination. Recalling \eqref{eq:do1}, we see that $\varphi_3$ and $i_\mr{L}$ have to be learned as the differential DOFs. In addition to these two, we also consider the algebraic variable $\varphi_2$ and we will compare the accuracy of learning $\varphi_2$ directly to recovering it from the two differential variables. \autoref{fig:do1_solution} shows the solution for $L = \SI{1.7}{\milli\henry}$ and $C = \SI{220}{\nano\farad}$, a combination which is not part of the training data.
\begin{figure}[t]
    \begin{center}
        \includegraphics[width=.4\textwidth]{do1_solution_phi} \hspace{1.5cm} \includegraphics[width=.4\textwidth]{do1_solution_i_L}
    \end{center}
    \caption{Solution of the first example circuit for $L = \SI{1.7}{\milli\henry}$ and $C = \SI{220}{\nano\farad}$.}
    \label{fig:do1_solution}
\end{figure}

\begin{figure}[b]
    \begin{center}
        \includegraphics[width=.4\textwidth]{do1_convergence}
    \end{center}
    \caption{Convergence of the relative prediction error $e$, for the variables shown in \autoref{fig:do1_solution}. The final accuracies correspond to approximately 180 different combinations of $L$ and $C$ for $\varphi_3$, 39 combinations for $\varphi_2$ and 37 for $i_\mr{L}$.}
    \label{fig:do1_convergence}
\end{figure}
In \autoref{fig:do1_convergence}, we see the convergence of the relative prediction errors with respect to the number of samples used by the greedy selection strategy. We observe that, depending on the qualitative complexity of the dynamics, the solution variables take different numbers of samples to reach the same prediction accuracy. It should be noted however, that the total number of samples $N_D$ also includes the number of time points that are sampled, and not only the number of distinct parameter combinations (simulations). The latter are listed in the caption of \autoref{fig:do1_convergence} and turn out to be considerably smaller. One may also note that we only execute the learning strategy up to a relatively large tolerance of $10^{-3}$. This is due to the fact that both the optimization of the hyperparameters, as well as the computation of the posterior mean and variance, scale badly for conventional GPs, leading to large computation times. Remedies for this exist, see e.g.~the book by Rasmussen and Williams\cite{rasmussen2006}, however this issue lies outside the scope of this article as our approach may be combined with any machine learning method of choice.
\begin{figure}[t]
    \begin{center}
        \includegraphics[width=.4\textwidth]{do1_error_phi_3} \hspace{1.5cm} \includegraphics[width=.4\textwidth]{do1_error_i_L}
    \end{center}
    \caption{Differences between the mean predictions ($\hat{\varphi}_3$ and $\hat{i}_\mr{L}$) and the corresponding simulations for the differential variables when considering $L = \SI{1.7}{\milli\henry}$ and $C = \SI{220}{\nano\farad}$. The predictions correspond to the smallest relative errors from \autoref{fig:do1_convergence}.}
    \label{fig:do1_error_d}
\end{figure}

\begin{figure}[b]
    \begin{center}
        \includegraphics[width=.4\textwidth]{do1_error_phi_2} \hspace{1.5cm} \includegraphics[width=.4\textwidth]{do1_consistency}
    \end{center}
    \caption{Differences between the mean prediction $\hat{\varphi}_2$, reconstruction $\bar{\varphi}_2$ and the corresponding simulation of the algebraic variable $\varphi_2$ (left) and the respective consistency errors (right). The predictions correspond to $L = \SI{1.7}{\milli\henry}$, $C = \SI{220}{\nano\farad}$ and the smallest relative error from \autoref{fig:do1_convergence}.}
    \label{fig:do1_error_a_consistency}
\end{figure}
The differences between the mean predictions and simulations for the differential DOFs, when using the predictions belonging to the smallest relative errors from \autoref{fig:do1_convergence}, are shown in \autoref{fig:do1_error_d}. The results again correspond to $L = \SI{1.7}{\milli\henry}$ and $C = \SI{220}{\nano\farad}$, and we observe that the differences are in line with the relative prediction errors of \autoref{fig:do1_convergence}. The differences between the mean prediction $\hat{\varphi}_2$, reconstruction $\bar{\varphi}_2$ and the simulation of the algebraic variable $\varphi_2$ are highlighted on the left of \autoref{fig:do1_error_a_consistency}. The results again correspond to the same parameter values $L = \SI{1.7}{\milli\henry}$ and $C = \SI{220}{\nano\farad}$. We observe that there is not much difference between the mean prediction and reconstruction, however the reconstruction does appear to have a slight advantage in terms of accuracy. Here, one should note that although the reconstruction is exact up to the accuracy of solving the algebraic equation \eqref{eq:do1}, it still contains the error from learning the differential variables, hence the overall difference between $\bar{\varphi}_2$ and $\varphi_2$. To better quantify the difference between the learned and reconstructed solutions, we therefore introduce consistency errors $\hat{e}(t)$ and $\bar{e}(t)$ based on \eqref{eq:do1}
\begin{align*}
    \hat{e}(t) \coloneqq \begin{Vmatrix}
        \hat{\varphi}_1 - v_\mr{s}\\
        \hat{\varphi}_2 + R \hat{i}_\mr{L} - v_\mr{s}\\
        \hat{i}_\mr{V} + i_\mr{L}
    \end{Vmatrix}_2, \quad \bar{e}(t) \coloneqq \begin{Vmatrix}
        \bar{\varphi}_1 - v_\mr{s}\\
        \bar{\varphi}_2 + R \hat{i}_\mr{L} - v_\mr{s}\\
        \bar{i}_\mr{V} + i_\mr{L}
    \end{Vmatrix}_2,
\end{align*}
where $\hat{\cdot}$ indicates learned quantities and $\bar{\cdot}$ refers to reconstructed quantities. For a consistent solution obeying all algebraic constraints, both errors are identically zero. The right of \autoref{fig:do1_error_a_consistency} shows these consistency errors when using the same predictions and reconstructions as on the left. We observe a very clear improvement of the reconstructions (consistency error on the order of machine precision) over the directly learned predictions (consistency error as large as $10^{-3}$). While the central benefit is of course this adherence to the constraints, only needing to learn two of the five solution variables also represents a significant reduction of the learning effort in this case.

\paragraph{Second example circuit}
For the second example from \autoref{fig:do2}, we select $i_\mr{s}(t) = 10^{-4} \sin(400 \pi t)\, \si{\volt}$, while all other parameters and the learning strategy remain the same. Recalling \eqref{eq:do2}, we find that $\varphi_3$ is the only differential variable that needs to be learned to reconstruct the remaining algebraic DOFs. We again focus on $\varphi_2$ as an algebraic variable to obtain a comparative example.
\begin{figure}[b]
    \begin{center}
        \includegraphics[width=.4\textwidth]{do2_solution_phi} \hspace{1.5cm} \includegraphics[width=.4\textwidth]{do2_convergence}
    \end{center}
    \caption{Solution of the second example circuit for $L = \SI{1.7}{\milli\henry}$ and $C = \SI{220}{\nano\farad}$ (left) and convergence of the relative prediction error for the same variables (right). The final accuracies correspond to 31 distinct combinations of $L$ and $C$ for $\varphi_3$ and 49 for $\varphi_2$.}
    \label{fig:do2_solution_convergence}
\end{figure}

Considering similar numerical studies as for the first example, the left panel of \autoref{fig:do2_solution_convergence} shows the solutions of $\varphi_3$ and $\varphi_2$ for $L = \SI{1.7}{\milli\henry}$ and $C = \SI{220}{\nano\farad}$, which are again not part of the training data. The right plot shows the convergence of the relative prediction errors with respect to the total number of samples $N_D$. We again note that the number of distinct parameter combinations is significantly smaller, as listed in the caption. We also see that the relative errors behave similarly for both variables in this case, which is to be expected given the similar outlook of their dynamics in the left plot.
\begin{figure}[t]
    \begin{center}
        \includegraphics[width=.4\textwidth]{do2_error_phi_3} \hspace{1.5cm} \includegraphics[width=.4\textwidth]{do2_error_phi_2}
    \end{center}
    \caption{Differences between the mean prediction and simulation for $\varphi_3$ (left) and between the mean prediction, reconstruction and simulation for $\varphi_2$ (right). The predictions are again made for $L = \SI{1.7}{\milli\henry}$ and $C = \SI{220}{\nano\farad}$ and correspond to the smallest relative errors from \autoref{fig:do2_solution_convergence}.}
    \label{fig:do2_error}
\end{figure}

\begin{figure}[b]
    \begin{center}
        \includegraphics[width=.4\textwidth]{do2_consistency}
    \end{center}
    \caption{Consistency errors $\hat{e}(t)$ and $\bar{e}(t)$ corresponding to the results from \autoref{fig:do2_error}. The additional term $\check{e}(t)$ refers to the consistency error of the alternative approach, when using the predictions and Euler corrections outlined in \autoref{fig:do2_euler}.}
    \label{fig:do2_consistency}
\end{figure}
At this point we also return to the discussion from \autoref{sec:ial} about parameters only appearing in the algebraic equation. During the learning process, the greedy strategy requested 21 unique values for $C$, all of which required full simulations according to \eqref{eq:do2}. The 7 unique values for $L$, that were requested for learning $\varphi_2$, did not require full simulations however, but rather could be reconstructed from \eqref{eq:do2}. We again emphasize that time is also included as a parameter within the sample selection strategy, such that the reconstruction not only constitutes a reduction from solving a nonlinear DAE in four variables to merely evaluating a scalar equation, but that this scalar equation furthermore only needs to be evaluated at the particular points in time that are requested by the strategy, rather than at all time points of the solution.

The difference between the mean prediction and simulation of $\varphi_3$, for $L = \SI{1.7}{\milli\henry}$ and $C = \SI{220}{\nano\farad}$, is shown on the left of \autoref{fig:do2_error}, while the right shows the differences between the mean prediction, reconstruction and simulation of $\varphi_2$. The predictions again correspond to the smallest relative errors from \autoref{fig:do2_solution_convergence} and the differences are of the same order. In this case, the reconstruction performs similar or even slightly worse compared to the mean prediction when only looking at the difference. However it still constitutes a reduction in learning effort, from four solution variables down to only one, and most importantly it adheres to the algebraic constraints of the DAE. To further illustrate this point, we again take a look at the consistency errors, now redefined based on \eqref{eq:do2}
\begin{align*}
    \hat{e}(t) \coloneqq \begin{Vmatrix}
        \hat{\varphi}_1 - \hat{\varphi}_2 + R i_\mr{s}\\
        \hat{\varphi}_2 - \hat{\varphi}_3 + L \ddt i_\mr{s}\\
        \hat{i}_\mr{L} - i_\mr{s}
    \end{Vmatrix}_2, \quad \bar{e}(t) \coloneqq \begin{Vmatrix}
        \bar{\varphi}_1 - \bar{\varphi}_2 + R i_\mr{s}\\
        \bar{\varphi}_2 - \hat{\varphi}_3 + L \ddt i_\mr{s}\\
        \bar{i}_\mr{L} - i_\mr{s}
    \end{Vmatrix}_2.
\end{align*}
In our example, the algebraic equation from \eqref{eq:do2} gives an explicit description of the algebraic variables once more, such that the reconstruction is accurate up to machine precision, compare $\bar{e}(t)$ in \autoref{fig:do2_consistency}. When learning all solution variables individually however, we observe a much larger maximum value of around $10^{-3}$ for the consistency error $\hat{e}(t)$ across all the predicted points in time. In general, the algebraic equation may be nonlinear such that the reconstruction still leads to a consistency error $\bar{e}(t)$ greater than machine precision, depending on the accuracy of the nonlinear system solver.

\subsection{Alternative approach}
\label{subsec:aa}
In a last step, we want to provide an alternative approach to reducing the consistency error, while avoiding an extra implementation to handle the algebraic constraints. The idea is based on the insight that two successive implicit Euler steps suffice to bring the algebraic variables into a consistent state with the differential ones, when starting with inconsistent initial values\cite{garcia2022}. In our context, one can then simply learn all solution variables independently, and when making a prediction, select a time point slightly before the intended prediction point, such that two small implicit Euler steps may be performed as a correction to obtain a consistent solution.
\begin{figure}[t]
    \begin{center}
        \includegraphics[width=.4\textwidth]{do2_euler_phi_3} \hspace{1.5cm} \includegraphics[width=.4\textwidth]{do2_euler_phi_2}
    \end{center}
    \caption{Differences between the mean predictions, Euler corrections and simulations for $\varphi_3$ (left) and $\varphi_2$ (right). The predictions are made for $L = \SI{1.7}{\milli\henry}$ and $C = \SI{220}{\nano\farad}$ and correspond to the smallest relative errors from \autoref{fig:do2_solution_convergence}.}
    \label{fig:do2_euler}
\end{figure}

This alternative approach has the advantage that it can be directly applied using simulations from any SPICE-like software, and using any machine learning method for the predictions, without worrying about the reconstruction of the algebraic variables. However it may not achieve the same level of accuracy as the approach outlined in \autoref{sec:ial} in practice, depending on the Euler step size and the accuracy of the nonlinear system solver\cite{garcia2022}. Furthermore, it requires predictions for all solution variables, which represents unnecessary computational cost, and there is no chance for parameters to shift completely to the algebraic equations, compare \autoref{sec:ial}.

\autoref{fig:do2_euler} shows the differences between the mean predictions ($\hat{\varphi}_3$, $\hat{\varphi}_2$), Euler corrections ($\check{\varphi}_3$, $\check{\varphi}_2$) and the simulations for the second example, using the same parameters as in \autoref{fig:do2_error}. We see that the two Euler steps have almost no effect on the differential variable $\varphi_3$, as expected. For the algebraic variable $\varphi_2$ however, the mean predictions $\hat{\varphi}_2$ and the corrected predictions after two Euler steps $\check{\varphi}_2$, differ significantly, with the latter resembling the reconstruction from \autoref{fig:do2_error}. Also defining a consistency error $\check{e}(t)$ based on \eqref{eq:do2} in this case
\begin{align*}
    \check{e}(t) \coloneqq \begin{Vmatrix}
        \check{\varphi}_1 - \check{\varphi}_2 + R i_\mr{s}\\
        \check{\varphi}_2 - \check{\varphi}_3 + L \ddt i_\mr{s}\\
        \check{i}_\mr{L} - i_\mr{s}
    \end{Vmatrix}_2
\end{align*}
and comparing it to the other approaches, see \autoref{fig:do2_consistency}, shows that this alternative approach represents a compromise in terms of accuracy in the algebraic constraints. The maximum value of the consistency error $\check{e}(t)$ of the corrected predictions comes out around $10^{-7}$, about four orders of magnitude smaller compared to before the correction. While it is theoretically possible to also force the error down to machine precision in this case, other factors become prohibitive as mentioned earlier.

\section{Conclusions and future research}
\label{sec:cfr}
This article introduced a new method for learning the time and parameter dependent behavior of electrical circuits. The method assumes the circuit to be modeled using MNA and then exploits the structure of MNA to improve the learning process by splitting it into a learning and a reconstruction stage, compare \autoref{fig:ial}. It achieves this by decoupling the underlying DAE into an ODE and a purely algebraic equation. Benefits of the method include a reduction in the number of variables that are to be learned during the learning stage and the exact adherence of the learned solution to the inherent constraints of the circuit model after the reconstruction stage. Two numerical examples illustrated both benefits. The examples also showed that the exact recovery of the constraints might improve the accuracy of the learned solutions. Furthermore, additional computational savings may be possible as some of the parameters of interest only appear in the reconstruction stage, which entirely avoids the need for training data with varying values of these parameters. We emphasize that the approach is independent of the machine learning method used during the learning stage, such that the learning method may be chosen according to the problem at hand. Lastly, we showcased an alternative approach that avoids the need for the decoupling and the associated implementation, while still improving the adherence of the learned solution to the algebraic constraints of the DAE.

Multiple extensions are possible within this index-aware learning framework. A natural first step could be to make use of the topological decomposition that was hinted at in \autoref{sec:dc}. This would pave the way for the inclusion of controlled sources within the workflow, and thus the industrial use of the approach, with the idea being the extension of the topological decomposition to also allow for controlled sources. Focusing in on the idea of adhering to physically meaningful constraints, one could work on extending the concept to a charge conserving variant of MNA, to potentially guarantee charge conservation even for the learned solutions. Yet another direction may be the application of index-aware learning to DAEs arising elsewhere, e.g.~in modified loop analysis. (The dissection concept applies to a more general class of DAEs, as \autoref{sec:dc} showed.) Finally, separate but related work may focus on improving the learning stage by developing new methods that are especially suited for learning ODEs.

% \begin{table}
%     \caption{}
%     \begin{center}
%         \begin{tabular}{}
%             \toprule
%             & &\\
%             \midrule
%             & &\\
%             \bottomrule
%         \end{tabular}
%     \end{center}
%     \begin{tablenotes}
%         \item notes
%     \end{tablenotes}
% \end{table}

% \begin{algorithm}
%     \begin{algorithmic}
%     \end{algorithmic}
% \end{algorithm}

\section*{Acknowledgments}
This work is supported by the Graduate School CE within the Centre for Computational Engineering at Technische Universit{\"a}t Darmstadt and the ECSEL Joint Undertaking (JU) under grant agreement No.~101007319. The JU receives support from the European Union's Horizon 2020 research and innovation programme and the Netherlands, Hungary, France, Poland, Austria, Germany, Italy and Switzerland. Note that this work only reflects the authors' views and that the JU is not responsible for any use that may be made of the information it contains.

% \subsection*{Author contributions}

% \subsection*{Financial disclosure}

\subsection*{Conflict of interest}
The authors declare no potential conflict of interests.

% \section*{Supporting information}

% \appendix
% \section{}

% \nocite{*}
\bibliography{biblio}

\clearpage

\section*{Author biography}
\begin{biography}{\includegraphics[width=66pt,height=86pt]{cortes}}{\textbf{Idoia Cortes Garcia} received her Bachelor in mathematics and Master in modelling for science and engineering from the Autonomous University of Barcelona. She obtained her Ph.D. in electrical engineering from the Technical University of Darmstadt in 2020. Since 2022 she is an assistant professor at the group of Dynamics and Control from the department of mechanical engineering at Eindhoven University of Technology. Her research interests include coupled multiphysical dynamical systems, differential algebraic equations, efficient time domain (co-)simulation methods and hybrid modelling approaches.}
\end{biography}
\vspace{0.5cm}
\begin{biography}{\includegraphics[width=66pt,height=86pt]{forster}}{\textbf{Peter Förster} received his B.Sc. and M.Sc. degrees in electrical engineering from Technical University of Darmstadt. Currently, he is a doctoral researcher at the centre for analysis, scientific computing and applications at Eindhoven University of Technology and in the computational electromagnetics group at Technical University of Darmstadt. His research interests include circuit simulation, differential-algebraic equations, machine learning and digital twins.}
\end{biography}
\vspace{0.5cm}
\begin{biography}{\includegraphics[width=66pt,height=86pt]{jansen}}{\textbf{Lennart Jansen} received his M.Sc. degree in mathematics from Universität zu Köln and his Ph.D. degree in mathematics from Universität zu Berlin. Afterwards, he held a post-doc position at Heinrich-Heine-Universität in Düsseldorf before starting work in the industry. One of the main research topics he worked on in the industry was AI-based recommendations for design improvement in the automobile industry. He switched to the photovoltaic industry in 2021 and currently works there on optimal control for combined electric-water-gas networks.}
\end{biography}
\vspace{0.5cm}
\begin{biography}{\includegraphics[width=66pt,height=86pt]{schilders}}{\textbf{Wil Schilders} received the M.Sc. degree in mathematics from Radboud University Nijmegen, Nijmegen, The Netherlands, in 1978, and the Ph.D. degree from the Trinity College Dublin, Dublin, Ireland, in 1980. He has been working in the electronics industry with Philips Research Laboratories, Eindhoven, The Netherlands, since 1980, and NXP, since 2006, where he developed algorithms for simulating semiconductor devices, electronic circuits, organic light emitting diodes, and electromagnetic problems (TV tubes, interconnects, and magnetic resonance imaging). Since 1999, he has been a Part-Time Professor of numerical mathematics for industry with the Technical University of Eindhoven, Eindhoven. He is currently the Managing Director of the Platform for Mathematics in The Netherlands. He is also the President of the European Consortium for Mathematics in Industry.}
\end{biography}
\vspace{0.5cm}
\begin{biography}{\includegraphics[width=66pt,height=86pt]{schoeps}}{\textbf{Sebastian Schöps} received the M.Sc. degree in business mathematics and the joint Ph.D. degree from Bergische Universität Wuppertal and Katholieke Universiteit Leuven in mathematics and physics. He was appointed as a Professor of Computational Electromagnetics at Technische Universität Darmstadt within the interdisciplinary center of computational engineering, in 2012. His current research interests include coupled multi-physical problems, bridging computer aided design and simulation, parallel algorithms for high performance computing, digital twins, uncertainty quantification, and machine learning.}
\end{biography}

\end{document}

### 2023-11-30 doc/fig/ial.tex
index-aware-learning

% \begin{tikzpicture}
%     \node[rectangle, rounded corners, draw=black, semithick] (dae) at (0,0) {DAE};
%     \node[rectangle, rounded corners, draw=black, semithick] (ode) at (-2,-2) {\textcolor{TUDa-1a}{ODE}};
%     \node[rectangle, rounded corners, draw=black, semithick] (ae) at (2,-2) {\textcolor{TUDa-9a}{AE}};
%     \node[rectangle, rounded corners, draw=black, semithick] (dd) at (-2,-4) {\textcolor{TUDa-1a}{Differential DOFs}};
%     \node[rectangle, rounded corners, draw=black, semithick] (ad) at (2,-7) {\textcolor{TUDa-9a}{Algebraic DOFs}};

%     \node[circle, fill=TUDa-9a, minimum size=3pt, inner sep=0] (ddae) at (2,-4) {};

%     \draw[->, >=latex, semithick, color=black] (dae) -- (ode) node[above, midway, sloped] {decouple};
%     \draw[->, >=latex, semithick, color=black] (dae) -- (ae) node[above, midway, sloped] {decouple};
%     \draw[<-, >=latex, semithick, color=TUDa-1a, dashed] (dd) -- (ode) node[above, midway, sloped] {learn};

%     \draw[-, semithick, color=TUDa-9a, dotted] (dd) -- (ddae);
%     \draw[-, semithick, color=TUDa-9a, dotted] (ae) -- (ddae);
%     \draw[<-, >=latex, semithick, color=TUDa-9a, dotted] (ad) -- (ddae) node[above, midway, sloped] {reconstruct};
% \end{tikzpicture}

% landscape
\begin{tikzpicture}
    \node[rectangle, rounded corners, draw=black, semithick] (dae) at (0,0) {DAE};
    \node[rectangle, rounded corners, draw=black, semithick] (ode) at (2,-2) {\textcolor{TUDa-1a}{ODE}};
    \node[rectangle, rounded corners, draw=black, semithick] (ae) at (2,2) {\textcolor{TUDa-9a}{AE}};
    \node[rectangle, rounded corners, draw=black, semithick] (dd) at (5.5,-2) {\textcolor{TUDa-1a}{Differential DOFs}};
    \node[rectangle, rounded corners, draw=black, semithick] (ad) at (9.5,2) {\textcolor{TUDa-9a}{Algebraic DOFs}};

    \node[circle, fill=TUDa-9a, minimum size=3pt, inner sep=0] (ddae) at (5.5,2) {};

    \draw[->, >=latex, semithick, color=black] (dae) -- (ode) node[above, midway, sloped] {decouple};
    \draw[->, >=latex, semithick, color=black] (dae) -- (ae) node[above, midway, sloped] {decouple};
    \draw[<-, >=latex, semithick, color=TUDa-1a, dashed] (dd) -- (ode) node[above, midway, sloped] {learn};

    \draw[-, semithick, color=TUDa-9a, dotted] (dd) -- (ddae);
    \draw[-, semithick, color=TUDa-9a, dotted] (ae) -- (ddae);
    \draw[<-, >=latex, semithick, color=TUDa-9a, dotted] (ad) -- (ddae) node[above, midway, sloped] {reconstruct};
\end{tikzpicture}

### 2023-11-30 code/diode_oscillator_doe.m

@@ -1,0 +1,173 @@
+problem  = 1; % 1, 2
+variable = 'i_V'; % phi_3, i_L, phi_2, phi_1, i_V
+tol      = [1e-2 5e-3 1e-3]; % 1e-2, 5e-3, 1e-3
+
+N_t = 1000;
+T_0 = 0;
+T   = 1e-2;
+
+% all available data
+L = [1e-3:0.1e-3:3e-3];
+C = [100e-9:10e-9:300e-9];
+t = linspace(T_0, T, N_t);
+
+% initial points
+t_i = [t(1); t(end); t(1); t(end); t(1); t(end); t(1); t(end)];
+L_i = [L(1); L(1); L(end); L(end); L(1); L(1); L(end); L(end)];
+C_i = [C(1); C(1); C(1); C(1); C(end); C(end); C(end); C(end)];
+
+if (problem == 1)
+    filename = ['do1_doe_N_t=' num2str(N_t) '.dat'];
+elseif (problem == 2)
+    filename = ['do2_doe_N_t=' num2str(N_t) '.dat'];
+end
+data = dlmread(filename);
+data = data(2:end,:);
+
+t_d = data(:,1);
+L_d = data(:,2);
+C_d = data(:,3);
+X   = [t_d L_d C_d];
+
+phi_3_d = data(:,4);
+
+R   = 500;
+i_s = @(t) 1e-4*sin(2*pi*200*t);
+if (problem == 1)
+    i_L_d   = data(:,5);
+    phi_2_d = data(:,6);
+    i_V_d   = -i_L_d;
+elseif (problem == 2)
+    phi_2_d = data(:,5);
+    phi_1_d = phi_2_d + R*i_s(t_d);
+end
+
+if (strcmp(variable, 'phi_3'))
+    y = phi_3_d;
+elseif (strcmp(variable, 'i_L'))
+    y = i_L_d;
+elseif (strcmp(variable, 'phi_2'))
+    y = phi_2_d;
+elseif (strcmp(variable, 'phi_1'))
+    y = phi_1_d;
+elseif (strcmp(variable, 'i_V'))
+    y = i_V_d;
+end
+
+i_i = NaN(length(t_i), 1);
+for i=1:length(i_i)
+    i_i(i) = find((t_i(i) == t_d) & (L_i(i) == L_d) & (C_i(i) == C_d));
+end
+
+f = stk_model('stk_gausscov_aniso', 3);
+
[…]

### 2023-11-30 code/diode_oscillator_doe_av.m

@@ -1,0 +1,151 @@
+problem  = 1; % 1, 2
+variable = 'phi_1'; % phi_1, i_L
+tol      = [1e-3]; % 1e-3
+
+N_t = 1000;
+T_0 = 0;
+T   = 1e-2;
+
+% all available data
+t = linspace(T_0, T, N_t);
+
+% initial points (need more than boundary)
+t_i = [t(1); t(500); t(end)];
+
+if (problem == 1)
+    filename = ['do1_doe_N_t=' num2str(N_t) '.dat'];
+elseif (problem == 2)
+    filename = ['do2_doe_N_t=' num2str(N_t) '.dat'];
+end
+data     = dlmread(filename);
+data     = data(2:end,:);
+
+t_d = data(1:N_t,1);
+X   = [t_d];
+
+R   = 500;
+v_s = @(t) sin(2*pi*300*t);
+i_s = @(t) 1e-4*sin(2*pi*200*t);
+if (strcmp(variable, 'phi_1'))
+    y = v_s(t_d);
+elseif (strcmp(variable, 'i_L'))
+    y = i_s(t_d);
+end
+
+i_i = NaN(length(t_i), 1);
+for i=1:length(i_i)
+    i_i(i) = find((t_i(i) == t_d));
+end
+
+f = stk_model('stk_gausscov_aniso', 1);
+
+f.lognoisevariance = NaN;
+% f.lognoisevariance = -Inf;
+tic;
+[theta, sigma] = stk_param_init(f, X(i_i,:), y(i_i));
+fprintf('\nstk_param_init: %d min\n', toc/60);
+
+tic;
+[f.param, f.lognoisevariance] = stk_param_estim(f, X(i_i,:), y(i_i), theta, sigma);
+% [f.param] = stk_param_estim(f, X, y, theta);
+fprintf('\nstk_param_estim: %d min\n', toc/60);
+
+tic;
+y_p = stk_predict(f, X(i_i,:), y(i_i), X);
+fprintf('\nstk_predict: %d min\n', toc/60);
+
+i_t = i_i;
+e_p = 1;
+for i_tol=1:length(tol)
+    fprintf('\ntol: %g\n', tol(i_tol));
+    if (problem == 1)
+        filename = ['do1_doe_i_t_' variable '_tol=' num2str(tol(i_tol)) '_N_t=' num2str(N_t) '.dat'];
+        if (exist(filename))
+            i_t = dlmread(filename);
+
[…]

### 2023-11-30 code/diode_oscillator_results.m

@@ -1,0 +1,481 @@
+problem = 2; % 1, 2
+
+plot_input       = 0; % 0, 1
+plot_convergence = 0; % 0, 1
+plot_solution    = 0; % 0, 1
+plot_consistency = 0; % 0, 1
+plot_euler       = 1; % 0, 1
+
+% only relevant for input/convergence
+variable = 'phi_3'; % phi_3, i_L, phi_2
+
+% only relevant for solution/consistency
+tol = 1e-3; % 1e-2, 5e-3, 1e-3
+
+if (plot_input == 1)
+    TOL = [1e-2, 5e-3, 1e-3];
+    N_t = 1000;
+
+    for tol=TOL
+        if (problem == 1)
+            filename = ['do1_doe_i_t_' variable '_tol=' num2str(tol) '_N_t=' num2str(N_t) '.dat'];
+            i_t      = dlmread(filename);
+
+            filename = ['do1_doe_N_t=' num2str(N_t) '.dat'];
+        elseif (problem == 2)
+            filename = ['do2_doe_i_t_' variable '_tol=' num2str(tol) '_N_t=' num2str(N_t) '.dat'];
+            i_t      = dlmread(filename);
+
+            filename = ['do2_doe_N_t=' num2str(N_t) '.dat'];
+        end
+
+        data = dlmread(filename);
+        data = data(2:end,:);
+
+        t = data(:,1);
+        L = data(:,2);
+        C = data(:,3);
+        X = [t L C];
+
+        if (problem == 1)
+            filename = ['do1_input_' variable '_tol=' num2str(tol) '.dat'];
+        elseif (problem == 2)
+            filename = ['do2_input_' variable '_tol=' num2str(tol) '.dat'];
+        end
+
+        fid = fopen(filename, 'w');
+        fprintf(fid, 'x  y\n');
+        dlmwrite(fid, X(i_t,2:3), 'delimiter', '  ', 'append', 'on');
+        fclose(fid);
+    end
+end
+
+if (plot_convergence == 1)
+    tol = [1e-2, 5e-3, 1e-3];
+    N_t = 1000;
+
+    N_s = NaN(length(tol), 1);
+    for it=1:length(tol)
+       if (problem == 1)
[…]

### 2023-11-30 code/do1_sc.m

@@ -1,0 +1,21 @@
+% syms, sym, simplify, jacobian, inv, null, rank, +, -, *, /, ^, .'
+pkg load symbolic
+
+syms phi_1 phi_2 phi_3 i_L i_V G C L g_D v_s
+
+M = [[0, 0, 0, 0, 0]; [0, 0, 0, 0, 0]; [0, 0, C, 0, 0]; [0, 0, 0, L, 0]; [0, 0, 0, 0, 0]];
+K = [[G, -G, 0, 0, 1]; [-G, G, 0, 1, 0]; [0, 0, g_D, -1, 0]; [0, -1, 1, 0, 0]; [-1, 0, 0, 0, 0]];
+f = [[0]; [0]; [0]; [0]; [-v_s]];
+
+% first basis functions
+Q = W = sym([[1, 0, 0]; [0, 1, 0]; [0, 0, 0]; [0, 0, 0]; [0, 0, 1]]);
+P = V = sym([[0, 0]; [0, 0]; [1, 0]; [0, 1]; [0, 0]]);
+
+% first stage
+Mt   = V.' * M * P;
+Kt_P = V.' * K * P;
+Kt_Q = V.' * K * Q;
+ft   = V.' * f;
+Kb_P = W.' * K * P;
+Kb_Q = W.' * K * Q;
+fb   = W.' * f;

### 2023-11-30 code/do2_sc.m

@@ -1,0 +1,47 @@
+% syms, sym, simplify, jacobian, inv, null, rank, +, -, *, /, ^, .'
+pkg load symbolic
+
+syms phi_1 phi_2 phi_3 i_L G C L g_D i_s i_sp
+
+M = [[0, 0, 0, 0]; [0, 0, 0, 0]; [0, 0, C, 0]; [0, 0, 0, L]];
+K = [[G, -G, 0, 0]; [-G, G, 0, 1]; [0, 0, g_D, -1]; [0, -1, 1, 0]];
+f = [[i_s]; [0]; [0]; [0]];
+
+% first basis functions
+Q = W = sym([[1, 0]; [0, 1]; [0, 0]; [0, 0]]);
+P = V = sym([[0, 0]; [0, 0]; [1, 0]; [0, 1]]);
+
+% first stage
+Mt   = V.' * M * P;
+Kt_P = V.' * K * P;
+Kt_Q = V.' * K * Q;
+ft   = V.' * f;
+Kb_P = W.' * K * P;
+Kb_Q = W.' * K * Q;
+fb   = W.' * f;
+
+% second basis functions
+Qb = Wb = sym([[1]; [1]])
+
+% Wb.' * Kb_P
+Qt  = sym([[1]; [0]])
+Qtp = sym([[0]; [0]]);
+Pt  = sym([[0]; [1]])
+Ptp = sym([[0]; [0]]);
+
+% xt_P = inv(Wb.' * Kb_P * Pt) * Wb.' * fb;
+xt_P  = i_s;
+xt_Pp = i_sp;
+
+% (Mt * Qt).';
+Wt = sym([[0]; [1]])
+Vt = sym([[1]; [0]])
+
+% second stage
+Mt_2  = Vt.' * Mt * Qt
+Kt_P2 = Vt.' * (Mt * Qtp + Kt_P * Qt)
+Kt_Q2 = Vt.' * Kt_Q * Qb
+ft_2  = Vt.' * (ft - Mt * Pt * xt_Pp - (Mt * Ptp + Kt_P * Pt) * xt_P)
+Kb_P2 = Wt.' * (Mt * Qtp + Kt_P * Qt)
+Kb_Q2 = Wt.' * Kt_Q * Qb
+fb_2  = Wt.' * (ft - Mt * Pt * xt_Pp - (Mt * Ptp + Kt_P * Pt) * xt_P)

### 2023-11-30 code/full_wave_rectifier.m

@@ -1,0 +1,150 @@
+samples     = 0; % 0, 1
+convergence = 1; % 0, 1
+
+t_0    = 0;
+t_f    = 5e-2;
+Deltat = 1e-5;
+N_t    = 1000;
+
+% T in K
+q   = 1.602e-19;
+k   = 1.38e-23;
+qk  = q/k;
+I_s = @(T) 1e-6 * (T./300).^3 .* exp( (T./300 - 1) .* 1.12*qk./T );
+g_D = @(v_D, T) I_s(T) .* qk./T .* exp(qk./T .* v_D);
+R   = 50;
+
+L_1  = 27.46e-6;
+L_12 = 27.57e-6;
+L_2  = 27.75e-6;
+
+% v_s = @(t) 325*sin(2*pi*50*t);
+v_s = @(t) 12*sin(2*pi*50*t);
+
+n_v = 4;
+n_L = 2;
+n_V = 1;
+
+A_R = [[0, 0, 0, 0, 0]; [1, 0, -1, 0, 0]; [-1, -1, 0, 0, 1]; [0, 0, 1, 1, -1]];
+A_L = [[1 0]; [0 1]; [0, 0]; [0, 0]];
+A_V = [[1]; [0]; [0]; [0]];
+
+G = @(x, T) [[g_D(x(2)-x(3), T(1)), 0, 0, 0, 0]; [0, g_D(-x(3), T(2)), 0, 0, 0]; [0, 0, g_D(x(4)-x(2), T(3)), 0, 0]; [0, 0, 0, g_D(x(4), T(4)), 0]; [0, 0, 0, 0, 1/R]];
+L = [[L_1 L_12]; [L_12 L_2]];
+
+M = [[zeros(n_v,n_v), zeros(n_v,n_L), zeros(n_v,n_V)]; [zeros(n_L,n_v), L, zeros(n_L,n_V)]; [zeros(n_V,n_v), zeros(n_V,n_L), zeros(n_V,n_V)];];
+K = @(x, T) [[A_R*G(x, T)*A_R.', A_L, A_V]; [-A_L.', zeros(n_L,n_L), zeros(n_L,n_V)]; [-A_V.', zeros(n_V,n_L), zeros(n_V,n_V)]];
+f = @(t) [zeros(n_v,1); zeros(n_L,1); v_s(t)];
+
+% first basis functions
+Q      = W      = zeros(7,5);
+Q(1,1) = W(1,1) = 1;
[…]

### 2023-11-30 code/full_wave_rectifier_doe.m

@@ -1,0 +1,205 @@
+variables = ['i_L_1'; 'i_L_2'; 'v_1'; 'v_2'; 'v_3'; 'v_4'; 'i_V'];
+
+t_0       = 0;
+t_f       = 5e-2;
+Deltat    = 1e-5;
+N_t       = 1000;
+epsilon_T = 1e-6;
+i_t_i     = [round(1:N_t/10:N_t) N_t];
+epsilon_t = 1e-16;
+w_1 = w_2 = 1;
+
+for iv = 1:length(variables)
+    variable = variables(iv,:);
+    t        = linspace(t_0, t_f, N_t);
+
+    % all sampling points
+    % T_1 = T_2 = T_3 = T_4 = [20:10:100] + 273.15;
+    T_1 = T_2 = [20:10:100] + 273.15;
+
+    % initial points
+    N_t_i = length(i_t_i);
+    % t_i   = repmat([t_0*ones(2^0, 1); t_f*ones(2^0, 1)], 2^2, 1);
+    % T_1_i = repmat([T_1(1)*ones(2^1, 1); T_1(end)*ones(2^1, 1)], 2^1, 1);
+    % T_2_i = repmat([T_2(1)*ones(2^2, 1); T_2(end)*ones(2^2, 1)], 2^0, 1);
+    t_i   = repmat(t(i_t_i)', 2^2, 1);
+    T_1_i = repmat([T_1(1)*ones(N_t_i*1, 1); T_1(end)*ones(N_t_i*1, 1)], 2^1, 1);
+    T_2_i = repmat([T_2(1)*ones(N_t_i*2, 1); T_2(end)*ones(N_t_i*2, 1)], 2^0, 1);
+    % T_3_i = repmat([T_3(1)*ones(2^3, 1); T_3(end)*ones(2^3, 1)], 2^1, 1);
+    % T_4_i = repmat([T_4(1)*ones(2^4, 1); T_4(end)*ones(2^4, 1)], 2^0, 1);
+
+    % define grid (becomes very large very quickly: true optimization worthwhile)
+    % N_t   = round((t_f - t_0) / Deltat) + 1;
+    N_T_1 = length(T_1);
+    N_T_2 = length(T_2);
+    % N_T_3 = length(T_3);
+    % N_T_4 = length(T_4);
+    X     = NaN(N_t*N_T_1*N_T_2, 3);
+    y     = NaN(N_t*N_T_1*N_T_2, 1);
+    for i_2 = 1:N_T_2
+        for i_1 = 1:N_T_1
[…]

### 2023-11-30 code/full_wave_rectifier_results.m

@@ -1,0 +1,286 @@
+plot_solution    = 0; % 0, 1
+plot_consistency = 1; % 0, 1
+
+variables = ['i_L_1'; 'i_L_2'; 'v_1'; 'v_2'; 'v_3'; 'v_4'; 'i_V'];
+tol       = [5e-3]; % 1e-2, 5e-3, 1e-3
+
+warning("off", "Octave:data-file-in-path");
+
+t_0       = 0;
+t_f       = 5e-2;
+Deltat    = 1e-5;
+N_t       = 1000;
+t         = linspace(t_0, t_f, N_t);
+epsilon_T = 1e-6;
+
+% T in K
+q   = 1.602e-19;
+k   = 1.38e-23;
+qk  = q/k;
+I_s = @(T) 1e-6 * (T./300).^3 .* exp( (T./300 - 1) .* 1.12*qk./T );
+g_D = @(v_D, T) I_s(T) .* qk./T .* exp(qk./T .* v_D);
+R   = 50;
+
+L_1  = 27.46e-6;
+L_12 = 27.57e-6;
+L_2  = 27.75e-6;
+
+v_s = @(t) 12*sin(2*pi*50*t);
+
+n_v = 4;
+n_L = 2;
+n_V = 1;
+
+A_R = [[0, 0, 0, 0, 0]; [1, 0, -1, 0, 0]; [-1, -1, 0, 0, 1]; [0, 0, 1, 1, -1]];
+A_L = [[1 0]; [0 1]; [0, 0]; [0, 0]];
+A_V = [[1]; [0]; [0]; [0]];
+
+G = @(x, T) [[g_D(x(2)-x(3), T(1)), 0, 0, 0, 0]; [0, g_D(-x(3), T(2)), 0, 0, 0]; [0, 0, g_D(x(4)-x(2), T(3)), 0, 0]; [0, 0, 0, g_D(x(4), T(4)), 0]; [0, 0, 0, 0, 1/R]];
+L = [[L_1 L_12]; [L_12 L_2]];
+
+M = [[zeros(n_v,n_v), zeros(n_v,n_L), zeros(n_v,n_V)]; [zeros(n_L,n_v), L, zeros(n_L,n_V)]; [zeros(n_V,n_v), zeros(n_V,n_L), zeros(n_V,n_V)];];
[…]

### 2023-11-30 code/fwr_sc.m

@@ -1,0 +1,40 @@
+% syms, sym, simplify, jacobian, inv, null, rank, +, -, *, /, ^, .'
+pkg load symbolic
+
+syms v_1 v_2 v_3 v_4 i_L_1 i_L_2 i_V
+syms L_1 L_2 L_12 g_D_1 g_D_2 g_D_3 g_D_4 R v_s
+syms g_D
+
+v   = [v_1; v_2; v_3; v_4];
+i_L = [i_L_1; i_L_2];
+
+n_v = length(v);
+n_L = length(i_L);
+
+A_R = [[0, 0, 0, 0, 0]; [1, 0, -1, 0, 0]; [-1, -1, 0, 0, 1]; [0, 0, 1, 1, -1]];
+A_L = [[1 0]; [0 1]; [0, 0]; [0, 0]];
+A_V = [[1]; [0]; [0]; [0]];
+
+G = [[g_D_1, 0, 0, 0, 0]; [0, g_D_2, 0, 0, 0]; [0, 0, g_D_3, 0, 0]; [0, 0, 0, g_D_4, 0]; [0, 0, 0, 0, 1/R]];
+% G = [[g_D, 0, 0, 0, 0]; [0, g_D, 0, 0, 0]; [0, 0, g_D, 0, 0]; [0, 0, 0, g_D, 0]; [0, 0, 0, 0, 1/R]];
+L = [[L_1 L_12]; [L_12 L_2]];
+
+M = [[zeros(n_v,n_v), zeros(n_v,n_L), zeros(n_v,1)]; [zeros(n_L,n_v), L, zeros(n_L,1)]; [zeros(1,n_v), zeros(1,n_L), zeros(1,1)];];
+K = [[A_R*G*A_R.', A_L, A_V]; [-A_L.', zeros(n_L,n_L), zeros(n_L,1)]; [-A_V.', zeros(1,n_L), zeros(1,1)]];
+f = [zeros(n_v,1); zeros(n_L,1); v_s];
+
+% first basis functions
+Q = W = null(M);
+P = V = sym([[zeros(4,2)]; [1, 0]; [0, 1]; [0, 0]]);
+
+% first stage
+Mt   = V.' * M * P;
+Kt_P = V.' * K * P;
+Kt_Q = V.' * K * Q;
+ft   = V.' * f;
+Kb_P = W.' * K * P;
+Kb_Q = W.' * K * Q;
+fb   = W.' * f;
+
+% second basis functions
+Qb = Wb = null(Kb_Q);

### 2023-11-30 code/implicit_euler.m

@@ -1,0 +1,38 @@
+%   [x, t] = implicit_euler (T_0, T, Deltat, M, K, f, x_0)
+%
+% Computes the numerical solution of an initial value problem of the form
+%      M x' + K(x) x = f(t) on [T_0,T]
+%               x(0) = x_0
+% using the implicit Euler method.
+%
+% INPUT:
+%       - T_0, T, Deltat: time interval [T_0,T] and timestep
+%       - M: constant matrix
+%       - K: @(x) nonlinear matrix
+%       - f: @(t) time dependent rhs
+%       - x_0: initial condition
+%
+% OUTPUT:
+%       - x: [N_x, N_t] solution vector at the discrete points in time
+%       - t: [1, N_t] vector containing the discrete points in time
+
+function [x, t] = implicit_euler (T_0, T, Deltat, M, K, f, x_0)
+    N_x = length(x_0);
+    N_t = round((T - T_0) / Deltat);
+    x   = NaN(N_x, N_t+1);
+    x(:,1) = x_0;
+
+    t = T_0:Deltat:(T_0 + N_t*Deltat);
+    for n=1:N_t
+        if (~mod(n, 1000))
+            fprintf('\niteration no.: %i/%i\n', n, N_t);
+        end
+
+        % M (x_npo - x_n) + Deltat K(x_npo) x_npo = Deltat f(t_npo)
+        g = @(x_npo) M*(x_npo - x(:,n)) + Deltat*K(x_npo)*x_npo - Deltat*f(t(n+1));
+
+        options.TolX   = 1e-16;
+        options.TolFun = 1e-16;
+        x(:,n+1)       = fsolve(g, x(:,n), options);
+    end
+end

### 2023-11-30 code/tc_sc.m

@@ -1,0 +1,137 @@
+pkg load symbolic
+
+syms v_1 v_2 v_3 v_4 v_5 v_6 v_7 v_8 v_9 v_10 v_11 v_12 v_13 v_14 v_15 v_16
+syms i_L_1 i_L_2 i_L_3 i_L_4 i_L_5 i_L_6 i_L_7
+syms i_V_1 i_V_2
+
+syms R_1 R_2 R_3 R_4 R_5 R_6 R_7
+syms L_1 L_2 L_12 L_3 L_4 L_34 L_5 L_6 L_56 L_7
+syms C_1 C_2 C_3 C_4 C_5 C_6 C_7 C_8 C_9
+syms v_s_1 v_s_2
+
+n_v   = 16;
+n_R   = 7;
+n_L   = 7;
+n_C   = 9;
+n_V   = 2;
+
+A_R       = zeros(n_v,n_R);
+A_R(1,1)  = 1;
+A_R(2,1)  = -1;
+A_R(8,2)  = 1;
+A_R(9,2)  = -1;
+A_R(3,3)  = 1;
+A_R(4,3)  = -1;
+A_R(10,4) = 1;
+A_R(11,4) = -1;
+A_R(5,5)  = 1;
+A_R(6,5)  = -1;
+A_R(12,6) = 1;
+A_R(13,6) = -1;
+A_R(12,7) = 1;
+A_R(15,7) = -1;
+
+A_L       = zeros(n_v,n_L);
+A_L(2,1)  = 1;
+A_L(3,1)  = -1;
+A_L(9,2)  = 1;
+A_L(10,2) = -1;
+A_L(4,3)  = 1;
+A_L(5,3)  = -1;
+A_L(11,4) = 1;
+A_L(12,4) = -1;
+A_L(6,5)  = 1;
+A_L(7,5)  = -1;
+A_L(13,6) = 1;
+A_L(14,6) = -1;
+A_L(15,7) = 1;
+A_L(16,7) = -1;
+
+A_C       = zeros(n_v,n_C);
+A_C(3,1)  = 1;
+A_C(10,2) = 1;
+A_C(3,2)  = -1;
+A_C(5,3)  = 1;
+A_C(12,4) = 1;
+A_C(5,4)  = -1;
+A_C(7,5)  = 1;
+A_C(14,6) = 1;
+A_C(7,6)  = -1;
+A_C(10,7) = -1;
+A_C(12,8) = -1;
[…]

### 2023-11-30 code/test_circuit.m

@@ -1,0 +1,176 @@
+T_0    = 0;
+T      = 3;
+Deltat = 0.1;
+
+G_1 = G_2 = G_3 = G_4 = G_5 = G_6 = G_7 = 1;
+L_1 = L_2 = L_3 = L_4 = L_5 = L_6 = L_7 = 1;
+L_12 = L_34 = L_56 = 0.1;
+C_1 = C_2 = C_3 = C_4 = C_5 = C_6 = C_7 = C_8 = C_9 = 1;
+
+v_s_1 = @(t) sin(4*t) + cos(2*t);
+v_s_2 = @(t) sin(3*t) + cos(t);
+
+n_v   = 16;
+n_R   = 7;
+n_L   = 7;
+n_C   = 9;
+n_V   = 2;
+
+A_R       = zeros(n_v,n_R);
+A_R(1,1)  = 1;
+A_R(2,1)  = -1;
+A_R(8,2)  = 1;
+A_R(9,2)  = -1;
+A_R(3,3)  = 1;
+A_R(4,3)  = -1;
+A_R(10,4) = 1;
+A_R(11,4) = -1;
+A_R(5,5)  = 1;
+A_R(6,5)  = -1;
+A_R(12,6) = 1;
+A_R(13,6) = -1;
+A_R(12,7) = 1;
+A_R(15,7) = -1;
+
+A_L       = zeros(n_v,n_L);
+A_L(2,1)  = 1;
+A_L(3,1)  = -1;
+A_L(9,2)  = 1;
+A_L(10,2) = -1;
+A_L(4,3)  = 1;
+A_L(5,3)  = -1;
+A_L(11,4) = 1;
+A_L(12,4) = -1;
+A_L(6,5)  = 1;
+A_L(7,5)  = -1;
+A_L(13,6) = 1;
+A_L(14,6) = -1;
+A_L(15,7) = 1;
+A_L(16,7) = -1;
+
+A_C       = zeros(n_v,n_C);
+A_C(3,1)  = 1;
+A_C(10,2) = 1;
+A_C(3,2)  = -1;
+A_C(5,3)  = 1;
+A_C(12,4) = 1;
+A_C(5,4)  = -1;
+A_C(7,5)  = 1;
+A_C(14,6) = 1;
+A_C(7,6)  = -1;
+A_C(10,7) = -1;
+A_C(12,8) = -1;
+A_C(16,9) = -1;
+
[…]

### 2023-11-30 code/trapezoidal_rule_MNA.m

@@ -1,0 +1,38 @@
+%   [x, t] = trapezoidal_rule_MNA (T_0, T, Deltat, M, K, f, x_0)
+%
+% Computes the numerical solution of an initial value problem of the form
+%       M x' + K(x) x + f(t) = 0 on [T_0,T]
+%     x(0) = x_0
+% using the trapezoidal rule.
+%
+% INPUT:
+%       - T_0, T, Deltat: time interval [T_0,T] and timestep
+%       - M: constant matrix
+%       - K: @(x) nonlinear matrix
+%       - f: @(t) time dependent source term
+%       - x_0: initial condition
+%
+% OUTPUT:
+%       - x: [N_x, N_t] solution vector at the discrete points in time
+%       - t: [1, N_t] vector containing the discrete points in time
+
+function [x, t] = trapezoidal_rule_MNA (T_0, T, Deltat, M, K, f, x_0)
+    N_x    = length(x_0);
+    N_t    = round((T - T_0) / Deltat);
+    x      = NaN(N_x, N_t+1);
+    x(:,1) = x_0;
+
+    t = T_0:Deltat:(T_0 + N_t*Deltat);
+    for n = 1:N_t
+        if (~mod(n, 1000))
+            fprintf('\niteration no.: %i/%i\n', n, N_t);
+        end
+
+        % M (x_npo - x_n) + Deltat/2 ( K(x_n) x_n - K(x_npo) x_npo + f(t_n) - f(t_npo) ) = 0
+        g = @(x_npo) M*(x_npo - x(:,n)) + Deltat/2 * ( K(x(:,n))*x(:,n) - K(x_npo)*x_npo + f(t(n)) - f(t(n+1)));
+
+        options.TolX   = 1e-16;
+        options.TolFun = 1e-16;
+        x(:,n+1)       = fsolve(g, x(:,n), options);
+    end
+end

### 2023-11-30 diode_oscillator_doe.m

@@ -1,157 +1,0 @@
-problem  = 1; % 1, 2
-variable = 'phi_3'; % phi_3, i_L, phi_2, phi_1
-tol      = [1e-2 5e-3 1e-3]; % 1e-2, 5e-3, 1e-3
-
-N_t = 1000;
-T_0 = 0;
-T   = 1e-2;
-
-% all available data
-L = [1e-3:0.1e-3:3e-3];
-C = [100e-9:10e-9:300e-9];
-t = linspace(T_0, T, N_t);
-
-% initial points
-t_i = [t(1); t(end); t(1); t(end); t(1); t(end); t(1); t(end)];
-L_i = [L(1); L(1); L(end); L(end); L(1); L(1); L(end); L(end)];
-C_i = [C(1); C(1); C(1); C(1); C(end); C(end); C(end); C(end)];
-
-if (problem == 1)
-    filename = ['do1_doe_N_t=' num2str(N_t) '.dat'];
-elseif (problem == 2)
-    filename = ['do2_doe_N_t=' num2str(N_t) '.dat'];
-end
-data = dlmread(filename);
-data = data(2:end,:);
-
-t_d = data(:,1);
-L_d = data(:,2);
-C_d = data(:,3);
-X   = [t_d L_d C_d];
-
-phi_3_d = data(:,4);
-
-R   = 500;
-i_s = @(t) 1e-4*sin(2*pi*200*t);
-if (problem == 1)
-    i_L_d   = data(:,5);
-    phi_2_d = data(:,6);
-elseif (problem == 2)
-    phi_2_d = data(:,5);
-    phi_1_d = phi_2_d + R*i_s(t_d);
-end
-
-if (strcmp(variable, 'phi_3'))
-    y = phi_3_d;
-elseif (strcmp(variable, 'i_L'))
-    y = i_L_d;
-elseif (strcmp(variable, 'phi_2'))
-    y = phi_2_d;
-elseif (strcmp(variable, 'phi_1'))
-    y = phi_1_d;
-end
-
-i_i = NaN(length(t_i), 1);
-for i=1:length(i_i)
-    i_i(i) = find((t_i(i) == t_d) & (L_i(i) == L_d) & (C_i(i) == C_d));
-end
-
-f = stk_model('stk_gausscov_aniso', 3);
-
-f.lognoisevariance = NaN;
-% f.lognoisevariance = -Inf;
-tic;
[…]

### 2023-11-30 diode_oscillator_doe_av.m

@@ -1,107 +1,0 @@
-problem  = 2; % 2
-variable = 'i_L'; % i_L
-tol      = [1e-3]; % 1e-3
-
-N_t = 1000;
-T_0 = 0;
-T   = 1e-2;
-
-% all available data
-t = linspace(T_0, T, N_t);
-
-% initial points (need more than boundary)
-t_i = [t(1); t(500); t(end)];
-
-filename = ['do2_doe_N_t=' num2str(N_t) '.dat'];
-data     = dlmread(filename);
-data     = data(2:end,:);
-
-t_d = data(1:N_t,1);
-X   = [t_d];
-
-R   = 500;
-i_s = @(t) 1e-4*sin(2*pi*200*t);
-if (strcmp(variable, 'i_L'))
-    y = i_s(t_d);
-end
-
-i_i = NaN(length(t_i), 1);
-for i=1:length(i_i)
-    i_i(i) = find((t_i(i) == t_d));
-end
-
-f = stk_model('stk_gausscov_aniso', 1);
-
-f.lognoisevariance = NaN;
-% f.lognoisevariance = -Inf;
-tic;
-[theta, sigma] = stk_param_init(f, X(i_i,:), y(i_i));
-fprintf('\nstk_param_init: %d min\n', toc/60);
-
-tic;
-[f.param, f.lognoisevariance] = stk_param_estim(f, X(i_i,:), y(i_i), theta, sigma);
-% [f.param] = stk_param_estim(f, X, y, theta);
-fprintf('\nstk_param_estim: %d min\n', toc/60);
-
-tic;
-y_p = stk_predict(f, X(i_i,:), y(i_i), X);
-fprintf('\nstk_predict: %d min\n', toc/60);
-
-i_t = i_i;
-e_p = 1;
-for i_tol=1:length(tol)
-    fprintf('\ntol: %g\n', tol(i_tol));
-    if (problem == 2)
-        filename = ['do2_doe_i_t_' variable '_tol=' num2str(tol(i_tol)) '_N_t=' num2str(N_t) '.dat'];
-        if (exist(filename))
-            i_t = dlmread(filename);
-
-            filename = ['do2_doe_param_' variable '_tol=' num2str(tol(i_tol)) '_N_t=' num2str(N_t) '.dat'];
-            f.param = dlmread(filename);
-
-            filename = ['do2_doe_lognoisevariance_' variable '_tol=' num2str(tol(i_tol)) '_N_t=' num2str(N_t) '.dat'];
[…]

### 2023-11-30 diode_oscillator_results.m

@@ -1,236 +1,0 @@
-problem  = 2; % 1, 2
-
-plot_input       = 0; % 0, 1
-plot_convergence = 0; % 0, 1
-plot_solution    = 0; % 0, 1
-plot_consistency = 1; % 0, 1
-
-% only relevant for input/convergence
-variable = 'phi_3'; % phi_3, i_L, phi_2
-
-% only relevant for solution
-tol = 1e-3; % 1e-2, 5e-3, 1e-3
-
-if (plot_input == 1)
-    TOL = [1e-2, 5e-3, 1e-3];
-    N_t = 1000;
-
-    for tol=TOL
-        if (problem == 1)
-            filename = ['do1_doe_i_t_' variable '_tol=' num2str(tol) '_N_t=' num2str(N_t) '.dat'];
-            i_t      = dlmread(filename);
-
-            filename = ['do1_doe_N_t=' num2str(N_t) '.dat'];
-        elseif (problem == 2)
-            filename = ['do2_doe_i_t_' variable '_tol=' num2str(tol) '_N_t=' num2str(N_t) '.dat'];
-            i_t      = dlmread(filename);
-
-            filename = ['do2_doe_N_t=' num2str(N_t) '.dat'];
-        end
-
-        data = dlmread(filename);
-        data = data(2:end,:);
-
-        t = data(:,1);
-        L = data(:,2);
-        C = data(:,3);
-        X = [t L C];
-
-        if (problem == 1)
-            filename = ['do1_input_' variable '_tol=' num2str(tol) '.dat'];
-        elseif (problem == 2)
-            filename = ['do2_input_' variable '_tol=' num2str(tol) '.dat'];
-        end
-
-        fid = fopen(filename, 'w');
-        fprintf(fid, 'x  y\n');
-        dlmwrite(fid, X(i_t,2:3), 'delimiter', '  ', 'append', 'on');
-        fclose(fid);
-    end
-end
-
-if (plot_convergence == 1)
-    tol = [1e-2, 5e-3, 1e-3];
-    N_t = 1000;
-
-    N_s = NaN(length(tol), 1);
-    for it=1:length(tol)
-       if (problem == 1)
-            filename = ['do1_doe_i_t_' variable '_tol=' num2str(tol(it)) '_N_t=' num2str(N_t) '.dat'];
[…]

### 2023-11-30 do1_sc.m

@@ -1,23 +1,0 @@
-% syms, sym, simplify, jacobian, inv, null, rank, +, -, *, /, ^, .'
-pkg load symbolic
-
-syms phi_1 phi_2 phi_3 i_L i_V G C L g_D v_s
-
-M = [[0, 0, 0, 0, 0]; [0, 0, 0, 0, 0]; [0, 0, C, 0, 0]; [0, 0, 0, L, 0]; [0, 0, 0, 0, 0]];
-
-K = [[G, -G, 0, 0, 1]; [-G, G, 0, 1, 0]; [0, 0, g_D, -1, 0]; [0, -1, 1, 0, 0]; [-1, 0, 0, 0, 0]];
-
-f = [[0]; [0]; [0]; [0]; [-v_s]];
-
-% first basis functions
-Q = W = sym([[1, 0, 0]; [0, 1, 0]; [0, 0, 0]; [0, 0, 0]; [0, 0, 1]]);
-P = V = sym([[0, 0]; [0, 0]; [1, 0]; [0, 1]; [0, 0]]);
-
-% first stage
-Mt   = V.' * M * P;
-Kt_P = V.' * K * P;
-Kt_Q = V.' * K * Q;
-ft   = V.' * f;
-Kb_P = W.' * K * P;
-Kb_Q = W.' * K * Q;
-fb   = W.' * f;

### 2023-11-30 do2_sc.m

@@ -1,49 +1,0 @@
-% syms, sym, simplify, jacobian, inv, null, rank, +, -, *, /, ^, .'
-pkg load symbolic
-
-syms phi_1 phi_2 phi_3 i_L G C L g_D i_s i_sp
-
-M = [[0, 0, 0, 0]; [0, 0, 0, 0]; [0, 0, C, 0]; [0, 0, 0, L]];
-
-K = [[G, -G, 0, 0]; [-G, G, 0, 1]; [0, 0, g_D, -1]; [0, -1, 1, 0]];
-
-f = [[i_s]; [0]; [0]; [0]];
-
-% first basis functions
-Q = W = sym([[1, 0]; [0, 1]; [0, 0]; [0, 0]]);
-P = V = sym([[0, 0]; [0, 0]; [1, 0]; [0, 1]]);
-
-% first stage
-Mt   = V.' * M * P;
-Kt_P = V.' * K * P;
-Kt_Q = V.' * K * Q;
-ft   = V.' * f;
-Kb_P = W.' * K * P;
-Kb_Q = W.' * K * Q;
-fb   = W.' * f;
-
-% second basis functions
-Qb = Wb = sym([[1]; [1]])
-
-% Wb.' * Kb_P
-Qt  = sym([[1]; [0]])
-Qtp = sym([[0]; [0]]);
-Pt  = sym([[0]; [1]])
-Ptp = sym([[0]; [0]]);
-
-% xt_P = inv(Wb.' * Kb_P * Pt) * Wb.' * fb;
-xt_P  = i_s;
-xt_Pp = i_sp;
-
-% (Mt * Qt).';
-Wt = sym([[0]; [1]])
-Vt = sym([[1]; [0]])
-
-% second stage
-Mt_2  = Vt.' * Mt * Qt
-Kt_P2 = Vt.' * (Mt * Qtp + Kt_P * Qt)
-Kt_Q2 = Vt.' * Kt_Q * Qb
-ft_2  = Vt.' * (ft - Mt * Pt * xt_Pp - (Mt * Ptp + Kt_P * Pt) * xt_P)
-Kb_P2 = Wt.' * (Mt * Qtp + Kt_P * Qt)
-Kb_Q2 = Wt.' * Kt_Q * Qb
-fb_2  = Wt.' * (ft - Mt * Pt * xt_Pp - (Mt * Ptp + Kt_P * Pt) * xt_P)

### 2023-11-30 implicit_euler.m

@@ -1,35 +1,0 @@
-%   [x, t] = implicit_euler (T_0, T, Deltat, M, K, f, x_0)
-%
-% Computes the numerical solution of an initial value problem of the form
-%      M x' + K(x) x = f(t) on [T_0,T]
-%               x(0) = x_0
-% using the implicit Euler method.
-%
-% INPUT:
-%       - T_0, T, Deltat: time interval [T_0,T] and timestep
-%       - M: constant matrix
-%       - K: @(x) nonlinear matrix
-%       - f: @(t) time dependent rhs
-%       - x_0: initial condition
-%
-% OUTPUT:
-%       - x: [N_x, N_t] solution vector at the discrete points in time
-%       - t: [1, N_t] vector containing the discrete points in time
-
-function [x, t] = implicit_euler (T_0, T, Deltat, M, K, f, x_0)
-    N_x = length(x_0);
-    N_t = round((T - T_0) / Deltat);
-    x   = NaN(N_x, N_t+1);
-    x(:,1) = x_0;
-
-    t = T_0:Deltat:(T_0 + N_t*Deltat);
-    for n=1:N_t
-        if (~mod(n, 1000))
-            fprintf('\niteration no.: %i/%i\n', n, N_t);
-        end
-
-        % M (x_npo - x_n) + Deltat K(x_npo) x_npo = Deltat f(t_npo)
-        g = @(x_npo) M*(x_npo - x(:,n)) + Deltat*K(x_npo)*x_npo - Deltat*f(t(n+1));
-        x(:,n+1) = fsolve(g, x(:,n));
-    end
-end

### 2023-12-03 code/fwr2_sc.m
article

@@ -1,0 +1,68 @@
+% syms, sym, simplify, jacobian, inv, null, rank, +, -, *, /, ^, .'
+pkg load symbolic
+
+syms phi_1 phi_2 phi_3 phi_4 i_L_1 i_L_2
+syms L_1 L_2 L_12 g_D_1 g_D_2 g_D_3 g_D_4 R i_s i_sp
+
+phi = [phi_1; phi_2; phi_3; phi_4];
+i_L = [i_L_1; i_L_2];
+i_V = [];
+
+n_phi = length(phi);
+n_L   = length(i_L);
+n_V   = length(i_V);
+
+A_R = [[0, 0, 0, 0, 0]; [1, 0, -1, 0, 0]; [-1, -1, 0, 0, 1]; [0, 0, 1, 1, -1]];
+A_L = [[1, 0]; [0, -1]; [0, 0]; [0, 0]];
+A_C = [0; 0; 1; -1];
+A_V = [];
+A_I = [-1; 0; 0; 0];
+
+G = [[g_D_1, 0, 0, 0, 0]; [0, g_D_2, 0, 0, 0]; [0, 0, g_D_3, 0, 0]; [0, 0, 0, g_D_4, 0]; [0, 0, 0, 0, 1/R]];
+L = [[L_1, L_12]; [L_12, L_2]];
+
+M = [[zeros(n_phi, n_phi), zeros(n_phi, n_L), zeros(n_phi, n_V)]; [zeros(n_L, n_phi), L, zeros(n_L, n_V)]; [zeros(n_V, n_phi), zeros(n_V, n_L), zeros(n_V, n_V)]];
+% K = [[A_R*G*A_R.', A_L, A_V]; [-A_L.', zeros(n_L, n_L), zeros(n_L, n_V)]; [-A_V.', zeros(n_V, n_L), zeros(n_V, n_V)]];
+K = [[A_R*G*A_R.', A_L, A_V]; [-A_L.', zeros(n_L, n_L), zeros(n_L, n_V)]];
+f = [A_I*i_s; zeros(n_L, 1); zeros(n_V, 1)];
+
+% first basis functions
+Q = W = null(M);
+P = V = sym([[zeros(4, 2)]; [1, 0]; [0, 1]]);
+
+% first stage
+Mt   = V.' * M * P;
+Kt_P = V.' * K * P;
[…]

### 2024-01-22 doc/article/article.tex
article
revision

@@
 
+\usepackage{amsthm}
 \usepackage{mathtools}
@@
 \usepackage{siunitx}
-\usepackage{subcaption}
+\usepackage[european]{circuitikz}
 
@@
 
+% \newtheorem{assumption}{Assumption}
+% \newtheorem{theorem}{Theorem}
+% \newtheorem{remark}{Remark}
+
+\newcommand{\assumptionautorefname}{Assumption}
+\renewcommand{\theoremautorefname}{Theorem}
+\newcommand{\remarkautorefname}{Remark}
+
+\usetikzlibrary{external}
+\tikzexternalize
+
 \begin{document}
@@
 
-\abstract[Summary]{Electrical circuits are present in a variety of technologies, making their design an important part of computer aided engineering.
-The growing number of tunable parameters that affect the final design leads to a need for new approaches of quantifying their impact.
+\abstract[Summary]{
+    Electrical circuits are present in a variety of technologies, making their design an important part of computer aided engineering.
+The growing number of parameters that affect the final design leads to a need for new approaches to quantify their impact.
 Machine learning may play a key role in this regard, however current approaches often make suboptimal use of existing knowledge about the system at hand.
@@
 hidden constraints that the solution needs to fulfill.
-We aim to use the recently introduced dissection concept for DAEs that can decouple a given system into ordinary differential equations, only depending on differential variables, and purely algebraic equations that describe the relations between differential and algebraic variables.
-The idea then is to only learn the differential variables and reconstruct the algebraic ones using the relations from the decoupling.
-This approach guarantees that the algebraic constraints are fulfilled up to the accuracy of the nonlinear system solver, which represents the main benefit highlighted in this article.}
-
-\keywords{electrical circuits, differential-algebraic equations, modified nodal analysis, dissection concept, machine learning}
+We use the recently introduced dissection index that can decouple a given system of DAEs into ordinary differential equations, only depending on differential variables, and purely algebraic equations, that describe the relations between differential and algebraic variables.
+The idea is to then only learn the differential variables and reconstruct the algebraic ones using the relations from the decoupling.
+This approach guarantees that the algebraic constraints are fulfilled up to the accuracy of the nonlinear system solver, and it may also reduce the learning effort as only the differential variables need to be learned.}
+
+\keywords{electrical circuits, differential-algebraic equations, modified nodal analysis, dissection index, machine learning}
 
@@
 \label{sec:int}
-Design optimization and uncertainty quantification are ubiquitous tools of modern computer aided  engineering.
-Due to the ever-increasing complexity of engineering systems, machine learning techniques have gained popularity for constructing surrogate models when the objective function becomes expensive to evaluate and a large number of (design or uncertainty) parameters are present.
-In these cases, classical model order reduction techniques\cite{schilders2008} or function approximation approaches\cite{xiu2010} suffer from the curse of dimensionality: the number of operations to construct and the memory required to store the model grow exponentially with respect to the number of parameters.
-Experimental and in some cases even theoretical evidence\cite{beneventano2020} shows that machine learning approaches may be able to overcome the curse of dimensionality and provide surrogates that are fast to evaluate, while requiring comparably little data.
-
-In the particular case of electrical circuit design, neural networks have been used for design optimization for over 20 years\cite{zaabab1995,fang2000}.
+Design optimization and uncertainty quantification are key tools of modern computer aided engineering, that both rely on objective functions to express quantities of interest in terms of the variables of the underlying system.
+Due to the increasing complexity of engineering systems, machine learning approaches have gained popularity for constructing surrogate models of objective functions when they become expensive to evaluate and a large number of (design or uncertainty) parameters are present.
+In such situations, classical model order reduction techniques\cite{schilders2008} or function approximation approaches\cite{xiu2010} suffer from the curse of dimensionality: the number of operations to construct and the memory required to store the surrogate model grow exponentially with respect to the number of parameters.
+Experimental and in some cases even theoretical evidence\cite{beneventano2020} shows that machine learning approaches may be able to overcome this curse of dimensionality and provide surrogate models that are fast to evaluate, while requiring comparably little data for their construction and storage.
+
+In the context of electrical circuit design, neural networks have been used for design optimization for over 20 years\cite{zaabab1995,fang2000}.
 More recently, Gaussian process regression has been employed for both uncertainty quantification and design optimization during analog integrated circuit design\cite{wang2018,sanabria2020}.
-The commonality between these approaches is that they focus on the learning part: they aim to provide a computationally efficient and accurate surrogate, given data produced by some circuit simulation software.
+The commonality between these approaches is that they focus on the learning part: they all aim to provide a computationally efficient and accurate surrogate model, given data produced by some circuit simulator.
 Thus, they all treat the circuit simulator as a black box that simply provides the data which is then used for constructing the surrogate.
@@
 Applying MNA to a given circuit leads to systems of differential-algebraic equations (DAEs), which can generally be written as systems of implicit differential equations \cite{denk1991}
-\begin{align*}
-    \mb{f}(\mb{x}', \mb{x}, t, \mb{p}) = \mb{0}, \quad \mb{x}(0) = \mb{x}_0,
-\end{align*}
-where the Jacobian $\mb{J}_{\mb{x}'}(\mb{f})$, of $\mb{f}$ w.r.t.~$\mb{x}'$, is singular and $\mb{p}$ are the design or uncertainty parameters.
-To give a more intuitive description, one can think of DAEs as ordinary differential equations (ODEs) that are constrained to a manifold determined by (hidden) constraints on the solution variables.
-Multiple so-called index concepts were introduced in order to characterize the difficulties that appear in the context of DAEs.
-A reference by Mehrmann\cite{mehrmann2015} outlines some of the most prominent, including the differentiation, tractability, strangeness and perturbation indices.
-For all of these, the index $\nu \in \mathbb{N}$ depends on structural properties of the DAE and (informally) measures the degree of difficulty associated with solving it.
-In this sense, higher index DAEs usually lead to more problems, both analytically and numerically.
-
-Various relationships between the indices, including situations when they coincide, are also known\cite{mehrmann2015} and the structure and index of MNA are well understood for both independent sources\cite{tischendorf1999} and controlled sources\cite{estevez2000ijcta}.
-Rather recently, the dissection concept was introduced\cite{jansen2014}.
-It combines features of the strangeness and tractability indices to decouple a DAE into an ODE and a set of purely algebraic equations.
-In particular, the entire dynamics of the solution may then be found only using the ODE.
-We aim to use this dissection concept to propose an approach for learning electrical circuits more accurately and efficiently.
-
-In the following, \autoref{sec:mna} introduces MNA in more detail and states some well-known related results.
-Afterwards, \autoref{sec:dc} outlines the dissection concept and showcases its properties using small example circuits.
-The same examples are then used to present the new approach on an abstract level in \autoref{sec:ial} and on a numerical level in \autoref{sec:ne}, where we also showcase an alternative method that trades some of the accuracy of our approach for an easier implementation.
-Preliminary conclusions about the effectiveness of our approach and future research directions are given in \autoref{sec:cfr}.
-
-\section{Modified nodal analysis}
-\label{sec:mna}
-We begin by looking at the system of DAEs that results from MNA.
-From here onward we will disregard controlled sources.
-As these are of crucial importance for engineering applications however, we will state whether extensions including controlled sources are available or missing at the appropriate times.
-Borrowing the notation from Tischendorf\cite{tischendorf1999}, and only explicitly expressing the time dependence of independent sources for brevity, the system of MNA reads
+\begin{align}
+    \mb{F}(\mb{x}', \mb{x}, t, \mb{p}) = \mb{0}, \quad \mb{x}(0) = \mb{x}_0, \label{eq:int_dae}
+\end{align}
+where the Jacobian $\mb{J}_{\mb{x}'}(\mb{F})$, of $\mb{F}$ w.r.t.~$\mb{x}'$, is singular and $\mb{p}$ are the design or uncertainty parameters.
+Intuitively, one can think of DAEs as ordinary differential equations (ODEs) that are constrained to a manifold determined by (hidden) constraints on the solution variables $\mb{x}$.
+We aim to exploit the special structure of the DAEs arising from MNA by using the dissection index\cite{jansen2014} to propose an approach for learning electrical circuits more accurately and efficiently.
+More concretely, we use the dissection index to decouple the DAEs into sets of ODEs and purely algebraic equations, such that the entire dynamics of the solution may then be found only using the ODEs, while the algebraic equations may be used to recover the entire solution.
+
+In the following, \autoref{sec:md} introduces MNA and DAEs in more detail and states some well-known results.
+Afterwards, \autoref{sec:di} outlines the dissection index and showcases its properties using example circuits.
+The new approach is then presented on an abstract level in \autoref{sec:ial}, and on a numerical level in \autoref{sec:ne}.
+Preliminary conclusions about the effectiveness of the approach and future research directions are given in \autoref{sec:cfr}.
+
+\section{MNA and DAEs}
+\label{sec:md}
+We first look at the system of DAEs that results from MNA when not considering controlled sources.
+As they are of crucial importance for engineering applications however, we will note whether extensions including controlled sources are available or missing at the appropriate times.
+Borrowing the notation from Tischendorf\cite{tischendorf1999}, the system of MNA reads
 \begin{subequations}
-    \label{eq:mnai}
+    \label{eq:md_mnai}
     \begin{align}
@@
 \end{subequations}
-where $\gR$, $\qC$ and $\phiL$ represent functions modeling resistive, capacitive or inductive devices respectively.
+where $\gR$, $\phiL$ and $\qC$ represent functions modeling resistive, capacitive or inductive devices respectively.
 The terms for independent current and voltage sources are given by $\i{s}(t)$ and $\v{s}(t)$, while $\vphi$ denotes the vector of nodal potentials and $\i{L} $, $\i{V}$ are the currents flowing through branches containing inductors or voltage sources respectively.
 The last ingredient is given by the incidence matrices $\A{*}$, where $*$ indicates the device type.
-These collect the branch to node relations of the underlying electrical network when considering the branches and nodes as edges and vertices of a graph.
-In the context of design optimization or uncertainty quantification, the device functions $\gR$, $\qC$ or $\phiL$ could e.g.~depend on (subsets of) the parameters $\mb{p}$.
-
-In order to obtain a version of \eqref{eq:mnai} that is better suited to analysis and implementation, we consider the device function Jacobians
-\begin{align}
-    \mb{L}(\i{L}) \coloneqq \mb{J}_{\i{L}}(\phiL), \quad \mb{C} \big( \AT{C} \vphi \big) \coloneqq \mb{J}_{\AT{C} \vphi}(\qC) \label{eq:dfj},
-\end{align}
-where we use the same notation for the Jacobians as in \autoref{sec:int}. We also assume that the device function for resistors may be expressed as
-\begin{align}
-    \gR \big( \AT{R} \vphi \big) = \mb{G} \big( \AT{R} \vphi \big) \AT{R} \vphi \label{eq:dfr}.
-\end{align}
-In practice, this is accomplished by either considering the chordal (static) conductance $\frac{[\gR]_i}{[\AT{R} \vphi]_i}$ for each resistor $\mr{R}_i$ separately or by using the Jacobian $\mathbf{J}_{ \AT{R} \vphi}(\gR)$.
-Inserting \eqref{eq:dfj} and \eqref{eq:dfr} into the original system and writing it in matrix form yields
+These collect the branch to node relations of the underlying electrical network, when considering the branches and nodes as edges and vertices of a directed graph.
+In the context of design optimization or uncertainty quantification, the device functions $\gR$, $\phiL$ or $\qC$ could e.g.~depend on (subsets of) the parameters $\mb{p}$.
+
+\textcolor{red}{In order to obtain a version of \eqref{eq:md_mnai} that is better suited to analysis and implementation, we consider the device function Jacobians
+\begin{align}
+    \mb{G} \big( \AT{R} \vphi \big) \coloneqq \mb{J}_{\AT{R} \vphi}(\gR), \quad \mb{L}(\i{L}) \coloneqq \mb{J}_{\i{L}}(\phiL), \quad \mb{C} \big( \AT{C} \vphi \big) \coloneqq \mb{J}_{\AT{C} \vphi}(\qC), \label{eq:md_dfj}
+\end{align}
+where we use the same notation for the Jacobians as in \autoref{sec:int}.} Inserting \eqref{eq:md_dfj} into the original system, while linearizing $\gR$ and writing everything in matrix form, yields
 \begin{align}
@@
         \v{s}(t)
-    \end{bmatrix} = \mb{0} \label{eq:mnaii},
-\end{align}
-where we assume that $\qC$ and $\phiL$ do not explicitly depend on time.
-MNA is usually presented in the form of \eqref{eq:mnaii} and for it, there exists the following well-known topological index result.
-\begin{quote}
-    Assuming the matrix-valued functions $\mathbf{G}$, $\mathbf{L}$, $\mathbf{C}$ from \eqref{eq:dfj} and \eqref{eq:dfr} positive definite:
+    \end{bmatrix} = \mb{0}, \label{eq:md_mnaii}
+\end{align}
+assuming that $\phiL$ and $\qC$ do not explicitly depend on time.
+
+\textcolor{red}{We note that \eqref{eq:md_mnaii} readily extends to multiport devices\cite{tischendorf1999}.
+A general inductive $n$-port for example, can also be modeled by a function $\phiL(\i{L}) = \big[ \phi_1(\i{L}), \dotsc, \phi_{n-1}(\i{L}) \big]^\T$, however the Jacobian is not necessarily diagonal in this case as the component functions of $\phiL$ can each depend on all the currents $\i{L} = [i_1, \cdots, i_{n-1}]^\T$.
+Regarding the incidence matrices, one chooses a reference node $\varphi_0$ for the multiport device and then considers $n-1$ branches from the remaining $n-1$ nodes to the reference.
+Interpreting the multiport as a single node and using Kirchhoff's current law then gives $i_0 = -\sum_{k=1}^{n-1} i_k$ for the reference current, when orienting all currents to point toward the device.
+Following this approach, there is then no difference in treating multiports compared to one-ports\cite{tischendorf1999}.}
+
+\subsection{Index concepts}
+\label{subsec:ic}
+\textcolor{red}{Before outlining the dissection index, we want to give a brief introduction to index concepts more generally.
+There is no unique definition of the index of a DAE, but rather there are multiple related concepts that each possess different strengths and weaknesses\cite{mehrmann2015}.
+One important aspect that unites these ideas is that they agree in key cases, e.g.~when looking at linear DAEs, and the same also holds true for the dissection index.
+To emphasize the practical importance of the notion of index, we take a closer look at the perturbation index.
+It is based on a perturbed version of the DAE \eqref{eq:int_dae} (we leave out the parameters $\mb{p}$ for conciseness)
+\begin{align}
+    \mb{F}(\mbh{x}', \mbh{x}, t) = \boldsymbol{\varepsilon}(\mbh{x}', \mbh{x}, t), \quad \mbh{x}(0) = \mbh{x}_0, \label{eq:ic_pdae}
+\end{align}
+where $\varepsilon$ is a sufficiently smooth perturbation, such that the required derivatives exist.
+\begin{definition}
+    \label{def:ic_pi}
+    Let $\mb{x}$ be a solution of the unperturbed DAE, then the DAE is said to have perturbation index $\nu \in \mathbb{N}$, if $\nu$ is the smallest natural number such that, for any sufficiently smooth solution $\mbh{x}$ of \eqref{eq:ic_pdae}, there exists $c \in \mathbb{R}$ with
+    \begin{align*}
+        \| \mbh{x} - \mb{x} \| \leq c \big( \| \mbh{x}_0 - \mb{x}_0 \| + \| \boldsymbol{\varepsilon} \|_\infty + \| \boldsymbol{\varepsilon}' \|_\infty + \dotsb + \| \boldsymbol{\varepsilon}^{(\nu - 1)} \|_\infty \big)
+    \end{align*}
+    for an appropriate norm $\| \cdot \|$ and the right hand side small enough.
+\end{definition}
+The idea behind this definition is to capture the impact of perturbations on the solution, as the name suggests.
+Usually these pertubations are assumed small, in the sense that $\| \boldsymbol{\varepsilon} \|_\infty \ll 1$, but fast changing, such that $\| \boldsymbol{\varepsilon}^{(\nu)} \|_\infty$ may grow very quickly in $\nu$.}
+
+\paragraph{Example}
+\begin{figure}[b]
+    \begin{center}
+        \includegraphics[width=0.35\textwidth]{vcr}
+    \end{center}
+    \caption{Small example circuit for illustrating the perturbation index.}
+    \label{fig:ic_vcr}
+\end{figure}
+\textcolor{red}{To illustrate the perturbation index, and also to hint at its relevance for circuit simulation, we consider the small example given in \autoref{fig:ic_vcr}.
+Applying MNA to the circuit and introducing a perturbation yields the following DAE
+\begin{subequations}
+    \label{eq:ic_vcr}
+    \begin{align}
+        C \ddt \hat{\varphi}_1 + \frac{1}{R} \hat{\varphi}_1 + \hat{i}_\mr{V} &= \varepsilon_1(\mbh{x}', \mbh{x}, t) \label{eq:ic_vcra}\\
+        -\hat{\varphi}_1 + v_\mr{s}(t) &= \varepsilon_2(\mbh{x}', \mbh{x}, t), \label{eq:ic_vcrb}
+    \end{align}
+\end{subequations}
+where $\mbh{x} = [\hat{\varphi}_1, \hat{i}_\mr{V}]^\T$. Using \eqref{eq:ic_vcrb} we find
+\begin{align}
+    \hat{\varphi}_1 = v_\mr{s}(t) - \varepsilon_2(\mbh{x}', \mbh{x}, t). \label{eq:ic_vcrphi}
+\end{align}
+Inserting \eqref{eq:ic_vcrphi} into \eqref{eq:ic_vcra} and rearranging then gives
+\begin{align*}
+    \hat{i}_\mr{V} &= \varepsilon_1(\mbh{x}', \mbh{x}, t) - C \ddt \big( v_\mr{s}(t) - \varepsilon_2(\mbh{x}', \mbh{x}, t) \big) - \frac{1}{R} \big( v_\mr{s}(t) - \varepsilon_2(\mbh{x}', \mbh{x}, t) \big).
+\end{align*}
+Noting that the solution $\mb{x}$ to the unperturbed problem follows directly from the perturbed solution by setting $\boldsymbol{\varepsilon} = \mb{0}$, we obtain
+\begin{align}
+    \| \mbh{x} - \mb{x} \| = \begin{Vmatrix}
+        -\varepsilon_2(\mbh{x}', \mbh{x}, t)\\
+        \varepsilon_1(\mbh{x}', \mbh{x}, t) + C \ddt \varepsilon_2(\mbh{x}', \mbh{x}, t) + \frac{1}{R} \varepsilon_2(\mbh{x}', \mbh{x}, t)
+    \end{Vmatrix}, \label{eq:ic_vcrpi}
+\end{align}
+so the unperturbed system corresponding to \eqref{eq:ic_vcr} has perturbation index $\nu = 2$ as \eqref{eq:ic_vcrpi} depends on the first derivative of $\varepsilon_2$.}
+
+\paragraph{Index of MNA}
+The structure and index of MNA are well understood when only considering independent sources\cite{tischendorf1999} (as in \eqref{eq:md_mnaii}), but also when including controlled sources\cite{estevez2000ijcta}.
+For the case without controlled sources, there exists the following well-known topological index result.
+\begin{theorem}
+    \label{th:ic_imna}
+    Assuming the matrix-valued functions $\mb{G} \big( \AT{R} \vphi \big)$, $\mb{L}(\i{L})$, $\mb{C} \big( \AT{C} \vphi \big)$ from \eqref{eq:md_dfj} positive definite:
     \begin{enumerate}
-        \item The index of \eqref{eq:mnaii} is $\nu \leq 2$.
-        \item The index of \eqref{eq:mnaii} is $\nu \leq 1$, if and only if there are no loops consisting only of capacitors and voltage sources containing at least one voltage source and no cutsets consisting only of inductors and current sources.
+        \item The index of \eqref{eq:md_mnaii} is $\nu \leq 2$.
+        \item The index of \eqref{eq:md_mnaii} is $\nu \leq 1$, if and only if there are no loops consisting only of capacitors and voltage sources containing at least one voltage source and no cutsets consisting only of inductors and current sources.
     \end{enumerate}
-    The result can be found in terms of the tractability index\cite{tischendorf1999}, the differentiation index \cite{estevez2000ijcta} and the dissection concept\cite{jansen2014}.
-% \cite[Theorem 4.28]{jansen2014}
-\end{quote}
-We remark that there also exist extensive results about when the index of MNA including controlled sources does not exceed $\nu = 2$\cite{estevez2000ijcta}.
-
-\section{Dissection concept}
-\label{sec:dc}
-This section serves as a general introduction to the dissection concept and more specifically its application to MNA as given in \eqref{eq:mnaii}.
-Consider a linear DAE in standard form\cite{jansen2014},
-\begin{align}
-    \mb{M} \ddt \mb{x} + \mb{K} \mb{x} + \mb{f}(t) = \mb{0}, \quad \ker \mb{M} \supsetneq \{ \mb{0} \} \label{eq:dae},
-\end{align}
-where we again keep the time dependence of $\mb{x}$ implicit.
-We now demonstrate the first two steps of the dissection concept when applied to systems of the form of \eqref{eq:dae} and remark that although \eqref{eq:mnaii} is a nonlinear DAE, the decoupling still takes the same form in this case due to the special structure of MNA\cite{jansen2014}.
+\end{theorem}
+The result can be found in terms of the tractability index\cite{tischendorf1999}, the differentiation index\cite{estevez2000ijcta} and the dissection index\cite{jansen2014}.
+We again note that there also exist extensive results about when the index of MNA including controlled sources does not exceed $\nu = 2$\cite{estevez2000ijcta}.
+
+\section{Dissection index}
+\label{sec:di}
+\textcolor{red}{We now focus on the dissection index.
+As mentioned in \autoref{sec:int}, the key property of this index concept is that it enables the decoupling of a DAE into an ODE and a set of purely algebraic equations.
+This is clearly different from the perturbation index, however other index concepts such as the tractability and differentiation indices also make such a decoupling possible.
+Still, the dissection index maintains some advantages over these concepts, as it provides a simple algorithmic procedure for the decoupling that is similar to the tractability index, and it poses less strict smoothness assumptions.
+In the case of MNA without controlled sources \eqref{eq:md_mnaii}, it is even possible to find a purely topological decoupling based on the dissection index\cite{jansen2014}.
+We will not make use of this topological decoupling here however, but rather consider the dissection index for a more general class of DAEs, to formulate the assumptions that are necessary for our method to work also for DAEs other than \eqref{eq:md_mnaii}.}
+
+\textcolor{red}{We consider a DAE in standard form\cite{jansen2014},
+\begin{align}
+    \mb{M}(\mb{x}) \ddt \mb{x} + \mb{K}(\mb{x}) \mb{x} + \mb{f}(t) = \mb{0}, \quad \ker \mb{M} \supsetneq \{ \mb{0} \}, \label{eq:di_dae}
+\end{align}
+where the matrix-valued functions $\mb{M}$ and $\mb{K}$ are derived from \eqref{eq:int_dae} by defining
+\begin{align*}
+    \mb{M}(\mb{x}) = \mb{J}_{\mb{x}'} \big( \mb{F}(\mb{x}', \mb{x}, t) \big), \quad \mb{K}(\mb{x}) = \mb{J}_{\mb{x}} \big( \mb{F}(\mb{x}', \mb{x}, t) \big).
+\end{align*}
+Note that the description of MNA in \eqref{eq:md_mnaii} is precisely of this form.
+In the following, we will demonstrate the first two steps of the dissection index when applied to systems of the form of \eqref{eq:di_dae}, while stating the assumptions of our approach.
+Appendix \autoref{sec:dim} contains additional remarks showing that these assumptions are fulfilled by \eqref{eq:md_mnaii}.}
 
 \subsection{Index one case}
-Assuming $\mb{M} \in \mathbb{R}^{n \times m}$, we define four basis functions $\mb{P}$, $\mb{Q}$, $\mb{V}$, $\mb{W}$ such that
-\begin{align*}
-    \mr{im\, } \mb{Q} = \ker \mb{M}, \quad \mr{im\, } \mb{W} = \ker \mb{M}^{\T}
-\end{align*}
-and the columns of $\mb{P}$ and $\mb{Q}$ together form a basis of $\mathbb{R}^n$, while the columns of $\mb{V}$ and $\mb{W}$ together form a basis of $\mathbb{R}^m$.
-
-\paragraph{Splitting the solution}
-The key idea is to use the basis functions to split the solution $\mb{x}$ into two parts
-\begin{align}
-    \mb{x} = \mb{P} \mbt{x} + \mb{Q} \mbb{x} \label{eq:vs},
+\label{subsec:ioc}
+Assuming $\mb{M}(\mb{x})$ to be sufficiently smooth with values in $\mathbb{R}^{n \times m}$, we define four basis functions $\mb{P}(\mb{x})$, $\mb{Q}(\mb{x})$, $\mb{V}(\mb{x})$, $\mb{W}(\mb{x})$ such that
+\begin{align*}
+    \mr{im\, } \mb{Q}(\mb{x}) = \ker \mb{M}(\mb{x}), \quad \mr{im\, } \mb{W}(\mb{x}) = \ker \mb{M}^\T (\mb{x})
+\end{align*}
+and the columns of $\mb{P}(\mb{x})$ and $\mb{Q}(\mb{x})$ together form a basis of $\mathbb{R}^n$, while the columns of $\mb{V}(\mb{x})$ and $\mb{W}(\mb{x})$ together form a basis of $\mathbb{R}^m$.
+We now make the following assumption.
+\textcolor{red}{\begin{assumption}
+    \label{as:ioc_bf}
+    The basis functions $\mb{P}$ and $\mb{Q}$ of $\mb{M}(\mb{x})$ are constant.
+\end{assumption}
+\autoref{as:ioc_bf} may seem restrictive, but it is fulfilled by many systems occurring in practical applications\cite{jansen2014}.
+This in particular includes MNA, as \autoref{re:dim_bf} shows.} The key idea of the dissection index is to use the basis functions to split the solution variables $\mb{x}$ into two parts
+\begin{align}
+    \mb{x} = \mb{P} \mbt{x} + \mb{Q} \mbb{x}, \label{eq:ioc_x}
 \end{align}
 where $\tilde{\cdot}$ is used to indicate differential (dynamic) variables and $\bar{\cdot}$ signifies algebraic (fixed) variables.
-When inserting the splitting \eqref{eq:vs} into the general system \eqref{eq:dae} we observe the following
-\begin{align}
-    \mb{M} \mb{P} \ddt \mbt{x} + \mb{K} \mb{P} \mbt{x} + \mb{K} \mb{Q} \mbb{x} + \mb{f}(t) = \mb{0} \label{eq:rs}
+When inserting the splitting \eqref{eq:ioc_x} into \eqref{eq:di_dae} we obtain
+\begin{align}
+    \mb{M}(\mb{x}) \mb{P} \ddt \mbt{x} + \mb{K}(\mb{x}) \mb{P} \mbt{x} + \mb{K}(\mb{x}) \mb{Q} \mbb{x} + \mb{f}(t) = \mb{0} \label{eq:ioc_dae}
 \end{align}
 and this motivates the notation, as only $\mbt{x}$ appears differentiated in time.
-
-\paragraph{Splitting the system}
-The procedure then continues by multiplying \eqref{eq:rs} with $\mb{V}^{\T}$ and $\mb{W}^{\T}$ from the left. This yields
+The procedure then continues by multiplying \eqref{eq:ioc_dae} once with $\mb{V}^\T (\mb{x})$ and $\mb{W}^\T (\mb{x})$ each from the left, to also split the system.
+This yields
 \begin{subequations}
-    \label{eq:io}
-    \begin{align}
-            \underbrace{\mb{V}^{\T} \mb{M} \mb{P}}_{\eqqcolon \mbt{M}} \ddt \mbt{x} + \underbrace{\mb{V}^{\T} \mb{K} \mb{P}}_{\eqqcolon \mbt{K}_\mb{P}} \mbt{x} + \underbrace{\mb{V}^{\T} \mb{K} \mb{Q}}_{\eqqcolon \mbt{K}_\mb{Q}} \mbb{x} + \underbrace{\mb{V}^{\T} \mb{f}(t)}_{\eqqcolon \mbt{f}(t)} &= \mb{0} \label{eq:ioa}\\
-            \underbrace{\mb{W}^{\T} \mb{K} \mb{P}}_{\eqqcolon \mbb{K}_\mb{P}} \mbt{x} + \underbrace{\mb{W}^{\T} \mb{K} \mb{Q}}_{\eqqcolon \mbb{K}_\mb{Q}} \mbb{x} + \underbrace{\mb{W}^{\T} \mb{f}(t)}_{\eqqcolon \mbb{f}(t)} &= \mb{0} \label{eq:iob},
+    \label{eq:ioc_io}
+    \begin{align}
+            \underbrace{\mb{V}^\T (\mb{x}) \mb{M}(\mb{x}) \mb{P}}_{\eqqcolon \mbt{M}(\mb{x})} \ddt \mbt{x} + \underbrace{\mb{V}^\T (\mb{x}) \mb{K}(\mb{x}) \mb{P}}_{\eqqcolon \mbt{K}_\mb{P}(\mb{x})} \mbt{x} + \underbrace{\mb{V}^\T (\mb{x}) \mb{K}(\mb{x}) \mb{Q}}_{\eqqcolon \mbt{K}_\mb{Q}(\mb{x})} \mbb{x} + \underbrace{\mb{V}^\T (\mb{x}) \mb{f}(t)}_{\eqqcolon \mbt{f}(t)} &= \mb{0} \label{eq:ioc_ioa}\\
+            \underbrace{\mb{W}^\T (\mb{x}) \mb{K}(\mb{x}) \mb{P}}_{\eqqcolon \mbb{K}_\mb{P}(\mb{x})} \mbt{x} + \underbrace{\mb{W}^\T (\mb{x}) \mb{K}(\mb{x}) \mb{Q}}_{\eqqcolon \mbb{K}_\mb{Q}(\mb{x})} \mbb{x} + \underbrace{\mb{W}^\T (\mb{x}) \mb{f}(t)}_{\eqqcolon \mbb{f}(t)} &= \mb{0}, \label{eq:ioc_iob}
     \end{align}
@@
 where we also introduce shorthands for the arising products.
-The key observations regarding \eqref{eq:io} are the following: $\mbt{M}$ is regular by construction, thus if $\mbb{K}_\mb{Q}$ is regular, \eqref{eq:iob} gives a purely algebraic equation for $\mbb{x}$ in terms of $\mbt{x}$.
-Solving this equation for $\mbb{x}$ and inserting the solution into \eqref{eq:ioa} then yields an ODE in the differential variables $\mbt{x}$.
-In this case, the \textbf{(dissection) index} of the DAE is defined to be $\nu = 1$.
+Using the fact that $\mbt{M}(\mb{x})$ is regular by construction\cite{jansen2014}, we can now define the index one case of the dissection index.
+\textcolor{red}{\begin{definition}
+    \label{def:ioc_io}
+    The DAE \eqref{eq:di_dae} has dissection index $\nu = 1$ if $\mbb{K}_\mb{Q}(\mb{x})$ is regular.
+\end{definition}}
+This is motivated by the observation that \eqref{eq:ioc_iob} is a purely algebraic equation with a locally unique solution for $\mbb{x}$ in terms of $\mbt{x}$, given that $\mbb{K}_\mb{Q}(\mb{x})$ is regular.
+In this case \eqref{eq:ioc_ioa} then describes an ODE in the differential variables $\mbt{x}$.
 
 \subsection{Index two case}
-There are of course many DAEs, including those described by MNA, which have a higher index.
-For these the dissection concept proceeds by introducing additional basis functions and continuing the splitting process in a similar fashion.
-
-\paragraph{Splitting the algebraic variables}
-We begin by considering basis functions $\mbb{P}, \mbb{Q}, \mbb{V}, \mbb{W}$ of $\mbb{K}_\mb{Q}$ and split the algebraic variables $\mbb{x}$ further as follows
-\begin{align}
-    \mbb{x} = \mbb{P} \mbb{x}_\mb{P} + \mbb{Q} \mbb{x}_\mb{Q} \label{eq:xb}.
-\end{align}
-Inserting this decomposition into \eqref{eq:iob} and multiplying once by $\mbb{V}^{\T}$ and $\mbb{W}^{\T}$ from the left yields
+\label{subsec:itc}
+As the example from \autoref{fig:ic_vcr} illustrates, there are many DAEs, including those described by MNA, which can have an index higher than one.
+In these cases the dissection index proceeds by introducing additional basis functions and continuing the splitting process in a similar fashion.
+We begin by focusing on the algebraic equation \eqref{eq:ioc_iob}, and consider basis functions $\mbb{P}(\mb{x})$, $\mbb{Q}(\mb{x})$, $\mbb{V}(\mb{x})$, $\mbb{W}(\mb{x})$ of $\mbb{K}_\mb{Q}(\mb{x})$, defined analogous to the ones for $\mb{M}(\mb{x})$.
+This allows us to further split the algebraic variables $\mbb{x}$ as follows
+\begin{align}
+    \mbb{x} = \mbb{P}(\mb{x}) \mbb{x}_\mb{P} + \mbb{Q}(\mb{x}) \mbb{x}_\mb{Q}. \label{eq:itc_xb}
+\end{align}
+Inserting this splitting into \eqref{eq:ioc_iob}, and multiplying once by $\mbb{V}^\T (\mb{x})$ and $\mbb{W}^\T (\mb{x})$ each from the left, splits the algebraic equation into two parts
 \begin{subequations}
-    \label{eq:iti}
-    \begin{align}
-        \mbb{V}^{\T} \mbb{K}_\mb{P} \mbt{x} + \mbb{V}^{\T} \mbb{K}_\mb{Q} \mbb{P} \mbb{x}_\mb{P} + \mbb{V}^{\T} \mbb{f}(t) &= \mb{0} \label{eq:itia}\\
-        \mbb{W}^{\T} \mbb{K}_\mb{P} \mbt{x} + \mbb{W}^{\T} \mbb{f}(t) &= \mb{0} \label{eq:itib}.
+    \label{eq:itc_iti}
+    \begin{align}
+        \mbb{V}^\T (\mb{x}) \mbb{K}_\mb{P}(\mb{x}) \mbt{x} + \mbb{V}^\T (\mb{x}) \mbb{K}_\mb{Q}(\mb{x}) \mbb{P}(\mb{x}) \mbb{x}_\mb{P} + \mbb{V}^\T (\mb{x}) \mbb{f}(t) &= \mb{0} \label{eq:itc_itia}\\
+        \mbb{W}^\T (\mb{x}) \mbb{K}_\mb{P}(\mb{x}) \mbt{x} + \mbb{W}^\T (\mb{x}) \mbb{f}(t) &= \mb{0}. \label{eq:itc_itib}
     \end{align}
 \end{subequations}
-
-\paragraph{Splitting the differential variables}
-We now need basis functions $\mbt{P}, \mbt{Q}$ of $\mbb{W}^{\T} \mbb{K}_\mb{P}$ to split the differential variables $\mbt{x}$ as
-\begin{align}
-    \mbt{x} = \mbt{P} \mbt{x}_\mb{P} +
-    \mbt{Q} \mbt{x}_\mb{Q} \label{eq:xt}
-\end{align}
-and inserting this splitting into \eqref{eq:itib} then allows for an explicit description of $\mbt{x}_\mb{P}$
-\begin{align*}
-    \mbt{x}_\mb{P} = -\big( \mbb{W}^{\T} \mbb{K}_\mb{P} \mbt{P} \big)^{-1} \mbb{W}^{\T} \mbb{f}(t) \eqqcolon \mbt{f}_\mb{P}(t),
-\end{align*}
-since $\mbb{W}^{\T} \mbb{K}_\mb{P} \mbt{P}$ is assumed to be regular (otherwise the DAE is underdetermined\cite{jansen2014}).
-
-\paragraph{Splitting the solution}
-Similarly, inserting \eqref{eq:xt} into \eqref{eq:itia} gives a description of $\mbb{x}_\mb{P}$ in terms of $\mbt{x}_\mb{Q}$ as $\mbb{V}^{\T} \mbb{K}_\mb{Q} \mbb{P}$ is guaranteed to be nonsingular
-\begin{align*}
-    \mbb{x}_\mb{P} = -\big( \mbb{V}^{\T} \mbb{K}_\mb{Q} \mbb{P} \big)^{-1} \big( \mbb{V}^{\T} \mbb{K}_\mb{P} \mbt{Q} \mbt{x}_\mb{Q} + \mbb{V}^{\T} \mbb{K}_\mb{P} \mbt{P} \mbt{x}_\mb{P} + \mbb{V}^{\T} \mbb{f}(t) \big) \eqqcolon \mbb{K} \mbt{x}_\mb{Q} + \mbb{f}_\mb{P}(t).
-\end{align*}
-Expanding $\mbt{x}$ and $\mbb{x}$ in \eqref{eq:ioa}, according to \eqref{eq:xb} and \eqref{eq:xt} respectively, and using the above expressions for $\mbt{x}_\mb{P}$ and $\mbb{x}_\mb{P}$ gives
-\begin{align}
-    \mbt{M} \mbt{Q} \ddt \mbt{x}_\mb{Q} + \big( \mbt{K}_\mb{P} \mbt{Q} + \mbt{K}_\mb{Q} \mbb{P} \mbb{K} \big) \mbt{x}_\mb{Q} + \mbt{K}_\mb{Q} \mbb{Q} \mbb{x}_\mb{Q} + \mbh{f}(t) = \mb{0} \label{eq:itii},
-\end{align}
-where we define
-\begin{align*}
-    \mbh{f}(t) \coloneqq \mbt{M} \mbt{P} \ddt \mbt{f}_\mb{P}(t) + \mbt{K}_\mb{P} \mbt{P} \mbt{f}_\mb{P}(t) + \mbt{K}_\mb{Q} \mbb{P} \mbb{f}_\mb{P}(t) + \mbt{f}(t).
-\end{align*}
-
-\paragraph{Splitting the system}
-Finally, we require basis functions $\mbt{V}$ and $\mbt{W}$ of $\mbt{M} \mbt{Q}$.
-Multiplying \eqref{eq:itii} from the left by $\mbt{V}^{\T}$ and $\mbt{W}^{\T}$ and redefining $\mbt{x}_2 \coloneqq \mbt{x}_\mb{Q}$ and $\mbb{x}_2 \coloneqq \mbb{x}_\mb{Q}$ yields
+We can now also split the differential variables $\mbt{x}$ further, by using basis functions $\mbt{P}(\mb{x})$ and $\mbt{Q}(\mb{x})$ of $\mbb{W}^\T (\mb{x}) \mbb{K}_\mb{P}(\mb{x})$
+\begin{align}
+    \mbt{x} = \mbt{P}(\mb{x}) \mbt{x}_\mb{P} +
+    \mbt{Q}(\mb{x}) \mbt{x}_\mb{Q}, \label{eq:itc_xt}
+\end{align}
+and inserting this splitting into the second algebraic equation \eqref{eq:itc_itib} yields
+\begin{align}
+    \mbb{W}^\T (\mb{x}) \mbb{K}_\mb{P}(\mb{x}) \mbt{P}(\mb{x}) \mbt{x}_\mb{P} + \mbb{W}^\T (\mb{x}) \mbb{f}(t) &= \mb{0}. \label{eq:itc_xtp}
+\end{align}
+We now make the following two assumptions.
+\textcolor{red}{\begin{assumption}
+    \label{as:itc_xtpdi}
+    The matrix product $\mbb{W}^\T (\mb{x}) \mbb{K}_\mb{P}(\mb{x}) \mbt{P}(\mb{x}) \mbt{x}_\mb{P}$ is regular.
+\end{assumption}
+\begin{assumption}
+    \label{as:itc_xtp}
+    The basis functions $\mbt{P}$ and $\mbt{Q}$ of $\mbb{W}^\T (\mb{x}) \mbb{K}_\mb{P}(\mb{x})$ are constant and \eqref{eq:itc_xtp} possesses a locally unique solution for $\mbt{x}_\mb{P}$ in terms of $\mbt{x}_\mb{Q}$ and $t$.
+\end{assumption}
+We note that while \autoref{as:itc_xtpdi} is equivalent to the DAE not being underdetermined\cite{jansen2014}, \autoref{as:itc_xtp} is more practical and important for the implementation, but not for the dissection index itself.
+In fact our approach still works if the basis functions $\mbt{P}(\mbt{x}_\mb{Q})$ and $\mbt{Q}(\mbt{x}_\mb{Q})$ depend on $\mbt{x}_\mb{Q}$, however we focus on the stronger assumption here, since it shortens the expressions in the following without impacting the general idea and \autoref{re:dim_xtp} shows that MNA fulfills an even stronger condition than \autoref{as:itc_xtp}.}
+
+Having a locally unique solution for $\mbt{x}_\mb{P}$ in terms of $\mbt{x}_\mb{Q}$ and $t$ at hand, we now turn to the first algebraic equation \eqref{eq:itc_itia}.
+Similar to \eqref{eq:itc_xtp}, inserting the splitting \eqref{eq:itc_xt} of the differential variables into \eqref{eq:itc_itia} gives a system with a locally unique solution for $\mbb{x}_\mb{P}$ in terms of $\mbt{x}_\mb{Q}$, $\mbt{x}_\mb{P}$, $\mbb{x}_\mb{Q}$ and $t$, as $\mbb{V}^\T (\mb{x}) \mbb{K}_\mb{Q}(\mb{x}) \mbb{P}(\mb{x})$ is regular by construction\cite{jansen2014}
+\begin{align}
+    \mbb{V}^\T (\mb{x}) \mbb{K}_\mb{P}(\mb{x}) \mbt{P} \mbt{x}_\mb{P} + \mbb{V}^\T (\mb{x}) \mbb{K}_\mb{P}(\mb{x}) \mbt{Q} \mbt{x}_\mb{Q} + \mbb{V}^\T (\mb{x}) \mbb{K}_\mb{Q}(\mb{x}) \mbb{P}(\mb{x}) \mbb{x}_\mb{P} + \mbb{V}^\T (\mb{x}) \mbb{f}(t) &= \mb{0}.
+\label{eq:itc_xbp}
+\end{align}
+Finally, we move towards the decoupled index two system by expanding $\mbt{x}$ and $\mbb{x}$ in \eqref{eq:ioc_ioa}, according to \eqref{eq:itc_xt} and \eqref{eq:itc_xb} respectively,
+\begin{align}
+    \mbt{M}(\mb{x}) \mbt{P} \ddt \mbt{x}_\mb{P} + \mbt{M}(\mb{x}) \mbt{Q} \ddt \mbt{x}_\mb{Q} + \mbt{K}_\mb{P}(\mb{x}) \mbt{P} \mbt{x}_\mb{P} + \mbt{K}_\mb{P}(\mb{x}) \mbt{Q} \mbt{x}_\mb{Q} + \mbt{K}_\mb{Q}(\mb{x}) \mbb{P}(\mb{x}) \mbb{x}_\mb{P} + \mbt{K}_\mb{Q}(\mb{x}) \mbb{Q}(\mb{x}) \mbb{x}_\mb{Q} + \mbt{f}(t) &= \mb{0}.
+\label{eq:itc_itii}
+\end{align}
+Using basis functions $\mbt{V}(\mb{x})$ and $\mbt{W}(\mb{x})$ of $\mbt{M}(\mb{x}) \mbt{Q}$, we can now split \eqref{eq:itc_itii} further by multiplying from the left by $\mbt{V}^\T (\mb{x})$ and $\mbt{W}^\T (\mb{x})$ once each.
+Reordering then yields
 \begin{subequations}
-    \label{eq:itiii}
-    \begin{align}
-            \underbrace{\mbt{V}^{\T} \mbt{M} \mbt{Q}}_{\eqqcolon \mbt{M}_2} \ddt \mbt{x}_2 + \underbrace{\mbt{V}^{\T} \big( \mbt{K}_\mb{P} \mbt{Q} + \mbt{K}_\mb{Q} \mbb{P} \mbb{K} \big)}_{\eqqcolon \mbt{K}_{\mb{P}_2}} \mbt{x}_2 + \underbrace{\mbt{V}^{\T} \mbt{K}_\mb{Q} \mbb{Q}}_{\eqqcolon \mbt{K}_{\mb{Q}_2}} \mbb{x}_2 + \underbrace{\mbt{V}^{\T} \mbh{f}(t)}_{\eqqcolon \mbt{f}_2(t)} &= \mb{0} \label{eq:itiiia}\\
-            \underbrace{\mbt{W}^{\T} \big( \mbt{K}_\mb{P} \mbt{Q} + \mbt{K}_\mb{Q} \mbb{P} \mbb{K} \big)}_{\eqqcolon \mbb{K}_{\mb{P}_2}} \mbt{x}_2 + \underbrace{\mbt{W}^{\T} \mbt{K}_\mb{Q} \mbb{Q}}_{\eqqcolon \mbb{K}_{\mb{Q}_2}} \mbb{x}_2 + \underbrace{\mbt{W}^{\T} \mbh{f}(t)}_{\eqqcolon \mbb{f}_2(t)} &= \mb{0}.
-\label{eq:itiiib}
+    \label{eq:itc_itiii}
+    \begin{align}
+            \mbt{V}^\T (\mb{x}) \mbt{M}(\mb{x}) \mbt{Q} \ddt \mbt{x}_\mb{Q} + \mbt{V}^\T (\mb{x}) \mbt{K}_\mb{P}(\mb{x}) \mbt{Q} \mbt{x}_\mb{Q} + \mbt{V}^\T (\mb{x}) \mbt{K}_\mb{Q}(\mb{x}) \mbb{Q}(\mb{x}) \mbb{x}_\mb{Q} + \mbt{V}^\T (\mb{x}) \big( \mbt{M}(\mb{x}) \mbt{P} \ddt \mbt{x}_\mb{P} + \mbt{K}_\mb{P}(\mb{x}) \mbt{P} \mbt{x}_\mb{P} + \mbt{K}_\mb{Q}(\mb{x}) \mbb{P}(\mb{x}) \mbb{x}_\mb{P} + \mbt{f}(t) \big) &= \mb{0} \label{eq:itc_itiiia}\\
+            \mbt{W}^\T (\mb{x}) \mbt{K}_\mb{P}(\mb{x}) \mbt{Q} \mbt{x}_\mb{Q} + \mbt{W}^\T (\mb{x}) \mbt{K}_\mb{Q}(\mb{x}) \mbb{Q}(\mb{x}) \mbb{x}_\mb{Q} + \mbt{W}^\T (\mb{x}) \big( \mbt{M}(\mb{x}) \mbt{P} \ddt \mbt{x}_\mb{P} + \mbt{K}_\mb{P}(\mb{x}) \mbt{P} \mbt{x}_\mb{P} + \mbt{K}_\mb{Q}(\mb{x}) \mbb{P}(\mb{x}) \mbb{x}_\mb{P} + \mbt{f}(t) \big) &= \mb{0}.
+\label{eq:itc_itiiib}
     \end{align}
 \end{subequations}
-We observe that \eqref{eq:itiii} is of the same form as \eqref{eq:io} and that $\mbt{M}_2$ is again regular by construction.
-Therefore, if $\mbb{K}_{\mb{Q}_2}$ is also regular we can extract an ODE analogous to the previous case by solving \eqref{eq:itiiib} for $\mbb{x}_2$ (in terms of $\mbt{x}_2$) and inserting that solution into \eqref{eq:itiiia}.
-The DAE is then said to have \textbf{dissection index} $\nu = 2$ and we call $\mbt{x}_2$ the differential variables and $\mbb{x}_2$ the algebraic variables.
-Furthermore, since \eqref{eq:itiii} and \eqref{eq:io} are structurally identical, we may simply apply the second step of the decoupling procedure to \eqref{eq:itiii} in order to deal with even higher index DAEs.
-
-To clarify the naming convention for the following sections, we call $\mbt{x}_2$ and $\mbb{x}_2$ the purely differential and algebraic variables of an index two DAE respectively, whereas $\mbb{x}$ denotes the explicit (obvious) algebraic variables and $\mbt{x}$ still contains hidden (implicit) algebraic variables that are identified in the second step of the dissection concept.
-
-\subsection{Dissection concept and MNA}
-\label{subsec:dcmna}
-Before we consider example circuits, we take a brief look at the first step of the dissection concept for MNA. For this, we recall the system of MNA \eqref{eq:mnaii}
-\begin{align*}
-    \begin{bmatrix}
-        \A{C}^{\phantom{\T}} \mb{C} \big( \AT{C} \vphi \big) \AT{C} & &\\
-        & \mb{L}(\i{L}) &\\
-        & & \mb{0}
-    \end{bmatrix} \ddt \begin{bmatrix}
-        \vphi\\
-        \i{L}\\
-        \i{V}
-    \end{bmatrix} + \begin{bmatrix}
-        \A{R}^{\phantom{\T}} \mb{G} \big( \AT{R} \vphi \big) \AT{R} & \A{L} & \A{V}\\
-        -\AT{L} & &\\
-        -\AT{V} & &
-    \end{bmatrix} \begin{bmatrix}
-        \vphi\\
-        \i{L}\\
-        \i{V}
-    \end{bmatrix} + \begin{bmatrix}
-        \A{I} \i{s}(t)\\
-        \mb{0}\\
-        \v{s}(t)
-    \end{bmatrix} = \mb{0}
-\end{align*}
-and determine the basis functions $\mb{P}$ and $\mb{Q}$.
-Taking into account the standard assumption, compare \autoref{sec:mna}, that $\mb{L}(\i{L})$ and $\mb{C} \big( \AT{C} \vphi \big)$ are positive definite, we find\cite{jansen2014}
-\begin{align*}
-    \mb{Q} = \begin{bmatrix}
-        \mb{Q}_\mr{C} &\\
-        & \mb{0}\\
-        & \mb{I}
-    \end{bmatrix}, \quad \mb{P} = \begin{bmatrix}
-        \mb{P}_\mr{C} &\\
-        & \mb{I}\\
-        & \mb{0}
-    \end{bmatrix},
-\end{align*}
-where $\mr{im\, } \mb{Q}_\mr{C} = \ker \AT{C}$, $\mb{0}$ and $\mb{I}$ are zero and identity matrices of appropriate dimension and the columns of $\mb{Q}_\mr{C}$ and $\mb{P}_\mr{C}$ together form a basis of $\mathbb{R}^{n_\varphi}$ with $n_\varphi$ the dimension of $\vphi$.
-The description of $\mb{Q}_\mr{C}$ results from the fact that (with $\mb{y} \coloneqq \AT{C} \mb{x}$)
-\begin{align*}
-    \mb{x}^{\T} \A{C}^{\phantom{\T}} \mb{C} \big( \AT{C} \vphi \big) \AT{C} \mb{x} = \mb{y}^{\T} \mb{C} \big( \AT{C} \vphi \big) \mb{y} > 0, \quad \forall \mb{y} \in \mathbb{R}^{n_\varphi} \setminus \{ \mb{0} \},
-\end{align*}
-since $\mb{C} \big( \AT{C} \vphi \big)$ is positive definite and hence the kernel of $\A{C}^{\phantom{\T}} \mb{C} \big( \AT{C} \vphi \big) \AT{C}$ is determined by the kernel of $\AT{C}$.
-As $\A{C}$ is an incidence (constant) matrix, we find that $\mb{Q}$ and $\mb{P}$ are also constant.
+We observe that \eqref{eq:itc_itiii} is of a form similar to \eqref{eq:ioc_io} and that $\mbt{V}^\T (\mb{x}) \mbt{M}(\mb{x}) \mbt{Q}$ is again regular by construction\cite{jansen2014}.
+Together with \eqref{eq:itc_xtp} and \eqref{eq:itc_xbp} providing locally unique solutions for $\mbt{x}_\mb{P}$ and $\mbb{x}_\mb{P}$ respectively, this motivates the following definition.
+\textcolor{red}{\begin{definition}
+    \label{def:itc_it}
+     The DAE has dissection index $\nu = 2$ if $\mbt{W}^\T (\mb{x}) \mbt{K}_\mb{Q}(\mb{x}) \mbb{Q}(\mb{x})$ is regular.
+\end{definition}
+This again follows from the purely algebraic equation \eqref{eq:itc_itiiib} having a locally unique solution for $\mbb{x}_\mb{Q}$ in terms of $\mbt{x}_\mb{Q}$, $\mbt{x}_\mb{P}$, $\mbb{x}_\mb{P}$ and $t$, given that $\mbt{W}^\T (\mb{x}) \mbt{K}_\mb{Q}(\mb{x}) \mbb{Q}(\mb{x})$ is regular.
+The differential part \eqref{eq:itc_itiiia} then describes an ODE in the index two differential variables $\mbt{x}_\mb{Q}$, analogous to the previous case, and we call $\mbt{x}_\mb{Q}$ the differential variables and $[\mbt{x}_\mb{P}, \mbb{x}_\mb{P}, \mbb{x}_\mb{Q}]^\T$ the algebraic variables.
+We note that the procedure may be continued for even higher index DAEs, by repeating the steps of the index two case with $\mbt{x}_\mb{Q}$ and $\mbb{x}_\mb{Q}$ playing the roles of $\mbt{x}$ and $\mbb{x}$ respectively.}
 
 \paragraph{First example circuit}
-\begin{figure}
+\begin{figure}[b]
     \begin{center}
@@
     \caption{First example circuit: simple diode oscillator.}
-    \label{fig:do1}
-\end{figure}
-We now demonstrate the dissection concept by applying it to the example circuit given in \autoref{fig:do1}.
-The circuit contains a voltage source $v_\mathrm{s}(t)$, a linear resistor with conductance $G$, a linear capacitor with capacitance $C$, a linear inductor with inductance $L$, as well as a diode $\mathrm{D}$ that is modeled by a nonlinear resistance $g_\mathrm{D}(\varphi_3)$.
-Comparing the conditions of the topological index result from \autoref{sec:mna} with the example circuit shows that the circuit has index $\nu = 1$, thus we only have to perform the first step of the dissection concept.
-
-We begin by writing out the system derived from applying MNA to the example
+    \label{fig:itc_do1}
+\end{figure}
+We now demonstrate the dissection index by applying it to the example circuit given in \autoref{fig:itc_do1}.
+The circuit contains a voltage source $v_\mr{s}(t)$, a linear resistor with resistance $R$, a linear capacitor with capacitance $C$, a linear inductor with inductance $L$, as well as a diode $\mr{D}$ that is modeled by a nonlinear resistance $g_\mr{D}(\varphi_3)$.
+Comparing the conditions of \autoref{th:ic_imna} with the example circuit shows that the circuit has index $\nu = 1$, thus we only have to perform the first step of the dissection index.
+
+We begin by writing out the system obtained from applying MNA to the example
 \begin{align*}
@@
         v_\mr{s}(t)
-    \end{bmatrix} = \mb{0}.
-\end{align*}
-For the first basis functions $\mb{Q}$ and $\mb{P}$ we find
+    \end{bmatrix} = \mb{0},
+\end{align*}
+where $G = 1/R$ is the inverse of the resistance. For the first basis functions $\mb{Q}$ and $\mb{P}$ we find
 \begin{align*}
@@
 \end{align*}
-and as the matrix equivalent to $\mb{M}$ in the context of \eqref{eq:dae} is symmetric, we have $\mb{W} = \mb{Q}$ and $\mb{V} = \mb{P}$ for the remaining two basis functions.
-Using these, we find the new unknowns $\mbt{x} = [\varphi_3\ i_\mr{L}]^{\T}$ and $\mbb{x} = [\varphi_1\ \varphi_2\ i_\mr{V}]^{\T}$, which allow us to obtain the systems corresponding to \eqref{eq:ioa} and \eqref{eq:iob}
+and as $\mb{M}(\mb{x})$ in the context of \eqref{eq:di_dae} is symmetric in this case, we have $\mb{W} = \mb{Q}$ and $\mb{V} = \mb{P}$ for the remaining two basis functions.
+Using these, we split the unknowns into $\mbt{x} = [\varphi_3, i_\mr{L}]^\T$ and $\mbb{x} = [\varphi_1, \varphi_2, i_\mr{V}]^\T$ according to \eqref{eq:ioc_x}, which allows us to obtain the systems corresponding to \eqref{eq:ioc_ioa} and \eqref{eq:ioc_iob}
 \begin{subequations}
-    \label{eq:ei}
+    \label{eq:itc_ei}
     \begin{align}
@@
             i_\mr{V}
-        \end{bmatrix} &= \mb{0} \label{eq:eia}\\
+        \end{bmatrix} &= \mb{0} \label{eq:itc_eia}\\
         \begin{bmatrix}
@@
             0\\
-            v_\mr{s}
-        \end{bmatrix} &= \mb{0} \label{eq:eib}.
+            v_\mr{s}(t)
+        \end{bmatrix} &= \mb{0}. \label{eq:itc_eib}
     \end{align}
 \end{subequations}
-Solving \eqref{eq:eib} for $\mbb{x}$ and subsequently inserting the solution into \eqref{eq:eia} then gives an ODE and a purely algebraic system as promised
+Since \eqref{eq:itc_eib} is linear, we can explicitly solve for $\mbb{x}$ in this case.
+Subsequently inserting this solution into \eqref{eq:itc_eia} then gives an ODE and a purely algebraic system as promised
 \begin{align}
@@
         0\\
-        -v_\mr{s}
-    \end{bmatrix} = \mb{0} \quad \text{and} \quad \begin{bmatrix}
+        -v_\mr{s}(t)
+    \end{bmatrix} = \mb{0}, \quad \begin{bmatrix}
         \varphi_1\\
@@
     \end{bmatrix} = \begin{bmatrix}
-        v_\mr{s}\\
-        v_\mr{s} - R i_\mr{L}\\
+        v_\mr{s}(t)\\
+        v_\mr{s}(t) - R i_\mr{L}\\
         -i_\mr{L}
-    \end{bmatrix}, \label{eq:do1}
-\end{align}
-where $R = 1/G$ is the inverse of the conductance.
+    \end{bmatrix}. \label{eq:itc_do1}
+\end{align}
 
 \paragraph{Second example circuit}
-\begin{figure}
+\begin{figure}[b]
     \begin{center}
@@
     \caption{Second example circuit: simple diode oscillator with current instead of voltage source.}
-    \label{fig:do2}
-\end{figure}
-We derive a second example from the first by substituting a current source $i_\mr{s}(t)$ for the voltage source, compare \autoref{fig:do2}.
-Looking at the index criteria from \autoref{sec:mna}, we observe that this circuit has index $\nu = 2$, as there now is a cutset consisting of the inductor and current source.
-Therefore, we need to execute two steps of the dissection concept in order to split the equations into purely differential and algebraic parts respectively.
+    \label{fig:itc_do2}
+\end{figure}
+We derive a second example circuit from the first by substituting a current source $i_\mr{s}(t)$ for the voltage source, compare \autoref{fig:itc_do2}.
+Looking at the index criteria from \autoref{th:ic_imna}, we observe that this circuit has index $\nu = 2$, as there now is a cutset consisting of the inductor and current source.
+Therefore, we need to execute two steps of the dissection index in order to split the equations into purely differential and algebraic parts respectively.
 
@@
         1
-    \end{bmatrix}, \quad \mbb{P} = \bar{\mathbf{V}} = \begin{bmatrix}
+    \end{bmatrix}, \quad \mbb{P} = \mbb{V} = \begin{bmatrix}
         1\\
@@
 \end{align*}
-and finally obtain an ODE in one variable only and a second purely algebraic system that recovers the remaining unknowns
-\begin{align}
-    C \ddt \varphi_3 + g_\mr{D}(\varphi_3) \varphi_3 - i_\mr{s} = 0 \quad
-    \text{and} \quad \begin{bmatrix}
+and finally obtain an ODE in one differential variable only, together with a purely algebraic system that recovers the remaining algebraic variables
+\begin{align}
+    C \ddt \varphi_3 + g_\mr{D}(\varphi_3) \varphi_3 - i_\mr{s}(t) = 0, \quad \begin{bmatrix}
         \varphi_1\\
@@
     \end{bmatrix} = \begin{bmatrix}
-        \varphi_2 + R i_\mr{s}\\
-        \varphi_3 + L \ddt i_\mr{s}\\
-        i_\mr{s}
-    \end{bmatrix} \label{eq:do2}.
-\end{align}
-A key observation at this point is that all basis functions of the second step are also constant, as was the case for the first step.
-Although we did not continue the abstract decoupling process for MNA in \autoref{subsec:dcmna}, this is indeed a general phenomenon.
-The decoupling can be organized based on topological conditions only\cite{jansen2014}, which agrees with the intuition given by the index result from \autoref{sec:mna}.
-% \cite[Section 7.1]{jansen2014}
-We remark that a similar result for circuits including controlled sources exists, however only for the first stage of the decoupling, as it is framed in the context of semi-explicit methods for which one only requires a DAE in semi-explicit form\cite{jansen2014}.
+        \varphi_2 + R i_\mr{s}(t)\\
+        \varphi_3 + L \ddt i_\mr{s}(t)\\
+        i_\mr{s}(t)
+    \end{bmatrix} \label{eq:itc_do2}.
+\end{align}
+
+We note that all basis functions of the second step are also constant for this example, as was the case for the first step.
+As hinted at in \autoref{sec:di}, it is possible to find a purely topological decoupling based on the dissection index\cite{jansen2014}, which agrees with the intuition given by the index result from \autoref{th:ic_imna}.
+\textcolor{red}{When considering large circuits, this topological decoupling along with its topological basis functions is to be preferred over other basis function choices, as it avoids the numerical computation of the basis functions, which becomes prohibitively expensive for large systems.} We also note that a similar topological decoupling result exists for circuits including controlled sources, however only for the first step of the decoupling, as it is framed in the context of semi-explicit methods for which one only requires a DAE in semi-explicit form\cite{jansen2014}.
 An extension of this result to circuits of higher index is within the scope of further research.
@@
 \label{sec:ial}
-In the following, we outline how to use the dissection concept in the context of machine learning.
-The workflow is illustrated in \autoref{fig:ial} and directly follows the structure of the dissection concept.
-We give a general description in a first step, followed by examples using the two circuits from \autoref{sec:dc}.
+\begin{figure}[b]
+    \begin{center}
+        \includegraphics[width=.7\textwidth]{ial}
+    \end{center}
+    \caption{Schematic workflow of index-aware learning.}
+    \label{fig:ial_ial}
+\end{figure}
+In the following, we outline the use of the dissection index in the context of machine learning.
+The workflow is illustrated in \autoref{fig:ial_ial} and directly follows the structure of the dissection index.
+We give a general description in a first step, followed by examples using the two circuits from \autoref{fig:itc_do1} and \autoref{fig:itc_do2}.
 \begin{enumerate}
-    \item Our approach begins by decoupling a given DAE into an ODE and a purely algebraic equation (AE) following the main steps of the dissection concept.
-
-    \item Afterwards, only the differential degrees of freedom (DOFs) of the ODE are learned.
-For a DAE of index $\nu$, these would be the entries of $\tilde{\mathbf{x}}_\nu$ using the notation of \autoref{sec:dc}.
-
-    \item The remaining algebraic DOFs $\bar{\mathbf{x}}_\nu$ may then be reconstructed using the algebraic equation.
+    \item Our approach begins by performing the decoupling of a given DAE into an ODE and a purely algebraic equation (AE) following the main steps of the dissection index.
+    \item Afterwards, only the differential variables of the ODE are learned.
+For a DAE of index one, these would be the entries of $\mbt{x}$, and for a DAE of index two, the entries of $\mbt{x}_\mb{Q}$ using the notation of \autoref{sec:di}.
+    \item The remaining algebraic variables, $\mbb{x}$ for index one or $[\mbt{x}_\mb{P}, \mbb{x}_\mb{P}, \mbb{x}_\mb{Q}]^\T$ for index two, may then be reconstructed using the algebraic equations.
 \end{enumerate}
-We remark that the identification of the differential DOFs and thus the advantages of the second point are in principle available for any Spice based simulator via the topological decoupling\cite{jansen2014}, whereas the reconstruction of the algebraic DOFs requires an additional implementation.
-However, it is possible to trade some of the accuracy of our approach for an easier implementation, see \autoref{subsec:aa}.
-\begin{figure}
-    \begin{center}
-        \includegraphics[width=.7\textwidth]{ial}
-    \end{center}
-    \caption{Schematic workflow of index-aware learning.}
-    \label{fig:ial}
-\end{figure}
-
-In terms of the example circuits from \autoref{subsec:dcmna} the workflow amounts to the following: for the first example we only consider $\varphi_3$ and $i_\mr{L}$ as variables that need to be learned and for the second example solely $\varphi_3$ is left.
-All remaining degrees of freedom may be recovered using \eqref{eq:do1} or \eqref{eq:do2} respectively.
-Thus the learning effort is already reduced quite significantly in these two examples; from five to two variables in the first and from four down to one variable in the second example.
+We remark that the identification of the differential variables, and thus the advantages of the second point, are in principle available for any Spice based simulator via the purely topological decoupling\cite{jansen2014}, whereas the reconstruction of the algebraic variables requires an additional implementation.
+We summarize the important steps of this additional implementation for the index one and two cases below.
+
+\paragraph{Index one case}
+\textcolor{red}{In order to recover the algebraic variables at time $t$ in the index one case, we only need to solve \eqref{eq:ioc_iob} for $\mbb{x}(t)$ using the learned $\mbt{x}(t)$.}
+
+\paragraph{Index two case}
+\textcolor{red}{In the index two case, we start by solving \eqref{eq:itc_xtp} for $\mbt{x}_\mb{P}(t)$ using the learned $\mbt{x}_\mb{Q}(t)$.
+Since we also require the derivative $\ddt \mbt{x}_\mb{P}(t)$ in \eqref{eq:itc_itiiib}, we consider a small time increment $\Delta t$ and approximate the derivative using a backward difference
+\begin{align*}
+    \ddt \mbt{x}_\mb{P}(t) \approx \frac{\mbt{x}_\mb{P}(t + \Delta t) - \mbt{x}_\mb{P}(t)}{\Delta t}.
+\end{align*}
+We note that this only reflects our implementation; in principle any finite difference (or similar) approximation is possible.
+Finally, we determine $\mbb{x}_\mb{P}(t)$ and $\mbb{x}_\mb{Q}(t)$ by jointly solving \eqref{eq:itc_xbp} and \eqref{eq:itc_itiiib} using the learned $\mbt{x}_\mb{Q}(t)$, $\mbt{x}_\mb{P}(t)$ and the approximation of $\ddt \mbt{x}_\mb{P}(t)$.}
+
+\paragraph{Example circuits}
+In terms of the example circuits from \autoref{sec:di}, the workflow amounts to the following: for the first example we consider $\varphi_3$ and $i_\mr{L}$ as the differential variables that need to be learned, and for the second example only $\varphi_3$ is left.
+All the remaining algebraic variables may be recovered using \eqref{eq:itc_do1} or \eqref{eq:itc_do2} respectively.
+Thus the learning effort is already reduced quite significantly in these two examples; from five to two variables in the first and from four down to one variable in the second.
 Another key benefit comes from the fact that the reconstructed algebraic variables exactly fulfill the inherent constraints of the DAE.
-This means that even though the learned solution variables (think of $\varphi_3$ for example) are only approximations, the reconstructions (e.g.
-$\varphi_1$ or $\varphi_2$) will still be consistent.
+This means that even though the learned solution variables (think of $\varphi_3$ for example) are only approximations, the reconstructions (e.g.~$\varphi_1$ or $\varphi_2$) will still be consistent.
 This may be of great importance for systems where the physical interpretability of the solution depends on it satisfying the constraints.
@@
 There is yet another, maybe less expected, benefit that might occur.
-Looking at the decoupled system from \eqref{eq:do2} we find that the resistance parameter $R$ and the inductance parameter $L$ only appear in the algebraic equation.
-In terms of our original goal of speeding up design optimization or uncertainty quantification, where a lot of solutions for varying parameter values are required, this means the following: instead of needing to solve the full system for a given combination of $R$ and $L$, we can instead merely solve the algebraic equation to obtain the full solution.
-While the algebraic equation might be more complicated than the simple linear relations of \eqref{eq:do1} and \eqref{eq:do2} in general, solving it is almost certainly much faster than having to integrate the entire system in time.
+Looking at the decoupled system from \eqref{eq:itc_do2} we find that the resistance parameter $R$ and the inductance parameter $L$ only appear in the algebraic equation.
+In terms of our original goal of speeding up design optimization or uncertainty quantification, where a lot of solutions for varying parameter values are required, this means the following: instead of having to solve the full system for a given combination of $R$ and $L$, we can instead simply solve the algebraic equation to obtain the full solution.
+While the algebraic equation might be more complicated than the simple linear relations of \eqref{eq:itc_do1} and \eqref{eq:itc_do2} in general, solving it is almost certainly much faster than having to integrate the entire system in time.
 This becomes an even bigger advantage when knowledge about the solution is only required for specific points in time, since the algebraic equation may be solved pointwise.
-
-Lastly, we emphasize that the approach is independent of the particular machine learning technique that is used for predicting the differential DOFs.
+\textcolor{red}{As of now however, we have no knowledge of an easy way to automatically determine which parameters appear in the ODE.
+But when combined with a sequential learning strategy, such as the one outlined in \autoref{sec:ne}, there may still be computational savings due to the learning method requiring less samples for the parameters not appearing in the ODE.}
+
+Lastly, we emphasize that the approach is independent of the particular machine learning method that is used for learning the differential variables.
 Thus methods developed especially for ODEs may be employed and exchanged depending on the problem at hand.
@@
 \label{sec:ne}
-Before we present numerical results regarding the two example circuits introduced in \autoref{subsec:dcmna}, we want to provide some background on our machine learning method of choice, Gaussian processes (GPs), and the particular learning strategy we employ.
+Before we present numerical results, we want to provide some background on our machine learning method of choice, Gaussian processes (GPs), and the particular learning strategy we employ.
 
 \paragraph{Gaussian processes}
-The following brief introduction is based on the textbook of Rasmussen and Williams\cite{rasmussen2006} and we refer to the book itself for more details.
-We consider the problem of approximating one component $x(t)$ of the DAE solution $\mb{x}(t)$, compare \eqref{eq:dae}, based on observations
+The following brief introduction is based on the textbook of Rasmussen and Williams\cite{rasmussen2006}, and we refer to the book itself for more details.
+We consider the problem of learning one component $x(t)$ of the DAE solution $\mb{x}(t)$, compare \eqref{eq:di_dae}, based on observations
 \begin{align*}
@@
 \end{align*}
-We will focus on the one-dimensional case for clarity of exposition, however we remark that the ideas extend to the case where the solution component $x(t, \mb{p})$ also depends on the parameters and thus more than one variable, compare also the textbook\cite{rasmussen2006}.
-A GP suitable for this problem is defined by a (prior) mean function $m: \mathbb{R} \to \mathbb{R}$ and covariance function $k: \mathbb{R} \times \mathbb{R} \to \mathbb{R}$.
+We will focus on the one-dimensional case for clarity of exposition, however we remark that the ideas extend to the case where the solution component $x(t, \mb{p})$ also depends on the parameters $\mb{p}$ and thus more than one variable, compare also the textbook\cite{rasmussen2006}.
+A GP suited for this problem is defined by a (prior) mean function $m: \mathbb{R} \to \mathbb{R}$ and covariance function $k: \mathbb{R} \times \mathbb{R} \to \mathbb{R}$.
 The learning problem is then tackled using Bayesian inference, such that one aims to obtain the posterior distribution $\hat{x}(t)$, given the observations $O$ and a point $t$, where $x$ is to be predicted.
@@
 \begin{subequations}
-    \label{eq:gp}
-    \begin{align}
-        \hat{m}(t) &= m(t) + \mb{w}(t)^\T (\mb{x} - \mb{m}) \label{eq:gpm}\\
-        \hat{k}(t) &= k(t, t) - \mb{w}(t)^\T \mb{k}(t) \label{eq:gpv},
+    \label{eq:ne_gp}
+    \begin{align}
+        \hat{m}(t) &= m(t) + \mb{w}(t)^\T (\mb{x} - \mb{m}) \label{eq:ne_gpa}\\
+        \hat{k}(t) &= k(t, t) - \mb{w}(t)^\T \mb{k}(t) \label{eq:ne_gpb},
     \end{align}
@@
 where $[\mb{K}]_{i,j} \coloneqq k(t_i, t_j)$ and $\sigma^2$ models i.i.d.~additive Gaussian noise on the observations.
-For a discussion on modeling choices where the assumptions leading to \eqref{eq:gp} are not fulfilled, see the review article by Swiler et al.\cite{swiler2020}.
+For a discussion on modeling choices where the assumptions leading to \eqref{eq:ne_gp} are not fulfilled, see the review article by Swiler et al.\cite{swiler2020}.
 
 The key model component influencing the learning process is the covariance (or kernel) function $k$, since it determines the approximation properties of the GP.
-It encodes prior knowledge about the function $x(t)$ that is to be learned, such as differentiability or characteristic length scales.
+It encodes prior knowledge about the function $x(t)$ that is to be learned, such as its differentiability or characteristic length scales.
 We opt for a radial basis function kernel, given by
@@
 
-In practice, the mean function $m$ is often taken to be zero as the data is assumed standardized and the hyperparameters are then determined by minimizing the negative log likelihood\cite{basak2021}
+In practice, the mean function $m$ is often taken to be zero as the data is assumed standardized, and the hyperparameters are then determined by minimizing the negative log likelihood\cite{basak2021}
 \begin{align*}
@@
 \paragraph{Learning strategy}
-The learning strategy aims to exploit one of the key features of GPs: that they provide both a mean prediction $\hat{m}$ and an associated variance estimate $\hat{k}$, as detailed in \eqref{eq:gp}.
-We use these properties in conjunction with a greedy sample selection strategy, that starts out with a small number of training data points and adds additional samples based on the variance estimate of the GP.
+The learning strategy aims to exploit one of the key features of GPs: they provide both a mean prediction $\hat{m}$ and an associated variance estimate $\hat{k}$, as detailed in \eqref{eq:ne_gp}.
+We use these properties in conjunction with a sequential sample selection strategy, that starts out with a small number of training data and adds further samples based on the variance estimate of the GP.
 This idea is not new, see again the textbook of Rasmussen and Williams\cite{rasmussen2006} for more references and details, however we still want to outline our particular approach to make the results better interpretable and reproducible.
@@
 \begin{enumerate}
-    \item We select a grid of time points $T = \{ t_i: 1 \leq i \leq N_T \}$ and parameter values $P = \{ \mb{p}_i: 1 \leq i \leq N_P \}$, where each $\mb{p}_i$ represents a specific combination of parameter values, for which we want to model the solution of the DAE using a GP.
+    \item We select a grid of time points $T = \{ t_i: 1 \leq i \leq N_T \}$ and parameter values $P = \{ \mb{p}_i: 1 \leq i \leq N_P \}$, where each $\mb{p}_i$ represents a specific combination of parameter values, for which we want to learn the solution of the DAE using a GP.
     \item We select a subset $D \subset T \times P$ and use the corresponding solutions for the initial training of a GP.
-    \item As a termination criterion, we compute the posterior mean $\hat{m}(t, \mb{p})$ using \eqref{eq:gpm} in all grid points $(t, \mb{p}) \in T \times P$ and check whether the relative prediction error
-    \begin{align*}
-        e \coloneqq \frac{\| \mbh{m} - \mb{x} \|_2}{ \| \mb{x} \|_2},
+    \item As a termination criterion, we compute the posterior mean $\hat{m}(t, \mb{p})$ using \eqref{eq:ne_gpa} in all grid points $(t, \mb{p}) \in T \times P$ and check whether the relative prediction error
+    \begin{align*}
+        e \coloneqq \frac{\| \mbh{m} - \mb{x} \|_2}{ \| \mb{x} \|_2}
     \end{align*}
     is below a desired tolerance, and if not, continue with 4.
-Here, $\mbh{m} \coloneqq [\hat{m}(t_1, \mb{p}_1), \dotsc, \hat{m}(t_{N_T}, \mb{p}_{N_P})]$ collects the mean predictions for all $(t, \mb{p}) \in T \times P$ and $\mb{x}$ contains the corresponding simulated solution values.
-    \item We compute the variance prediction $\hat{k}(t, \mb{p})$ using \eqref{eq:gpv} for all $(t, \mb{p}) \in (T \times P) \setminus D$ and add the point of maximum variance to the training data set $D$.
-    \item Finally, we retrain the model and continue with 3.
+Here, $\mbh{m} \coloneqq \big[ \hat{m}(t_1, \mb{p}_1), \dotsc, \hat{m}(t_{N_T}, \mb{p}_{N_P}) \big]$ collects the mean predictions for all $(t, \mb{p}) \in T \times P$ and $\mb{x}$ contains the corresponding simulated solution values.
+    \item We compute the variance prediction $\hat{k}(t, \mb{p})$ using \eqref{eq:ne_gpb} for all $(t, \mb{p}) \in (T \times P) \setminus D$ and add a point of maximum variance to the training data set $D$.
+    \item Finally, we retrain the GP and continue with 3.
 \end{enumerate}
@@
 \paragraph{First example circuit}
-We again consider the example of \autoref{fig:do1}, where we choose $R = \SI{500}{\ohm}$, $v_\mr{s}(t) = \sin(600 \pi t)\, \si{\volt}$ and the diode is modeled by $g_\mr{D}(\varphi_3) = 10^{-14} \big( e^{(\varphi_3 / \SI{26}{\mV})} - 1 \big)\, \si{\siemens}$.
-The remaining parameter values are chosen according to the greedy sample selection strategy with $\SI{1}{\milli\henry} \leq L \leq \SI{3}{\milli\henry}$ and $\SI{100}{\nano\farad} \leq C \leq \SI{300}{\nano\farad}$, i.e.~$\mb{p} = [L\ C]$ in the context of \autoref{sec:int}.
+\begin{figure}[t]
+    \begin{center}
+        \includegraphics[width=.4\textwidth]{do1_solution_phi} \hspace{1.5cm} \includegraphics[width=.4\textwidth]{do1_solution_i_L}
+    \end{center}
+    \caption{Solution of the first example circuit for $L = \SI{1.7}{\milli\henry}$ and $C = \SI{220}{\nano\farad}$.}
+    \label{fig:ne_do1_solution}
+\end{figure}
+We again consider the example of \autoref{fig:itc_do1}, where we choose $v_\mr{s}(t) = \sin(600 \pi t)\, \si{\volt}$, $R = \SI{500}{\ohm}$ and the diode is modeled by $g_\mr{D}(\varphi_3) = 10^{-14} \big( e^{(\varphi_3 / \SI{26}{\mV})} - 1 \big)\, \si{\siemens}$.
+The remaining parameter values are chosen according to the sequential sample selection strategy with $\SI{1}{\milli\henry} \leq L \leq \SI{3}{\milli\henry}$ and $\SI{100}{\nano\farad} \leq C \leq \SI{300}{\nano\farad}$, i.e.~$\mb{p} = [L, C]^\T$ in the context of \autoref{sec:int}.
 The starting points for the strategy are given by all combinations of the boundary values for $L$ and $C$ together with $t = \SI{0}{\ms}$ and $t = \SI{10}{\ms}$ once for each combination.
-Recalling \eqref{eq:do1}, we see that $\varphi_3$ and $i_\mr{L}$ have to be learned as the differential DOFs.
-In addition to these two, we also consider the algebraic variable $\varphi_2$ and we will compare the accuracy of learning $\varphi_2$ directly to recovering it from the two differential variables.
-\autoref{fig:do1_solution} shows the solution for $L = \SI{1.7}{\milli\henry}$ and $C = \SI{220}{\nano\farad}$, a combination which is not part of the training data.
-\begin{figure}[t]
-    \begin{center}
-        \includegraphics[width=.4\textwidth]{do1_solution_phi} \hspace{1.5cm} \includegraphics[width=.4\textwidth]{do1_solution_i_L}
-    \end{center}
-    \caption{Solution of the first example circuit for $L = \SI{1.7}{\milli\henry}$ and $C = \SI{220}{\nano\farad}$.}
-    \label{fig:do1_solution}
-\end{figure}
-
+\textcolor{red}{We note that the choice to consider $L$ and $C$, and not for example $R$, as parameters here is arbitrary and only serves to illustrate the approach.
+One could still follow the same approach when considering e.g.~the initial conditions, or the diode model, as being parameterized.} Recalling \eqref{eq:itc_do1}, we see that $\varphi_3$ and $i_\mr{L}$ have to be learned as the differential variables.
+In addition to these two, we also consider the algebraic variable $\varphi_2$ and we then compare the accuracy of learning $\varphi_2$ directly to recovering it from the two differential variables.
+\autoref{fig:ne_do1_solution} shows the solution for $L = \SI{1.7}{\milli\henry}$ and $C = \SI{220}{\nano\farad}$, a combination which is not part of the training data.
 \begin{figure}[b]
@@
     \end{center}
-    \caption{Convergence of the relative prediction error $e$, for the variables shown in \autoref{fig:do1_solution}.
+    \caption{Convergence of the relative prediction error $e$, for the variables shown in \autoref{fig:ne_do1_solution}.
 The final accuracies correspond to approximately 180 different combinations of $L$ and $C$ for $\varphi_3$, 39 combinations for $\varphi_2$ and 37 for $i_\mr{L}$.}
-    \label{fig:do1_convergence}
-\end{figure}
-In \autoref{fig:do1_convergence}, we see the convergence of the relative prediction errors with respect to the number of samples used by the greedy selection strategy.
+    \label{fig:ne_do1_convergence}
+\end{figure}
+
+In \autoref{fig:ne_do1_convergence}, we see the convergence of the relative prediction errors with respect to the number of samples used by the sample selection strategy.
 We observe that, depending on the qualitative complexity of the dynamics, the solution variables take different numbers of samples to reach the same prediction accuracy.
 It should be noted however, that the total number of samples $N_D$ also includes the number of time points that are sampled, and not only the number of distinct parameter combinations (simulations).
-The latter are listed in the caption of \autoref{fig:do1_convergence} and turn out to be considerably smaller.
+The latter are listed in the caption of \autoref{fig:ne_do1_convergence} and turn out to be considerably smaller.
 One may also note that we only execute the learning strategy up to a relatively large tolerance of $10^{-3}$.
@@
     \caption{Differences between the mean predictions ($\hat{\varphi}_3$ and $\hat{i}_\mr{L}$) and the corresponding simulations for the differential variables when considering $L = \SI{1.7}{\milli\henry}$ and $C = \SI{220}{\nano\farad}$.
-The predictions correspond to the smallest relative errors from \autoref{fig:do1_convergence}.}
-    \label{fig:do1_error_d}
+The predictions correspond to the smallest relative errors from \autoref{fig:ne_do1_convergence}.}
+    \label{fig:ne_do1_error_d}
 \end{figure}
@@
     \caption{Differences between the mean prediction $\hat{\varphi}_2$, reconstruction $\bar{\varphi}_2$ and the corresponding simulation of the algebraic variable $\varphi_2$ (left) and the respective consistency errors (right).
-The predictions correspond to $L = \SI{1.7}{\milli\henry}$, $C = \SI{220}{\nano\farad}$ and the smallest relative error from \autoref{fig:do1_convergence}.}
-    \label{fig:do1_error_a_consistency}
-\end{figure}
-The differences between the mean predictions and simulations for the differential DOFs, when using the predictions belonging to the smallest relative errors from \autoref{fig:do1_convergence}, are shown in \autoref{fig:do1_error_d}.
-The results again correspond to $L = \SI{1.7}{\milli\henry}$ and $C = \SI{220}{\nano\farad}$, and we observe that the differences are in line with the relative prediction errors of \autoref{fig:do1_convergence}.
-The differences between the mean prediction $\hat{\varphi}_2$, reconstruction $\bar{\varphi}_2$ and the simulation of the algebraic variable $\varphi_2$ are highlighted on the left of \autoref{fig:do1_error_a_consistency}.
+The predictions correspond to $L = \SI{1.7}{\milli\henry}$, $C = \SI{220}{\nano\farad}$ and the smallest relative error from \autoref{fig:ne_do1_convergence}.}
+    \label{fig:ne_do1_error_a_consistency}
+\end{figure}
+The differences between the mean predictions and simulations for the differential variables, when using the predictions belonging to the smallest relative errors from \autoref{fig:ne_do1_convergence}, are shown in \autoref{fig:ne_do1_error_d}.
+The results again correspond to $L = \SI{1.7}{\milli\henry}$ and $C = \SI{220}{\nano\farad}$, and we observe that the differences are in line with the relative prediction errors of \autoref{fig:ne_do1_convergence}.
+The differences between the mean prediction $\hat{\varphi}_2$, reconstruction $\bar{\varphi}_2$ and the simulation of the algebraic variable $\varphi_2$ are highlighted on the left of \autoref{fig:ne_do1_error_a_consistency}.
 The results again correspond to the same parameter values $L = \SI{1.7}{\milli\henry}$ and $C = \SI{220}{\nano\farad}$.
 We observe that there is not much difference between the mean prediction and reconstruction, however the reconstruction does appear to have a slight advantage in terms of accuracy.
-Here, one should note that although the reconstruction is exact up to the accuracy of solving the algebraic equation \eqref{eq:do1}, it still contains the error from learning the differential variables, hence the overall difference between $\bar{\varphi}_2$ and $\varphi_2$.
-To better quantify the difference between the learned and reconstructed solutions, we therefore introduce consistency errors $\hat{e}(t)$ and $\bar{e}(t)$ based on \eqref{eq:do1}
+Here, one should note that although the reconstruction is exact up to the accuracy of solving the algebraic equation in \eqref{eq:itc_do1}, it still contains the error from learning the differential variables, hence the overall difference between $\bar{\varphi}_2$ and $\varphi_2$.
+To better quantify the difference between the learned and reconstructed solutions, we introduce consistency errors $\hat{e}(t)$ and $\bar{e}(t)$ based on \eqref{eq:itc_do1}
 \begin{align*}
     \hat{e}(t) \coloneqq \begin{Vmatrix}
-        \hat{\varphi}_1 - v_\mr{s}\\
-        \hat{\varphi}_2 + R \hat{i}_\mr{L} - v_\mr{s}\\
-        \hat{i}_\mr{V} + i_\mr{L}
+        \hat{\varphi}_1 - v_\mr{s}(t)\\
+        \hat{\varphi}_2 + R \hat{i}_\mr{L} - v_\mr{s}(t)\\
+        \hat{i}_\mr{V} + \hat{i}_\mr{L}
     \end{Vmatrix}_2, \quad \bar{e}(t) \coloneqq \begin{Vmatrix}
-        \bar{\varphi}_1 - v_\mr{s}\\
-        \bar{\varphi}_2 + R \hat{i}_\mr{L} - v_\mr{s}\\
-        \bar{i}_\mr{V} + i_\mr{L}
+        \bar{\varphi}_1 - v_\mr{s}(t)\\
+        \bar{\varphi}_2 + R \hat{i}_\mr{L} - v_\mr{s}(t)\\
+        \bar{i}_\mr{V} + \hat{i}_\mr{L}
     \end{Vmatrix}_2,
 \end{align*}
-where $\hat{\cdot}$ indicates learned quantities and $\bar{\cdot}$ refers to reconstructed quantities.
+where $\hat{\cdot}$ indicates learned variables and $\bar{\cdot}$ refers to reconstructed variables.
 For a consistent solution obeying all algebraic constraints, both errors are identically zero.
-The right of \autoref{fig:do1_error_a_consistency} shows these consistency errors when using the same predictions and reconstructions as on the left.
+The right of \autoref{fig:ne_do1_error_a_consistency} shows these consistency errors when using the same predictions and reconstructions as on the left.
 We observe a very clear improvement of the reconstructions (consistency error on the order of machine precision) over the directly learned predictions (consistency error as large as $10^{-3}$).
-While the central benefit is of course this adherence to the constraints, only needing to learn two of the five solution variables also represents a significant reduction of the learning effort in this case.
+While the central benefit is this adherence to the constraints, only having to learn two of the five solution variables also represents a significant reduction of the learning effort in this case.
 
 \paragraph{Second example circuit}
-For the second example from \autoref{fig:do2}, we select $i_\mr{s}(t) = 10^{-4} \sin(400 \pi t)\, \si{\volt}$, while all other parameters and the learning strategy remain the same.
-Recalling \eqref{eq:do2}, we find that $\varphi_3$ is the only differential variable that needs to be learned to reconstruct the remaining algebraic DOFs.
+For the second example from \autoref{fig:itc_do2}, we select $i_\mr{s}(t) = 10^{-4} \sin(400 \pi t)\, \si{\volt}$, while all other parameters and the learning strategy remain the same.
+Recalling \eqref{eq:itc_do2}, we find that $\varphi_3$ is the only differential variable that needs to be learned to reconstruct the remaining algebraic variables.
 We again focus on $\varphi_2$ as an algebraic variable to obtain a comparative example.
+Considering similar numerical studies as for the first example, the left panel of \autoref{fig:ne_do2_solution_convergence} shows the solutions of $\varphi_3$ and $\varphi_2$ for $L = \SI{1.7}{\milli\henry}$ and $C = \SI{220}{\nano\farad}$, which are again not part of the training data.
+The right plot shows the convergence of the relative prediction errors with respect to the total number of samples $N_D$.
+We again note that the number of distinct parameter combinations is significantly smaller, as listed in the caption.
+We also see that the relative errors behave similar for both variables in this case, which is to be expected given the similar outlook of their dynamics in the left plot.
 \begin{figure}[b]
@@
 The final accuracies correspond to 31 distinct combinations of $L$ and $C$ for $\varphi_3$ and 49 for $\varphi_2$.}
-    \label{fig:do2_solution_convergence}
-\end{figure}
-
-Considering similar numerical studies as for the first example, the left panel of \autoref{fig:do2_solution_convergence} shows the solutions of $\varphi_3$ and $\varphi_2$ for $L = \SI{1.7}{\milli\henry}$ and $C = \SI{220}{\nano\farad}$, which are again not part of the training data.
-The right plot shows the convergence of the relative prediction errors with respect to the total number of samples $N_D$.
-We again note that the number of distinct parameter combinations is significantly smaller, as listed in the caption.
-We also see that the relative errors behave similarly for both variables in this case, which is to be expected given the similar outlook of their dynamics in the left plot.
+    \label{fig:ne_do2_solution_convergence}
+\end{figure}
+
+At this point we also return to the discussion from \autoref{sec:ial} about parameters only appearing in the algebraic equation.
+During the learning process, the sample selection strategy requested 21 unique values for $C$, all of which required full simulations according to \eqref{eq:itc_do2}.
+The 7 unique values for $L$, that were requested for learning $\varphi_2$, did not require full simulations however, but rather could be reconstructed from \eqref{eq:itc_do2}.
+\textcolor{red}{This results from the fact that the ODE in \eqref{eq:itc_do2} does not depend on $L$, such that the solution of $\varphi_3$ also does not depend on that parameter.
+The same goes for the algebraic variable $i_\mr{L}$, thus $\varphi_2$ and $\varphi_1$ can be reconstructed only based on the knowledge of $\varphi_3$.} We again emphasize that time is also included as a parameter within the sample selection strategy, such that the reconstructions of the algebraic variables only need to be evaluated at the particular points in time that are requested by the strategy, rather than at all time points of the solution.
 \begin{figure}[t]
@@
     \caption{Differences between the mean prediction and simulation for $\varphi_3$ (left) and between the mean prediction, reconstruction and simulation for $\varphi_2$ (right).
-The predictions are again made for $L = \SI{1.7}{\milli\henry}$ and $C = \SI{220}{\nano\farad}$ and correspond to the smallest relative errors from \autoref{fig:do2_solution_convergence}.}
-    \label{fig:do2_error}
-\end{figure}
-
-\begin{figure}[b]
-    \begin{center}
-        \includegraphics[width=.4\textwidth]{do2_consistency}
-    \end{center}
-    \caption{Consistency errors $\hat{e}(t)$ and $\bar{e}(t)$ corresponding to the results from \autoref{fig:do2_error}.
-The additional term $\check{e}(t)$ refers to the consistency error of the alternative approach, when using the predictions and Euler corrections outlined in \autoref{fig:do2_euler}.}
-    \label{fig:do2_consistency}
-\end{figure}
-At this point we also return to the discussion from \autoref{sec:ial} about parameters only appearing in the algebraic equation.
-During the learning process, the greedy strategy requested 21 unique values for $C$, all of which required full simulations according to \eqref{eq:do2}.
-The 7 unique values for $L$, that were requested for learning $\varphi_2$, did not require full simulations however, but rather could be reconstructed from \eqref{eq:do2}.
-We again emphasize that time is also included as a parameter within the sample selection strategy, such that the reconstruction not only constitutes a reduction from solving a nonlinear DAE in four variables to merely evaluating a scalar equation, but that this scalar equation furthermore only needs to be evaluated at the particular points in time that are requested by the strategy, rather than at all time points of the solution.
-
-The difference between the mean prediction and simulation of $\varphi_3$, for $L = \SI{1.7}{\milli\henry}$ and $C = \SI{220}{\nano\farad}$, is shown on the left of \autoref{fig:do2_error}, while the right shows the differences between the mean prediction, reconstruction and simulation of $\varphi_2$.
-The predictions again correspond to the smallest relative errors from \autoref{fig:do2_solution_convergence} and the differences are of the same order.
+The predictions are again made for $L = \SI{1.7}{\milli\henry}$ and $C = \SI{220}{\nano\farad}$ and correspond to the smallest relative errors from \autoref{fig:ne_do2_solution_convergence}.}
+    \label{fig:ne_do2_error}
+\end{figure}
+
+The difference between the mean prediction and simulation of $\varphi_3$, for $L = \SI{1.7}{\milli\henry}$ and $C = \SI{220}{\nano\farad}$, is shown on the left of \autoref{fig:ne_do2_error}, while the right shows the differences between the mean prediction, reconstruction and simulation of $\varphi_2$.
+The predictions again correspond to the smallest relative errors from \autoref{fig:ne_do2_solution_convergence}, and the differences are of the same order.
 In this case, the reconstruction performs similar or even slightly worse compared to the mean prediction when only looking at the difference.
-However it still constitutes a reduction in learning effort, from four solution variables down to only one, and most importantly it adheres to the algebraic constraints of the DAE.
-To further illustrate this point, we again take a look at the consistency errors, now redefined based on \eqref{eq:do2}
+However it still constitutes a reduction in learning effort, from four solution variables down to only one, and most importantly the reconstruction adheres to the algebraic constraints of the DAE.
+To further illustrate this point, we again take a look at the consistency errors, now redefined based on \eqref{eq:itc_do2}
 \begin{align*}
     \hat{e}(t) \coloneqq \begin{Vmatrix}
-        \hat{\varphi}_1 - \hat{\varphi}_2 + R i_\mr{s}\\
-        \hat{\varphi}_2 - \hat{\varphi}_3 + L \ddt i_\mr{s}\\
-        \hat{i}_\mr{L} - i_\mr{s}
+        \hat{\varphi}_1 - \hat{\varphi}_2 + R i_\mr{s}(t)\\
+        \hat{\varphi}_2 - \hat{\varphi}_3 + L \ddt i_\mr{s}(t)\\
+        \hat{i}_\mr{L} - i_\mr{s}(t)
     \end{Vmatrix}_2, \quad \bar{e}(t) \coloneqq \begin{Vmatrix}
-        \bar{\varphi}_1 - \bar{\varphi}_2 + R i_\mr{s}\\
-        \bar{\varphi}_2 - \hat{\varphi}_3 + L \ddt i_\mr{s}\\
-        \bar{i}_\mr{L} - i_\mr{s}
+        \bar{\varphi}_1 - \bar{\varphi}_2 + R i_\mr{s}(t)\\
+        \bar{\varphi}_2 - \hat{\varphi}_3 + L \ddt i_\mr{s}(t)\\
+        \bar{i}_\mr{L} - i_\mr{s}(t)
     \end{Vmatrix}_2.
 \end{align*}
-In our example, the algebraic equation from \eqref{eq:do2} gives an explicit description of the algebraic variables once more, such that the reconstruction is accurate up to machine precision, compare $\bar{e}(t)$ in \autoref{fig:do2_consistency}.
+In our example, the algebraic equation from \eqref{eq:itc_do2} gives an explicit description of the algebraic variables once more, such that the reconstruction is accurate up to machine precision, compare $\bar{e}(t)$ in \autoref{fig:ne_do2_consistency}.
 When learning all solution variables individually however, we observe a much larger maximum value of around $10^{-3}$ for the consistency error $\hat{e}(t)$ across all the predicted points in time.
 In general, the algebraic equation may be nonlinear such that the reconstruction still leads to a consistency error $\bar{e}(t)$ greater than machine precision, depending on the accuracy of the nonlinear system solver.
-
-\subsection{Alternative approach}
-\label{subsec:aa}
-In a last step, we want to provide an alternative approach to reducing the consistency error, while avoiding an extra implementation to handle the algebraic constraints.
-The idea is based on the insight that two successive implicit Euler steps suffice to bring the algebraic variables into a consistent state with the differential ones, when starting with inconsistent initial values\cite{garcia2022}.
-In our context, one can then simply learn all solution variables independently, and when making a prediction, select a time point slightly before the intended prediction point, such that two small implicit Euler steps may be performed as a correction to obtain a consistent solution.
+\begin{figure}[b]
+    \begin{center}
+        \includegraphics[width=.4\textwidth]{do2_consistency}
+    \end{center}
+    \caption{Consistency errors $\hat{e}(t)$ and $\bar{e}(t)$ corresponding to the results from \autoref{fig:ne_do2_error}.}
+    \label{fig:ne_do2_consistency}
+\end{figure}
+
+\paragraph{Rectifier circuit}
 \begin{figure}[t]
     \begin{center}
-        \includegraphics[width=.4\textwidth]{do2_euler_phi_3} \hspace{1.5cm} \includegraphics[width=.4\textwidth]{do2_euler_phi_2}
-    \end{center}
-    \caption{Differences between the mean predictions, Euler corrections and simulations for $\varphi_3$ (left) and $\varphi_2$ (right).
-The predictions are made for $L = \SI{1.7}{\milli\henry}$ and $C = \SI{220}{\nano\farad}$ and correspond to the smallest relative errors from \autoref{fig:do2_solution_convergence}.}
-    \label{fig:do2_euler}
-\end{figure}
-
-This alternative approach has the advantage that it can be directly applied using simulations from any SPICE-like software, and using any machine learning method for the predictions, without worrying about the reconstruction of the algebraic variables.
-However it may not achieve the same level of accuracy as the approach outlined in \autoref{sec:ial} in practice, depending on the Euler step size and the accuracy of the nonlinear system solver\cite{garcia2022}.
-Furthermore, it requires predictions for all solution variables, which represents unnecessary computational cost, and there is no chance for parameters to shift completely to the algebraic equations, compare \autoref{sec:ial}.
-
-\autoref{fig:do2_euler} shows the differences between the mean predictions ($\hat{\varphi}_3$, $\hat{\varphi}_2$), Euler corrections ($\check{\varphi}_3$, $\check{\varphi}_2$) and the simulations for the second example, using the same parameters as in \autoref{fig:do2_error}.
-We see that the two Euler steps have almost no effect on the differential variable $\varphi_3$, as expected.
-For the algebraic variable $\varphi_2$ however, the mean predictions $\hat{\varphi}_2$ and the corrected predictions after two Euler steps $\check{\varphi}_2$, differ significantly, with the latter resembling the reconstruction from \autoref{fig:do2_error}.
-Also defining a consistency error $\check{e}(t)$ based on \eqref{eq:do2} in this case
-\begin{align*}
-    \check{e}(t) \coloneqq \begin{Vmatrix}
-        \check{\varphi}_1 - \check{\varphi}_2 + R i_\mr{s}\\
-        \check{\varphi}_2 - \check{\varphi}_3 + L \ddt i_\mr{s}\\
-        \check{i}_\mr{L} - i_\mr{s}
-    \end{Vmatrix}_2
-\end{align*}
-and comparing it to the other approaches, see \autoref{fig:do2_consistency}, shows that this alternative approach represents a compromise in terms of accuracy in the algebraic constraints.
-The maximum value of the consistency error $\check{e}(t)$ of the corrected predictions comes out around $10^{-7}$, about four orders of magnitude smaller compared to before the correction.
-While it is theoretically possible to also force the error down to machine precision in this case, other factors become prohibitive as mentioned earlier.
+        \includegraphics[width=0.5\textwidth]{fwr2}
+    \end{center}
+    \caption{Full wave rectifier circuit.}
+    \label{fig:ne_fwr2}
+\end{figure}
+
+\begin{figure}[b]
+    \begin{center}
+        \includegraphics[width=.4\textwidth]{fwr2_solution_v34} \hspace{1.5cm} \includegraphics[width=.4\textwidth]{fwr2_solution_i_L_1}
+    \end{center}
+    \caption{Solutions of the differential variables of the full wave rectifier circuit for $T_1 = 65\, \si{\celsius}$ and $T_2 = 85\, \si{\celsius}$.}
+    \label{fig:ne_fwr2_solution_d}
+\end{figure}
+\textcolor{red}{As a third and larger example we consider the rectifier circuit from \autoref{fig:ne_fwr2}.
+Aside from requiring the full index two implementation from \autoref{sec:ial}, this example also showcases another potential application in the context of electrothermal simulations.
+Electrothermal simulations are used to investigate the thermal behavior of a circuit, by coupling the power that is dissipated in the circuit to a set of equations describing the temperature distribution in the circuit, and by using the temperatures of some components as parameters for certain device functions.
+In our case, we again model the diodes as nonlinear resistors, however this time the model also includes a temperature dependence\cite{tietze2008}
+\begin{align*}
+    g_\mr{D}(v_\mr{D}, T_\mr{D}) = I_\mr{s}(T_\mr{D}) \frac{qk}{T_\mr{D}} e^{(qk/T_\mr{D}) v_\mr{D}},
+\end{align*}
+where $v_\mr{D}$ is the voltage across the diode in forward direction, $T_\mr{D}$ is the temperature of the diode, $q$ is the charge of an electron, $k$ the Boltzmann constant and $I_\mr{s}(T_\mr{D})$ the temperature dependent reverse saturation current\cite{tietze2008}.
+The current source is given by $i_\mr{s}(t) = 0.1 \cos(100 \pi t)\, \si{\ampere}$ and the capacitor and resistor are linear with $C = 1\, \si{\milli\farad}$, $R = 50\, \si{\ohm}$.
+The circuit also contains a transformer $\mr{T}_{12}$ modeled as a nonlinear inductive multiport with
+\begin{align*}
+    \mb{L}(\i{L}) = \begin{bmatrix}
+        L(i_{\mr{L}_1}) & \sqrt{0.09 L(i_{\mr{L}_1}) L(i_{\mr{L}_2})}\\
+        \sqrt{0.09 L(i_{\mr{L}_1}) L(i_{\mr{L}_2})} & 0.1 L(i_{\mr{L}_2})
+    \end{bmatrix},
+\end{align*}
+where $L(i_\mr{L})$ is a fourth order polynomial in $i_\mr{L}$\cite{lullo2017}.
+We also note that rectifiers play an important role in many power electronics applications; they are key when converting high AC to lower DC voltages.
+In the example we still use a current source, since this leads to an index two circuit in this case, compare \autoref{th:ic_imna}, which requires the full index two approach from \autoref{sec:ial}.}
+\begin{figure}[t]
+    \begin{center}
+        \includegraphics[width=.4\textwidth]{fwr2_solution_phi} \hspace{1.5cm} \includegraphics[width=.4\textwidth]{fwr2_solution_i_L_2}
+    \end{center}
+    \caption{Solutions of the algebraic variables of the full wave rectifier circuit for $T_1 = 65\, \si{\celsius}$ and $T_2 = 85\, \si{\celsius}$.}
+    \label{fig:ne_fwr2_solution_a}
+\end{figure}
+
+\begin{figure}[b]
+    \begin{center}
+        \includegraphics[width=.4\textwidth]{fwr2_error_phi} \hspace{1.5cm} \includegraphics[width=.4\textwidth]{fwr2_error_i_L}
+    \end{center}
+    \caption{Differences between the mean predictions ($\hat{\cdot}$) and simulations of all differential and algebraic variables for $T_1 = 65\, \si{\celsius}$ and $T_2 = 85\, \si{\celsius}$.}
+    \label{fig:ne_fwr2_error}
+\end{figure}
+\textcolor{red}{When decoupling the DAE arising from the circuit of \autoref{fig:ne_fwr2} using the dissection index, one finds
+\begin{align*}
+    \mbt{x}_\mb{Q} = \begin{bmatrix}
+        v_{34}\\
+        i_{\mr{L}_2}
+    \end{bmatrix}, \quad \mbt{x}_\mb{P} = i_{\mr{L}_1}, \quad \mbb{x}_\mb{P} = \begin{bmatrix}
+        \varphi_2\\
+        \varphi_4
+    \end{bmatrix}, \quad \mbb{x}_\mb{Q} = \varphi_1
+\end{align*}
+for the differential and algebraic variables using the notation from \autoref{sec:di}.
+We first observe that there are only two differential variables $\mbt{x}_\mb{Q}$, again leading to a significant reduction in the learning effort.
+We also see that the first differential variable $v_{34} \coloneqq \varphi_3 - \varphi_4$ is a linear combination of the original variables of the DAE.
+This is a general phenomenon and does not interfere with our approach, as the original variables may always be reconstructed from the sets of differential and algebraic variables by reversing the splitting using \eqref{eq:ioc_x}, \eqref{eq:itc_xb} and \eqref{eq:itc_xt} from \autoref{sec:di}
+\begin{align*}
+    \mb{x} = \mb{P} \mbt{x} + \mb{Q} \mbb{x} = \mb{P} \big( \mbt{P} \mbt{x}_\mb{P} + \mbt{Q} \mbt{x}_\mb{Q} \big) + \mb{Q} \big( \mbb{P} \mbb{x}_\mb{P} + \mbb{Q} \mbb{x}_\mb{Q} \big).
+\end{align*}
+To keep the overall learning effort manageable, we consider the outer diodes $\mr{D}_1$ and $\mr{D}_4$ to depend on the same temperature $T_1$ and the inner diodes $\mr{D}_2$ and $\mr{D}_3$ to depend on the same temperature $T_2$, such that the vector of parameters is now given by $\mb{p} = [T_1, T_2]^\T$.
+We emphasize that there is no need to restrict the vector of parameters to two entries, this choice is only made to reduce the learning effort.
+For the training data we consider temperatures $\SI{20}{\celsius} \leq T_1, T_2 \leq \SI{90}{\celsius}$, and the starting values for the sample selection strategy are again obtained by considering all combinations of the boundary values together with $t = \SI{0}{\ms}$ and $t = \SI{50}{\ms}$.
+We then execute the learning strategy up to a tolerance of $5 \cdot 10^{-3}$ and choose the prediction point $T_1 = \SI{65}{\celsius}$, $T_2 = \SI{85}{\celsius}$ that is not included in the training data.}
+
+\textcolor{red}{The solutions of the differential and algebraic variables, for $T_1 = 65\, \si{\celsius}$ and $T_2 = 85\, \si{\celsius}$, are shown in \autoref{fig:ne_fwr2_solution_d} and \autoref{fig:ne_fwr2_solution_a}, while \autoref{fig:ne_fwr2_error} shows the differences between the mean predictions and simulations for the same temperature values.
+The comparatively larger differences for $i_{\mr{L}_1}$ in \autoref{fig:ne_fwr2_error} stem from the prediction error being defined as a relative quantity and the absolute values of $i_{\mr{L}_1}$ being much larger than the others, compare \autoref{fig:ne_fwr2_solution_d}.
+Defining a function $\mb{g}(\mbt{x}_\mb{Q}, \mbt{x}_\mb{P}, \mbb{x}_\mb{P}, \mbb{x}_\mb{Q}, t, \mb{p})$ by stacking the nonlinear systems \eqref{eq:itc_xtp}, \eqref{eq:itc_xbp} and \eqref{eq:itc_itiiib} required to recover the algebraic variables, compare \autoref{sec:ial}, we can again define consistency errors
+\begin{align*}
+    \hat{e}(t) \coloneqq \big\| \mb{g}(\hat{\mbt{x}}_\mb{Q}, \hat{\mbt{x}}_\mb{P}, \hat{\mbb{x}}_\mb{P}, \hat{\mbb{x}}_\mb{Q}, t, \mb{p}) \big\|_2, \quad \bar{e}(t) \coloneqq \big\| \mb{g}(\hat{\mbt{x}}_\mb{Q}, \bar{\mbt{x}}_\mb{P}, \bar{\mbb{x}}_\mb{P}, \bar{\mbb{x}}_\mb{Q}, t, \mb{p}) \big\|_2,
+\end{align*}
+where the additional $\hat{\cdot}$ and $\bar{\cdot}$ again refer to learned or reconstructed variables respectively.
+For $T_1 = 65\, \si{\celsius}$ and $T_2 = 85\, \si{\celsius}$ the consistency errors can be seen in \autoref{fig:ne_fwr2_consistency}.
+We observe that the reconstructed solution again outperforms the directly learned solution by several orders of magnitude.}
+\begin{figure}
+    \begin{center}
+        \includegraphics[width=.4\textwidth]{fwr2_consistency}
+    \end{center}
+    \caption{Consistency errors $\hat{e}(t)$ and $\bar{e}(t)$ corresponding to the results from \autoref{fig:ne_fwr2_error}}
+    \label{fig:ne_fwr2_consistency}
+\end{figure}
 
@@
 \label{sec:cfr}
-This article introduced a new method for learning the time and parameter dependent behavior of electrical circuits.
-The method assumes the circuit to be modeled using MNA and then exploits the structure of MNA to improve the learning process by splitting it into a learning and a reconstruction stage, compare \autoref{fig:ial}.
+This article introduced a new approach for learning the time and parameter dependent solutions of electrical circuits.
+The approach assumes the circuit to be modeled using MNA, and then exploits the structure of MNA to improve the learning process by splitting it into a learning and a reconstruction step, compare \autoref{fig:ial_ial}.
 It achieves this by decoupling the underlying DAE into an ODE and a purely algebraic equation.
-Benefits of the method include a reduction in the number of variables that are to be learned during the learning stage and the exact adherence of the learned solution to the inherent constraints of the circuit model after the reconstruction stage.
-Two numerical examples illustrated both benefits.
+Benefits of the approach include a reduction in the number of variables that are to be learned during the learning step, and the exact adherence of the learned solution to the inherent constraints of the circuit model after the reconstruction step.
+Numerical examples illustrated both benefits.
 The examples also showed that the exact recovery of the constraints might improve the accuracy of the learned solutions.
-Furthermore, additional computational savings may be possible as some of the parameters of interest only appear in the reconstruction stage, which entirely avoids the need for training data with varying values of these parameters.
-We emphasize that the approach is independent of the machine learning method used during the learning stage, such that the learning method may be chosen according to the problem at hand.
-Lastly, we showcased an alternative approach that avoids the need for the decoupling and the associated implementation, while still improving the adherence of the learned solution to the algebraic constraints of the DAE.
+Furthermore, additional computational savings may be possible as some of the parameters of interest only appear in the reconstruction step, which avoids the need for training data with varying values of these parameters entirely.
+We emphasize that the approach is independent of the machine learning method used during the learning step, such that the learning method may be chosen according to the problem at hand.
 
 Multiple extensions are possible within this index-aware learning framework.
-A natural first step could be to make use of the topological decomposition that was hinted at in \autoref{sec:dc}.
-This would pave the way for the inclusion of controlled sources within the workflow, and thus the industrial use of the approach, with the idea being the extension of the topological decomposition to also allow for controlled sources.
-Focusing in on the idea of adhering to physically meaningful constraints, one could work on extending the concept to a charge conserving variant of MNA, to potentially guarantee charge conservation even for the learned solutions.
+A natural first step could be to make use of the topological decoupling that was hinted at in \autoref{sec:di}.
+This would pave the way for the inclusion of controlled sources within the workflow, and thus the industrial use of the approach, with the idea being the extension of the topological decoupling to also allow for controlled sources.
+Focusing in on the idea of adhering to physically meaningful constraints, one could work on extending the approach to a charge conserving variant of MNA, to potentially guarantee charge conservation even for the learned solutions.
 Yet another direction may be the application of index-aware learning to DAEs arising elsewhere, e.g.~in modified loop analysis.
-(The dissection concept applies to a more general class of DAEs, as \autoref{sec:dc} showed.) Finally, separate but related work may focus on improving the learning stage by developing new methods that are especially suited for learning ODEs.
+(The dissection index applies to a more general class of DAEs, as \autoref{sec:di} showed.) Finally, separate but related work may focus on improving the learning step by developing new methods that are especially suited for learning ODEs.
 
@@
 
-% \appendix
-% \section{}
-
-% \nocite{*}
-\bibliography{biblio}
+\appendix
+\section{Dissection index and MNA}
+\label{sec:dim}
+\begin{remark}
+    \label{re:dim_bf}
+    Recalling \eqref{eq:md_mnaii}, we observe
+    \begin{align*}
+        \mb{M}(\mb{x}) = \begin{bmatrix}
+            \A{C}^{\phantom{\T}} \mb{C} \big( \AT{C} \vphi \big) \AT{C} & &\\
+            & \mb{L}(\i{L}) &\\
+            & & \mb{0}
+        \end{bmatrix}
+    \end{align*}
+    in the case of MNA.
+We now determine the basis functions $\mb{Q}$ and $\mb{P}$.
+Taking into account the standard assumption, compare \autoref{th:ic_imna}, that $\mb{L}(\i{L})$ and $\mb{C} \big( \AT{C} \vphi \big)$ are positive definite, we find\cite{jansen2014}
+    \begin{align*}
+        \mb{Q} = \begin{bmatrix}
+            \mb{Q}_\mr{C} &\\
+            & \mb{0}\\
+            & \mb{I}
+        \end{bmatrix}, \quad \mb{P} = \begin{bmatrix}
+            \mb{P}_\mr{C} &\\
+            & \mb{I}\\
+            & \mb{0}
+        \end{bmatrix},
+    \end{align*}
+    where $\mr{im\, } \mb{Q}_\mr{C} = \ker \AT{C}$, $\mb{0}$ and $\mb{I}$ are zero and identity matrices of appropriate dimensions and the columns of $\mb{Q}_\mr{C}$ and $\mb{P}_\mr{C}$ together form a basis of $\mathbb{R}^{n_{\vphi}}$ with $n_{\vphi}$ the dimension of $\vphi$.
+The description of $\mb{Q}_\mr{C}$ results from the fact that (with $\mb{y} \coloneqq \AT{C} \mb{x}$)
+    \begin{align*}
+        \mb{x}^\T \A{C}^{\phantom{\T}} \mb{C} \big( \AT{C} \vphi \big) \AT{C} \mb{x} = \mb{y}^\T \mb{C} \big( \AT{C} \vphi \big) \mb{y} > 0, \quad \forall \mb{y} \in \mathbb{R}^{n_{\vphi}} \setminus \{ \mb{0} \},
+    \end{align*}
+    since $\mb{C} \big( \AT{C} \vphi \big)$ is positive definite and hence the kernel of $\A{C}^{\phantom{\T}} \mb{C} \big( \AT{C} \vphi \big) \AT{C}$ is determined by the kernel of $\AT{C}$.
+As $\A{C}$ is an incidence (constant) matrix, we find that $\mb{Q}$ and $\mb{P}$ are also constant.
+\end{remark}
+
+\textcolor{red}{\begin{remark}
+    \label{re:dim_xtp}
+    Again recalling \eqref{eq:md_mnaii} we observe
+    \begin{align*}
+        \mb{K}(\mb{x}) = \begin{bmatrix}
+            \A{R}^{\phantom{\T}} \mb{G} \big( \AT{R} \vphi \big) \AT{R} & \A{L} & \A{V}\\
+            -\AT{L} & &\\
+            -\AT{V} & &
+        \end{bmatrix}.
+    \end{align*}
+    Noting that $\mb{W} = \mb{Q}$ for MNA, due to the symmetry of $\mb{M}(\mb{x})$ in \autoref{re:dim_bf}, we find
+    \begin{align*}
+        \mbb{K}_\mb{Q}(\mb{x}) = \mb{Q}^\T \mb{K}(\mb{x}) \mb{Q} = \begin{bmatrix}
+            \mb{Q}_\mr{C}^\T \A{R}^{\phantom{\T}} \mb{G} \big( \AT{R} \vphi \big) \AT{R} \mb{Q}_\mr{C} & \mb{Q}_\mr{C}^\T \A{V}\\
+            -\AT{V} \mb{Q}_\mr{C} &
+        \end{bmatrix}, \quad \mbb{K}_\mb{P}(\mb{x}) = \mb{Q}^\T \mb{K}(\mb{x}) \mb{P} = \begin{bmatrix}
+            \mb{Q}_\mr{C}^\T \A{R}^{\phantom{\T}} \mb{G} \big( \AT{R} \vphi \big) \AT{R} \mb{P}_\mr{C} & \mb{Q}_\mr{C}^\T \A{L}\\
+            -\AT{V} \mb{P}_\mr{C} &
+        \end{bmatrix}.
+    \end{align*}
+    Determining $\mbb{W}(\mb{x})$ using $\mbb{K}_\mb{Q}(\mb{x})$ gives\cite{jansen2014}
+    \begin{align*}
+        \mbb{W} = \begin{bmatrix}
+            \mb{Q}_\mr{V} \mb{Q}_\mr{R} &\\
+            & \mb{W}_\mr{V}
+        \end{bmatrix},
+    \end{align*}
+    where $\mr{im\, } \mb{Q}_\mr{V} = \ker \AT{V} \mb{Q}_\mr{C}$, $\mr{im\, } \mb{Q}_\mr{R} = \ker \AT{R} \mb{Q}_\mr{C} \mb{Q}_\mr{V}$ and $\mr{im\, } \mb{W}_\mr{V} = \ker \AT{V} \mb{Q}_\mr{C}$ are all constant such that $\mbb{W}$ is constant as well.
+This finally yields
+    \begin{align*}
+        \mbb{W}^\T \mbb{K}_\mb{P}(\mb{x}) = \begin{bmatrix}
+            \mb{Q}_\mr{R}^\T \mb{Q}_\mr{V}^\T \mb{Q}_\mr{C}^\T \A{R}^{\phantom{\T}} \mb{G} \big( \AT{R} \vphi \big) \AT{R} \mb{P}_\mr{C} & \mb{Q}_\mr{R}^\T \mb{Q}_\mr{V}^\T \mb{Q}_\mr{C}^\T \A{L}\\
+            -\mb{W}_\mr{V}^\T \AT{V} \mb{P}_\mr{C} &
+        \end{bmatrix} = \begin{bmatrix}
+            \mb{0} & \mb{Q}_\mr{R}^\T \mb{Q}_\mr{V}^\T \mb{Q}_\mr{C}^\T \A{L}\\
+            -\mb{W}_\mr{V}^\T \AT{V} \mb{P}_\mr{C} &
+        \end{bmatrix},
+    \end{align*}
+    thus the basis functions $\mbt{P}$ and $\mbt{Q}$ of $\mbb{W}^\T \mbb{K}_\mb{P}(\mb{x})$ are constant as well, and since $\mbb{W}^\T \mbb{f}(t)$ only depends on $t$, we find a unique solution for $\mbt{x}_\mb{P}$ that also only depends on $t$.
+\end{remark}}
+
+\bibliography{../../../ai-twilight/biblio}
+% \bibliography{biblio}
+
+% \input{../fig/vcr}
+% \input{../fig/do1}
+% \input{../fig/do2}
+% \input{../fig/ial}
+% \input{../fig/solution}
+% \input{../fig/convergence}
+% \input{../fig/full_wave_rectifier}
 

### 2024-01-22 doc/article/letter.tex
revision

\documentclass[a4paper, 10pt,
    ]{article}

\usepackage[top=1.75cm, bottom=1.75cm, left=3cm, right=3cm]{geometry}
\usepackage[ngerman]{babel}
\usepackage[T1]{fontenc}

% font
\usepackage[scaled=.98,sups,osf]{XCharter}
\usepackage[scaled=1.04,varqu,varl]{inconsolata}
\usepackage[type1]{cabin}
\usepackage[charter,vvarbb,scaled=1.05]{newtxmath}
\usepackage[cal=boondoxo]{mathalfa}
\linespread{1.04}

\usepackage{tuda-pgfplots}

\pagestyle{empty}
\setlength{\parindent}{0cm}

\newcommand{\mb}[1]{\mathbf{#1}}
\newcommand{\mbt}[1]{\tilde{\mathbf{#1}}}
\newcommand{\mbb}[1]{\bar{\mathbf{#1}}}
\newcommand{\mbh}[1]{\hat{\mathbf{#1}}}
\newcommand{\mr}[1]{\mathrm{#1}}
\newcommand{\T}{{\!\top}}
\newcommand{\ddt}{\frac{\mathrm{d}}{\mathrm{d}t}}
\newcommand{\A}[1]{\mb{A}_\mr{#1}}
\newcommand{\AT}[1]{\mb{A}_\mr{#1}^{\T}}
\newcommand{\qC}{\mb{q}_\mr{C}}
\newcommand{\gR}{\mb{g}_\mr{R}}
\newcommand{\phiL}{\boldsymbol{\phi}_\mr{L}}
\newcommand{\vphi}{\boldsymbol{\varphi}}
\renewcommand{\i}[1]{\mb{i}_\mr{#1}}
\renewcommand{\v}[1]{\mb{v}_\mr{#1}}

\begin{document}
    Dear editors of the International Journal of Circuit Theory and Applications,\\

    thank you for the feedback on our manuscript ``Index-aware learning of circuits". We have reviewed the comments (\textcolor{TUDa-0c}{repeated in gray}) and made changes (\textcolor{red}{highlighted in red}) in order to address the criticism and improve the paper.\\

    \paragraph{Reviewing: 1}
    \begin{itemize}
        \item[1)] \textcolor{TUDa-0c}{Both examples are well selected from a didactic point of view. They contain constant $C$ and $L$ and present a rather trivial DAE-structure, since the basis considered in the dissection concept consist of standard unit vectors $e_i$. However, they do clarify all challenges that may appear in general. It should be explained that due to the properties of the incidence matrices, in general the ODEs have to be formulated for linear combinations of components. Moreover, for nonlinear $C$ and $G$ corresponding to multiport devices, it should be discussed if such a decoupling is still possible. On the one hand the theory should be more precise and on the other hand at least one additional larger nonlinear index-2 example with a more complex structure should be included, e.g.~a CV-loop involving several nonlinear capacitors.}
        \begin{itemize}
            \item We added a paragraph in section 2 to illustrate that the decoupling also works for nonlinear multiport devices.\\
            ``\textcolor{red}{We note that \dots\ readily extends to multiport devices. A general inductive $n$-port for example, can also be modeled by a function $\phiL(\i{L}) = \big[ \phi_1(\i{L}), \dotsc, \phi_{n-1}(\i{L}) \big]^\T$, however the Jacobian is not necessarily diagonal in this case as the component functions of $\phiL$ can each depend on all the currents $\i{L} = [i_1, \cdots, i_{n-1}]^\T$. Regarding the incidence matrices, one chooses a reference node $\varphi_0$ for the multiport device and then considers $n-1$ branches from the remaining $n-1$ nodes to the reference. Interpreting the multiport as a single node and using Kirchhoff's current law then gives $i_0 = -\sum_{k=1}^{n-1} i_k$ for the reference current, when orienting all currents to point toward the device. Following this approach, there is then no difference in treating multiports compared to one-ports.}'' (p.2)
            \item We made the mathematical presentation of the dissection index more precise, reformulating section 3 in terms of nonlinear DAEs in standard form while providing the assumptions that are necessary for our approach. We also added an appendix with remarks showing that MNA fulfills all assumptions.\\
            ``\textcolor{red}{We consider a DAE in standard form,
            \begin{align}
                \mb{M}(\mb{x}) \ddt \mb{x} + \mb{K}(\mb{x}) \mb{x} + \mb{f}(t) = \mb{0}, \quad \ker \mb{M} \supsetneq \{ \mb{0} \},
            \end{align}
            where the matrix-valued functions $\mb{M}$ and $\mb{K}$ are derived from \dots\ by defining
            \begin{align*}
                \mb{M}(\mb{x}) = \mb{J}_{\mb{x}'} \big( \mb{F}(\mb{x}', \mb{x}, t) \big), \quad \mb{K}(\mb{x}) = \mb{J}_{\mb{x}} \big( \mb{F}(\mb{x}', \mb{x}, t) \big).
            \end{align*}
            Note that the description of MNA in \dots\ is precisely of this form. In the following, we will demonstrate the first two steps of the dissection index when applied to systems of the form of \dots, while stating the assumptions of our approach. Appendix \dots\ contains additional remarks showing that these assumptions are fulfilled by \dots.}'' (p.4)\\
            See also Assumption 1 (p.4), Assumption 2 and Assumption 3 (p.5) as well as Remark 2 (p.19).\\
            ``\textcolor{red}{Again recalling \dots\ we observe
            \begin{align*}
                \mb{K}(\mb{x}) = \begin{bmatrix}
                    \A{R}^{\phantom{\T}} \mb{G} \big( \AT{R} \vphi \big) \AT{R} & \A{L} & \A{V}\\
                    -\AT{L} & &\\
                    -\AT{V} & &
                \end{bmatrix}.
            \end{align*}
            Noting that $\mb{W} = \mb{Q}$ for MNA, due to the symmetry of $\mb{M}(\mb{x})$ in \dots, we find
            \begin{align*}
                \mbb{K}_\mb{Q}(\mb{x}) &= \mb{Q}^\T \mb{K}(\mb{x}) \mb{Q} = \begin{bmatrix}
                    \mb{Q}_\mr{C}^\T \A{R}^{\phantom{\T}} \mb{G} \big( \AT{R} \vphi \big) \AT{R} \mb{Q}_\mr{C} & \mb{Q}_\mr{C}^\T \A{V}\\
                    -\AT{V} \mb{Q}_\mr{C} &
                \end{bmatrix}\\
                \mbb{K}_\mb{P}(\mb{x}) &= \mb{Q}^\T \mb{K}(\mb{x}) \mb{P} = \begin{bmatrix}
                    \mb{Q}_\mr{C}^\T \A{R}^{\phantom{\T}} \mb{G} \big( \AT{R} \vphi \big) \AT{R} \mb{P}_\mr{C} & \mb{Q}_\mr{C}^\T \A{L}\\
                    -\AT{V} \mb{P}_\mr{C} &
                \end{bmatrix}.
            \end{align*}
            Determining $\mbb{W}(\mb{x})$ using $\mbb{K}_\mb{Q}(\mb{x})$ gives
            \begin{align*}
                \mbb{W} = \begin{bmatrix}
                    \mb{Q}_\mr{V} \mb{Q}_\mr{R} &\\
                    & \mb{W}_\mr{V}
                \end{bmatrix},
            \end{align*}
            where $\mr{im\, } \mb{Q}_\mr{V} = \ker \AT{V} \mb{Q}_\mr{C}$, $\mr{im\, } \mb{Q}_\mr{R} = \ker \AT{R} \mb{Q}_\mr{C} \mb{Q}_\mr{V}$ and $\mr{im\, } \mb{W}_\mr{V} = \ker \AT{V} \mb{Q}_\mr{C}$ are all constant such that $\mbb{W}$ is constant as well. This finally yields
            \begin{align*}
                \mbb{W}^\T \mbb{K}_\mb{P}(\mb{x}) = \begin{bmatrix}
                    \mb{Q}_\mr{R}^\T \mb{Q}_\mr{V}^\T \mb{Q}_\mr{C}^\T \A{R}^{\phantom{\T}} \mb{G} \big( \AT{R} \vphi \big) \AT{R} \mb{P}_\mr{C} & \mb{Q}_\mr{R}^\T \mb{Q}_\mr{V}^\T \mb{Q}_\mr{C}^\T \A{L}\\
                    -\mb{W}_\mr{V}^\T \AT{V} \mb{P}_\mr{C} &
                \end{bmatrix} = \begin{bmatrix}
                    \mb{0} & \mb{Q}_\mr{R}^\T \mb{Q}_\mr{V}^\T \mb{Q}_\mr{C}^\T \A{L}\\
                    -\mb{W}_\mr{V}^\T \AT{V} \mb{P}_\mr{C} &
                \end{bmatrix},
            \end{align*}
            thus the basis functions $\mbt{P}$ and $\mbt{Q}$ of $\mbb{W}^\T \mbb{K}_\mb{P}(\mb{x})$ are constant as well, and since $\mbb{W}^\T \mbb{f}(t)$ only depends on $t$, we find a unique solution for $\mbt{x}_\mb{P}$ that also only depends on $t$.}'' (p.19)
            \item We added a larger index two example containing a nonlinear multiport device to illustrate the approach for a more complex example. (p.15-17)
            \item This example also demonstrates that the differential variables may be linear combinations of the original variables.\\
            ``\textcolor{red}{When decoupling the DAE arising from the circuit of \dots\ using the dissection index, one finds
            \begin{align*}
                \mbt{x}_\mb{Q} = \begin{bmatrix}
                    v_{34}\\
                    i_{\mr{L}_2}
                \end{bmatrix}, \quad \mbt{x}_\mb{P} = i_{\mr{L}_1}, \quad \mbb{x}_\mb{P} = \begin{bmatrix}
                    \varphi_2\\
                    \varphi_4
                \end{bmatrix}, \quad \mbb{x}_\mb{Q} = \varphi_1
            \end{align*}
            for the differential and algebraic variables using the notation from \dots. We first observe that there are only two differential variables $\mbt{x}_\mb{Q}$, again leading to a significant reduction in the learning effort. We also see that the first differential variable $v_{34} \coloneqq \varphi_3 - \varphi_4$ is a linear combination of the original variables of the DAE. This is a general phenomenon and does not interfere with our approach, as the original variables may always be reconstructed from the sets of differential and algebraic variables by reversing the splitting using \dots
            \begin{align*}
                \mb{x} = \mb{P} \mbt{x} + \mb{Q} \mbb{x} = \mb{P} \big( \mbt{P} \mbt{x}_\mb{P} + \mbt{Q} \mbt{x}_\mb{Q} \big) + \mb{Q} \big( \mbb{P} \mbb{x}_\mb{P} + \mbb{Q} \mbb{x}_\mb{Q} \big).
            \end{align*}}'' (p.16)
        \end{itemize}

        \item[2)] \textcolor{TUDa-0c}{In the introduction, many general applications for machine learning in complex engineering applications are mentioned. However, I did not understand exactly the application that is considered in the article. In both examples, the parameter $R$ appearing in the explicit constraint is known. Please explain if this is realistic if e.g.~an application after model order reduction was the goal. If not, explain the applications you have in mind.}
        \begin{itemize}
        \item We added a remark that the choice of parameters is arbitrary and only serves to illustrate the approach.\\
        ``\textcolor{red}{We note that the choice to consider $L$ and $C$, and not for example $R$, as parameters here is arbitrary and only serves to illustrate the approach. One could still follow the same approach when considering e.g.~the initial conditions, or the diode model, as being parameterized.}'' (p.11)
        \item We outlined a potential application in the context of electrothermal simulations using the new example. (p.15)
        \item Other potential applications include constrained optimization (with constraints on e.g.~currents through or voltages across certain devices) and validating the behavior of a circuit over its entire range of operating conditions (input currents/voltages, environment temperatures, device parameters) at the end of its development.
        \end{itemize}

        \item[3)] \textcolor{TUDa-0c}{Page 8, line 34-35: Can we always decouple the parameters into those that (may) appear in the algebraic equations and those that (may) appear in the ODE? Are there topological criteria for this? And is your approach reasonable if such a classification does not exist?}
        \begin{itemize}
            \item For the topological basis functions mentioned in section 3 this might be possible, but we have not explored this in more detail.
            \item We pointed this out more explicitly in the manuscript and remarked that for sequential learning approaches, such as the one used in the paper, one can still benefit from parameters not appearing in the ODE without any explicit knowledge.\\
            ``\textcolor{red}{As of now however, we have no knowledge of an easy way to automatically determine which parameters appear in the ODE. But when combined with a sequential learning strategy, such as the one outlined in \dots, there may still be computational savings due to the learning method requiring less samples for the parameters not appearing in the ODE.}'' (p.9)
            \item The approach is reasonable independently of this, since the two main benefits (a reduced number of variables to learn and better accuracy in the algebraic constraints) are independent of this anyhow.
        \end{itemize}

        \item[4)] \textcolor{TUDa-0c}{Page 12, line 32: Please explain how you reconstruct the values for $L$.}
        \begin{itemize}
            \item We added a paragraph that explains the reconstruction in more detail.\\
            ``\textcolor{red}{This results from the fact that the ODE in \dots\ does not depend on $L$, such that the solution of $\varphi_3$ also does not depend on that parameter. The same goes for the algebraic variable $i_\mr{L}$, thus $\varphi_2$ and $\varphi_1$ can be reconstructed only based on the knowledge of $\varphi_3$.}'' (p.13)
        \end{itemize}

        \item[5)] \textcolor{TUDa-0c}{The bases of the dissection index are not unique, I guess. Are the results comparable for all bases? Is it e.g.~favorable to choose orthogonal bases?}
        \begin{itemize}
            \item The bases are indeed not unique. In the context of MNA there is a clear benefit to using the topological basis functions mentioned in section 3, and we added a remark explaining this.\\
            ``\textcolor{red}{When considering large circuits, this topological decoupling along with its topological basis functions is to be preferred over other basis function choices, as it avoids the numerical computation of the basis functions, which becomes prohibitively expensive for large systems.}'' (p.8)
        \end{itemize}

        \item[6)] \textcolor{TUDa-0c}{You emphasize that the approach is independent of the machine learning method used. Is it also applicable for any other decoupling of DAEs into an inherent ODE and the algebraic constraints, also if it is based on other index concepts?}
        \begin{itemize}
            \item In principle it is possible to use any index concept that provides such a decoupling. There are however clear advantages when using the dissection index, in particular in the context of MNA, see also the answer to 5).
            \item We added a remark about this at the beginning of section 3 and also motivated the dissection index more strongly.\\
            ``\textcolor{red}{We now focus on the dissection index. As mentioned in \dots, the key property of this index concept is that it enables the decoupling of a DAE into an ODE and a set of purely algebraic equations. This is clearly different from the perturbation index, however other index concepts such as the tractability and differentiation indices also make such a decoupling possible. Still, the dissection index maintains some advantages over these concepts, as it provides a simple algorithmic procedure for the decoupling that is similar to the tractability index, and it poses less strict smoothness assumptions. In the case of MNA without controlled sources \dots, it is even possible to find a purely topological decoupling based on the dissection index. We will not make use of this topological decoupling here however, but rather consider the dissection index for a more general class of DAEs, to formulate the assumptions that are necessary for our method to work also for DAEs other than \dots.}'' (p.4)
        \end{itemize}
    \end{itemize}

    \paragraph{Reviewing: 2}
    \begin{enumerate}
        \item \textcolor{TUDa-0c}{The manuscript seems to present new and significant results, however the presentation quality is low. The reader is not properly guided through the mathematical derivations and the explanation of the proposed algorithms is weak and sometimes unclear. The proposed algorithm is mixed with derivations of formulas and it is difficult to deduce the proposed procedure starting from equations. For these reasons, although it is interesting, my recommendation is to reject the paper.}
        \begin{itemize}
            \item We reworked section 3, that introduces the dissection index, to more clearly link the ideas presented in the introduction and section 2 with our approach in section 4. Compare also the answer to 6) above, that motivates the dissection index more strongly.

            \item We also used additional assumption (p.4-5) and definition (p.5-6) environments to better highlight the key points of section 3 and to give the section more structure, see also the answer to 1) above.

            \item Furthermore, we shifted math heavy remarks to the appendix, compare the reply to 1) above, in order to streamline the explanation.

            \item To more clearly distinguish the derivation of the dissection index from the introduction of our approach, we also expanded section 4. It now contains additional paragraphs describing the implementation for the index one and two cases.\\
            ``\textcolor{red}{In order to recover the algebraic variables at time $t$ in the index one case, we only need to solve \dots\ for $\mbb{x}(t)$ using the learned $\mbt{x}(t)$.}\\
            \textcolor{red}{In the index two case, we start by solving \dots\ for $\mbt{x}_\mb{P}(t)$ using the learned $\mbt{x}_\mb{Q}(t)$. Since we also require the derivative $\ddt \mbt{x}_\mb{P}(t)$ in \dots, we consider a small time increment $\Delta t$ and approximate the derivative using a backward difference
            \begin{align*}
                \ddt \mbt{x}_\mb{P}(t) \approx \frac{\mbt{x}_\mb{P}(t + \Delta t) - \mbt{x}_\mb{P}(t)}{\Delta t}.
            \end{align*}
            We note that this only reflects our implementation; in principle any finite difference (or similar) approximation is possible. Finally, we determine $\mbb{x}_\mb{P}(t)$ and $\mbb{x}_\mb{Q}(t)$ by jointly solving \dots\ using the learned $\mbt{x}_\mb{Q}(t)$, $\mbt{x}_\mb{P}(t)$ and the approximation of $\ddt \mbt{x}_\mb{P}(t)$.}'' (p.8-9)

            \item Lastly, we removed the alternative approach (previous section 5.1) to achieve a clearer presentation of our main approach.
        \end{itemize}

        \item \textcolor{TUDa-0c}{Since the concept of ``Index'' is key in understanding the manuscript the authors should define it properly and not just cite existent literature.}
        \begin{itemize}
            \item To better introduce the idea of an index, we shortened the introduction and devoted the entire subsection 2.1 to explaining index concepts in more detail. This also includes the definition of a second index concept (the perturbation index) along with an example to better link the term index with electrical circuits, and to motivate the importance of the notion of index in this context. (p.3)
            \item The additional definition environments in section 3 (p.5-6) also serve to emphasize the definition of the dissection index.
        \end{itemize}

        \item \textcolor{TUDa-0c}{Moreover, some symbols and variables are poorly explained or defined, see as an example matrix $G$ appearing in equation 3.}
        \begin{itemize}
            \item We amended the definition of $\mb{G}$ to be directly compatible with the new explanation of multiport devices and the definition of a DAE in standard form, compare 1) above.\\
            ``\textcolor{red}{In order to obtain a version of \dots\ that is better suited to analysis and implementation, we consider the device function Jacobians
            \begin{align}
                \mb{G} \big( \AT{R} \vphi \big) \coloneqq \mb{J}_{\AT{R} \vphi}(\gR), \quad \mb{L}(\i{L}) \coloneqq \mb{J}_{\i{L}}(\phiL), \quad \mb{C} \big( \AT{C} \vphi \big) \coloneqq \mb{J}_{\AT{C} \vphi}(\qC),
            \end{align}
            where we use the same notation for the Jacobians as in \dots.}'' (p.2)
            \item During the reworking of section 2 and section 3, we also took great care to define all of the used symbols/variables.
        \end{itemize}
    \end{enumerate}

    We are confident that, considering our changes and additions, the manuscript now meets all acceptance criteria.
\end{document}

### 2024-01-22 doc/fig/ial.tex
revision

@@
     \node[rectangle, rounded corners, draw=black, semithick] (ae) at (2,2) {\textcolor{TUDa-9a}{AE}};
-    \node[rectangle, rounded corners, draw=black, semithick] (dd) at (5.5,-2) {\textcolor{TUDa-1a}{Differential DOFs}};
-    \node[rectangle, rounded corners, draw=black, semithick] (ad) at (9.5,2) {\textcolor{TUDa-9a}{Algebraic DOFs}};
+    \node[rectangle, rounded corners, draw=black, semithick] (dd) at (5.5,-2) {\textcolor{TUDa-1a}{Differential variables}};
+    \node[rectangle, rounded corners, draw=black, semithick] (ad) at (9.5,2) {\textcolor{TUDa-9a}{Algebraic variables}};
 

### 2024-01-22 code/full_wave_rectifier2_results.m

@@ -353,9 +353,11 @@ if (plot_consistency)
 
     semilogy(t, e_s, t, eh, t, eb);
 
+    return
+
     filename = 'fwr2_solution.csv';
-    csvwrite(filename, [t'; xt_Q; xt_Q_p; xb; xb_p; xb_r]);
+    csvwrite(filename, [t'; xt_Q; xt_Q_p; xb; xb_p; xb_r]');
 
     filename = 'fwr2_consistency.csv';
-    csvwrite(filename, [t'; e_s; eh; eb]);
+    csvwrite(filename, [t'; e_s; eh; eb]');
 end

### 2024-01-22 code/fwr2_sc.m

@@ -46,6 +46,10 @@ Pb = Vb = sym([[0, 0]; [1, 0]; [0, 1]]);
 Qt = null(Wb.' * Kb_P);
 Pt = sym([0; 1; 0]);
 
+% DAE well determined
+Wb.' * Kb_P * Pt
+return
+
 Wt = null((Mt * Qt).');
 Wt = [0; -L_2; L_12];
 Vt = [[1, 0]; [0, L_2]; [0 L_12]];

### 2024-01-23 doc/article/article.tex
revision

@@
 \end{subequations}
-where $\gR$, $\phiL$ and $\qC$ represent functions modeling resistive, capacitive or inductive devices respectively.
+where the left hand side as a whole corresponds to $\mb{F}$ in \eqref{eq:int_dae}, and the solution variables are given by $\mb{x} = [\vphi, \i{L}, \i{V}]^\T$.
+The functions $\gR$, $\phiL$ and $\qC$ model resistive, capacitive or inductive devices respectively, that may each depend on the parameters $\mb{p}$.
 The terms for independent current and voltage sources are given by $\i{s}(t)$ and $\v{s}(t)$, while $\vphi$ denotes the vector of nodal potentials and $\i{L} $, $\i{V}$ are the currents flowing through branches containing inductors or voltage sources respectively.
@@
 These collect the branch to node relations of the underlying electrical network, when considering the branches and nodes as edges and vertices of a directed graph.
-In the context of design optimization or uncertainty quantification, the device functions $\gR$, $\phiL$ or $\qC$ could e.g.~depend on (subsets of) the parameters $\mb{p}$.
 
@@
 \textcolor{red}{Before outlining the dissection index, we want to give a brief introduction to index concepts more generally.
-There is no unique definition of the index of a DAE, but rather there are multiple related concepts that each possess different strengths and weaknesses\cite{mehrmann2015}.
+There are multiple definitions of the index of a DAE, along with related index concepts that each possess different strengths and weaknesses\cite{mehrmann2015}.
 One important aspect that unites these ideas is that they agree in key cases, e.g.~when looking at linear DAEs, and the same also holds true for the dissection index.
@@
 \begin{align}
-    \mb{F}(\mbh{x}', \mbh{x}, t) = \boldsymbol{\varepsilon}(\mbh{x}', \mbh{x}, t), \quad \mbh{x}(0) = \mbh{x}_0, \label{eq:ic_pdae}
+    \mb{F}(\mbh{x}', \mbh{x}, t) = \boldsymbol{\varepsilon}(t), \quad \mbh{x}(0) = \mbh{x}_0, \label{eq:ic_pdae}
 \end{align}
@@
     \begin{align}
-        C \ddt \hat{\varphi}_1 + \frac{1}{R} \hat{\varphi}_1 + \hat{i}_\mr{V} &= \varepsilon_1(\mbh{x}', \mbh{x}, t) \label{eq:ic_vcra}\\
-        -\hat{\varphi}_1 + v_\mr{s}(t) &= \varepsilon_2(\mbh{x}', \mbh{x}, t), \label{eq:ic_vcrb}
+        C \ddt \hat{\varphi}_1 + \frac{1}{R} \hat{\varphi}_1 + \hat{i}_\mr{V} &= \varepsilon_1(t) \label{eq:ic_vcra}\\
+        -\hat{\varphi}_1 + v_\mr{s}(t) &= \varepsilon_2(t), \label{eq:ic_vcrb}
     \end{align}
@@
 \begin{align}
-    \hat{\varphi}_1 = v_\mr{s}(t) - \varepsilon_2(\mbh{x}', \mbh{x}, t). \label{eq:ic_vcrphi}
+    \hat{\varphi}_1 = v_\mr{s}(t) - \varepsilon_2(t). \label{eq:ic_vcrphi}
 \end{align}
@@
 \begin{align*}
-    \hat{i}_\mr{V} &= \varepsilon_1(\mbh{x}', \mbh{x}, t) - C \ddt \big( v_\mr{s}(t) - \varepsilon_2(\mbh{x}', \mbh{x}, t) \big) - \frac{1}{R} \big( v_\mr{s}(t) - \varepsilon_2(\mbh{x}', \mbh{x}, t) \big).
+    \hat{i}_\mr{V} &= \varepsilon_1(t) - C \ddt \big( v_\mr{s}(t) - \varepsilon_2(t) \big) - \frac{1}{R} \big( v_\mr{s}(t) - \varepsilon_2(t) \big).
 \end{align*}
@@
     \| \mbh{x} - \mb{x} \| = \begin{Vmatrix}
-        -\varepsilon_2(\mbh{x}', \mbh{x}, t)\\
-        \varepsilon_1(\mbh{x}', \mbh{x}, t) + C \ddt \varepsilon_2(\mbh{x}', \mbh{x}, t) + \frac{1}{R} \varepsilon_2(\mbh{x}', \mbh{x}, t)
+        -\varepsilon_2(t)\\
+        \varepsilon_1(t) + C \ddt \varepsilon_2(t) + \frac{1}{R} \varepsilon_2(t)
     \end{Vmatrix}, \label{eq:ic_vcrpi}
@@
 \end{theorem}
-The result can be found in terms of the tractability index\cite{tischendorf1999}, the differentiation index\cite{estevez2000ijcta} and the dissection index\cite{jansen2014}.
+The result can be found in terms of the perturbation index\cite{gunther1996}, the tractability index\cite{tischendorf1999}, the differentiation index\cite{estevez2000ijcta} and the dissection index\cite{jansen2014}.
 We again note that there also exist extensive results about when the index of MNA including controlled sources does not exceed $\nu = 2$\cite{estevez2000ijcta}.
@@
 \label{sec:di}
-\textcolor{red}{We now focus on the dissection index.
-As mentioned in \autoref{sec:int}, the key property of this index concept is that it enables the decoupling of a DAE into an ODE and a set of purely algebraic equations.
-This is clearly different from the perturbation index, however other index concepts such as the tractability and differentiation indices also make such a decoupling possible.
-Still, the dissection index maintains some advantages over these concepts, as it provides a simple algorithmic procedure for the decoupling that is similar to the tractability index, and it poses less strict smoothness assumptions.
+\textcolor{red}{We focus on the dissection index, as it enables the decoupling of a DAE into an ODE and a set of purely algebraic equations.
+This is conceptually different from the perturbation index, however other index concepts such as the tractability and differentiation indices also use decoupling strategies.
+Still, the dissection index maintains some advantages over these concepts, as it provides a simple algorithmic procedure for the decoupling that is similar to the tractability index, but poses less strict smoothness assumptions.
 In the case of MNA without controlled sources \eqref{eq:md_mnaii}, it is even possible to find a purely topological decoupling based on the dissection index\cite{jansen2014}.
-We will not make use of this topological decoupling here however, but rather consider the dissection index for a more general class of DAEs, to formulate the assumptions that are necessary for our method to work also for DAEs other than \eqref{eq:md_mnaii}.}
+We will not make use of this topological decoupling in the derivation however, but rather consider the dissection index for a more general class of DAEs, to formulate the assumptions that are necessary for our method to work also for DAEs other than \eqref{eq:md_mnaii}.}
 
@@
 \end{assumption}
-We note that while \autoref{as:itc_xtpdi} is equivalent to the DAE not being underdetermined\cite{jansen2014}, \autoref{as:itc_xtp} is more practical and important for the implementation, but not for the dissection index itself.
+We note that while \autoref{as:itc_xtpdi} is equivalent to the DAE not being underdetermined\cite{jansen2014}, \autoref{as:itc_xtp} is rather important for the implementation, but not for the dissection index itself.
 In fact our approach still works if the basis functions $\mbt{P}(\mbt{x}_\mb{Q})$ and $\mbt{Q}(\mbt{x}_\mb{Q})$ depend on $\mbt{x}_\mb{Q}$, however we focus on the stronger assumption here, since it shortens the expressions in the following without impacting the general idea and \autoref{re:dim_xtp} shows that MNA fulfills an even stronger condition than \autoref{as:itc_xtp}.}
@@
 This becomes an even bigger advantage when knowledge about the solution is only required for specific points in time, since the algebraic equation may be solved pointwise.
-\textcolor{red}{As of now however, we have no knowledge of an easy way to automatically determine which parameters appear in the ODE.
+\textcolor{red}{As of now, we have no easy way to automatically determine which parameters appear in the ODE.
 But when combined with a sequential learning strategy, such as the one outlined in \autoref{sec:ne}, there may still be computational savings due to the learning method requiring less samples for the parameters not appearing in the ODE.}
@@
     \end{center}
-    \caption{Full wave rectifier circuit.}
+    \caption{Full wave rectifier circuit. The resistances of the diodes depend on the parameters $\mb{p} = [T_1, T_2]^\T$.}
     \label{fig:ne_fwr2}

### 2024-01-23 doc/article/letter.tex
revision

@@
 But when combined with a sequential learning strategy, such as the one outlined in \dots, there may still be computational savings due to the learning method requiring less samples for the parameters not appearing in the ODE.}'' (p.9)
-            \item The approach is reasonable independently of this, since the two main benefits (a reduced number of variables to learn and better accuracy in the algebraic constraints) are independent of this anyhow.
+            \item The approach is also reasonable if the number of parameters cannot be reduced by the decoupling, since two significant advantages remain: better accuracy in the algebraic constraints and a reduced number of (differential) variables to learn.
         \end{itemize}
@@
             \item We added a remark about this at the beginning of section 3 and also motivated the dissection index more strongly.\\
-            ``\textcolor{red}{We now focus on the dissection index.
-As mentioned in \dots, the key property of this index concept is that it enables the decoupling of a DAE into an ODE and a set of purely algebraic equations.
-This is clearly different from the perturbation index, however other index concepts such as the tractability and differentiation indices also make such a decoupling possible.
-Still, the dissection index maintains some advantages over these concepts, as it provides a simple algorithmic procedure for the decoupling that is similar to the tractability index, and it poses less strict smoothness assumptions.
+            ``\textcolor{red}{We focus on the dissection index, as it enables the decoupling of a DAE into an ODE and a set of purely algebraic equations.
+This is conceptually different from the perturbation index, however other index concepts such as the tractability and differentiation indices also use decoupling strategies.
+Still, the dissection index maintains some advantages over these concepts, as it provides a simple algorithmic procedure for the decoupling that is similar to the tractability index, but poses less strict smoothness assumptions.
 In the case of MNA without controlled sources \dots, it is even possible to find a purely topological decoupling based on the dissection index.
-We will not make use of this topological decoupling here however, but rather consider the dissection index for a more general class of DAEs, to formulate the assumptions that are necessary for our method to work also for DAEs other than \dots.}'' (p.4)
+We will not make use of this topological decoupling in the derivation however, but rather consider the dissection index for a more general class of DAEs, to formulate the assumptions that are necessary for our method to work also for DAEs other than \dots.}'' (p.4)
         \end{itemize}
@@
 
-            \item Lastly, we removed the alternative approach (previous section 5.1) to achieve a clearer presentation of our main approach.
+            \item Lastly, we removed the alternative approach (previous section 5.1) to achieve a clearer presentation of our main message.
         \end{itemize}

### 2024-01-25 doc/article/arxiv/article.tex
graphical abstract

--- 2024-01-23 doc/article/article.tex
@@
 \usetikzlibrary{external}
-\tikzexternalize
+% \tikzexternalize
 
@@
 
-\textcolor{red}{In order to obtain a version of \eqref{eq:md_mnai} that is better suited to analysis and implementation, we consider the device function Jacobians
+In order to obtain a version of \eqref{eq:md_mnai} that is better suited to analysis and implementation, we consider the device function Jacobians
 \begin{align}
@@
 \end{align}
-where we use the same notation for the Jacobians as in \autoref{sec:int}.} Inserting \eqref{eq:md_dfj} into the original system, while linearizing $\gR$ and writing everything in matrix form, yields
+where we use the same notation for the Jacobians as in \autoref{sec:int}. Inserting \eqref{eq:md_dfj} into the original system, while linearizing $\gR$ and writing everything in matrix form, yields
 \begin{align}
@@
 
-\textcolor{red}{We note that \eqref{eq:md_mnaii} readily extends to multiport devices\cite{tischendorf1999}.
+We note that \eqref{eq:md_mnaii} readily extends to multiport devices\cite{tischendorf1999}.
 A general inductive $n$-port for example, can also be modeled by a function $\phiL(\i{L}) = \big[ \phi_1(\i{L}), \dotsc, \phi_{n-1}(\i{L}) \big]^\T$, however the Jacobian is not necessarily diagonal in this case as the component functions of $\phiL$ can each depend on all the currents $\i{L} = [i_1, \cdots, i_{n-1}]^\T$.
@@
 Interpreting the multiport as a single node and using Kirchhoff's current law then gives $i_0 = -\sum_{k=1}^{n-1} i_k$ for the reference current, when orienting all currents to point toward the device.
-Following this approach, there is then no difference in treating multiports compared to one-ports\cite{tischendorf1999}.}
+Following this approach, there is then no difference in treating multiports compared to one-ports\cite{tischendorf1999}.
 
@@
 \label{subsec:ic}
-\textcolor{red}{Before outlining the dissection index, we want to give a brief introduction to index concepts more generally.
+Before outlining the dissection index, we want to give a brief introduction to index concepts more generally.
 There are multiple definitions of the index of a DAE, along with related index concepts that each possess different strengths and weaknesses\cite{mehrmann2015}.
@@
 The idea behind this definition is to capture the impact of perturbations on the solution, as the name suggests.
-Usually these pertubations are assumed small, in the sense that $\| \boldsymbol{\varepsilon} \|_\infty \ll 1$, but fast changing, such that $\| \boldsymbol{\varepsilon}^{(\nu)} \|_\infty$ may grow very quickly in $\nu$.}
+Usually these pertubations are assumed small, in the sense that $\| \boldsymbol{\varepsilon} \|_\infty \ll 1$, but fast changing, such that $\| \boldsymbol{\varepsilon}^{(\nu)} \|_\infty$ may grow very quickly in $\nu$.
 
@@
 \end{figure}
-\textcolor{red}{To illustrate the perturbation index, and also to hint at its relevance for circuit simulation, we consider the small example given in \autoref{fig:ic_vcr}.
+To illustrate the perturbation index, and also to hint at its relevance for circuit simulation, we consider the small example given in \autoref{fig:ic_vcr}.
 Applying MNA to the circuit and introducing a perturbation yields the following DAE
@@
 \end{align}
-so the unperturbed system corresponding to \eqref{eq:ic_vcr} has perturbation index $\nu = 2$ as \eqref{eq:ic_vcrpi} depends on the first derivative of $\varepsilon_2$.}
+so the unperturbed system corresponding to \eqref{eq:ic_vcr} has perturbation index $\nu = 2$ as \eqref{eq:ic_vcrpi} depends on the first derivative of $\varepsilon_2$.
 
@@
 \label{sec:di}
-\textcolor{red}{We focus on the dissection index, as it enables the decoupling of a DAE into an ODE and a set of purely algebraic equations.
+We focus on the dissection index, as it enables the decoupling of a DAE into an ODE and a set of purely algebraic equations.
 This is conceptually different from the perturbation index, however other index concepts such as the tractability and differentiation indices also use decoupling strategies.
@@
 In the case of MNA without controlled sources \eqref{eq:md_mnaii}, it is even possible to find a purely topological decoupling based on the dissection index\cite{jansen2014}.
-We will not make use of this topological decoupling in the derivation however, but rather consider the dissection index for a more general class of DAEs, to formulate the assumptions that are necessary for our method to work also for DAEs other than \eqref{eq:md_mnaii}.}
-
-\textcolor{red}{We consider a DAE in standard form\cite{jansen2014},
+We will not make use of this topological decoupling in the derivation however, but rather consider the dissection index for a more general class of DAEs, to formulate the assumptions that are necessary for our method to work also for DAEs other than \eqref{eq:md_mnaii}.
+
+We consider a DAE in standard form\cite{jansen2014},
 \begin{align}
@@
 In the following, we will demonstrate the first two steps of the dissection index when applied to systems of the form of \eqref{eq:di_dae}, while stating the assumptions of our approach.
-Appendix \autoref{sec:dim} contains additional remarks showing that these assumptions are fulfilled by \eqref{eq:md_mnaii}.}
+Appendix \autoref{sec:dim} contains additional remarks showing that these assumptions are fulfilled by \eqref{eq:md_mnaii}.
 
@@
 We now make the following assumption.
-\textcolor{red}{\begin{assumption}
+\begin{assumption}
     \label{as:ioc_bf}
@@
 \autoref{as:ioc_bf} may seem restrictive, but it is fulfilled by many systems occurring in practical applications\cite{jansen2014}.
-This in particular includes MNA, as \autoref{re:dim_bf} shows.} The key idea of the dissection index is to use the basis functions to split the solution variables $\mb{x}$ into two parts
+This in particular includes MNA, as \autoref{re:dim_bf} shows.
+The key idea of the dissection index is to use the basis functions to split the solution variables $\mb{x}$ into two parts
 \begin{align}
@@
 Using the fact that $\mbt{M}(\mb{x})$ is regular by construction\cite{jansen2014}, we can now define the index one case of the dissection index.
-\textcolor{red}{\begin{definition}
+\begin{definition}
     \label{def:ioc_io}
     The DAE \eqref{eq:di_dae} has dissection index $\nu = 1$ if $\mbb{K}_\mb{Q}(\mb{x})$ is regular.
-\end{definition}}
+\end{definition}
 This is motivated by the observation that \eqref{eq:ioc_iob} is a purely algebraic equation with a locally unique solution for $\mbb{x}$ in terms of $\mbt{x}$, given that $\mbb{K}_\mb{Q}(\mb{x})$ is regular.
@@
 We now make the following two assumptions.
-\textcolor{red}{\begin{assumption}
+\begin{assumption}
     \label{as:itc_xtpdi}
@@
 We note that while \autoref{as:itc_xtpdi} is equivalent to the DAE not being underdetermined\cite{jansen2014}, \autoref{as:itc_xtp} is rather important for the implementation, but not for the dissection index itself.
-In fact our approach still works if the basis functions $\mbt{P}(\mbt{x}_\mb{Q})$ and $\mbt{Q}(\mbt{x}_\mb{Q})$ depend on $\mbt{x}_\mb{Q}$, however we focus on the stronger assumption here, since it shortens the expressions in the following without impacting the general idea and \autoref{re:dim_xtp} shows that MNA fulfills an even stronger condition than \autoref{as:itc_xtp}.}
+In fact our approach still works if the basis functions $\mbt{P}(\mbt{x}_\mb{Q})$ and $\mbt{Q}(\mbt{x}_\mb{Q})$ depend on $\mbt{x}_\mb{Q}$, however we focus on the stronger assumption here, since it shortens the expressions in the following without impacting the general idea and \autoref{re:dim_xtp} shows that MNA fulfills an even stronger condition than \autoref{as:itc_xtp}.
 
@@
 Together with \eqref{eq:itc_xtp} and \eqref{eq:itc_xbp} providing locally unique solutions for $\mbt{x}_\mb{P}$ and $\mbb{x}_\mb{P}$ respectively, this motivates the following definition.
-\textcolor{red}{\begin{definition}
+\begin{definition}
     \label{def:itc_it}
@@
 The differential part \eqref{eq:itc_itiiia} then describes an ODE in the index two differential variables $\mbt{x}_\mb{Q}$, analogous to the previous case, and we call $\mbt{x}_\mb{Q}$ the differential variables and $[\mbt{x}_\mb{P}, \mbb{x}_\mb{P}, \mbb{x}_\mb{Q}]^\T$ the algebraic variables.
-We note that the procedure may be continued for even higher index DAEs, by repeating the steps of the index two case with $\mbt{x}_\mb{Q}$ and $\mbb{x}_\mb{Q}$ playing the roles of $\mbt{x}$ and $\mbb{x}$ respectively.}
+We note that the procedure may be continued for even higher index DAEs, by repeating the steps of the index two case with $\mbt{x}_\mb{Q}$ and $\mbb{x}_\mb{Q}$ playing the roles of $\mbt{x}$ and $\mbb{x}$ respectively.
 
@@
 As hinted at in \autoref{sec:di}, it is possible to find a purely topological decoupling based on the dissection index\cite{jansen2014}, which agrees with the intuition given by the index result from \autoref{th:ic_imna}.
-\textcolor{red}{When considering large circuits, this topological decoupling along with its topological basis functions is to be preferred over other basis function choices, as it avoids the numerical computation of the basis functions, which becomes prohibitively expensive for large systems.} We also note that a similar topological decoupling result exists for circuits including controlled sources, however only for the first step of the decoupling, as it is framed in the context of semi-explicit methods for which one only requires a DAE in semi-explicit form\cite{jansen2014}.
+When considering large circuits, this topological decoupling along with its topological basis functions is to be preferred over other basis function choices, as it avoids the numerical computation of the basis functions, which becomes prohibitively expensive for large systems.
+We also note that a similar topological decoupling result exists for circuits including controlled sources, however only for the first step of the decoupling, as it is framed in the context of semi-explicit methods for which one only requires a DAE in semi-explicit form\cite{jansen2014}.
 An extension of this result to circuits of higher index is within the scope of further research.
@@
 \paragraph{Index one case}
-\textcolor{red}{In order to recover the algebraic variables at time $t$ in the index one case, we only need to solve \eqref{eq:ioc_iob} for $\mbb{x}(t)$ using the learned $\mbt{x}(t)$.}
+In order to recover the algebraic variables at time $t$ in the index one case, we only need to solve \eqref{eq:ioc_iob} for $\mbb{x}(t)$ using the learned $\mbt{x}(t)$.
 
 \paragraph{Index two case}
-\textcolor{red}{In the index two case, we start by solving \eqref{eq:itc_xtp} for $\mbt{x}_\mb{P}(t)$ using the learned $\mbt{x}_\mb{Q}(t)$.
+In the index two case, we start by solving \eqref{eq:itc_xtp} for $\mbt{x}_\mb{P}(t)$ using the learned $\mbt{x}_\mb{Q}(t)$.
 Since we also require the derivative $\ddt \mbt{x}_\mb{P}(t)$ in \eqref{eq:itc_itiiib}, we consider a small time increment $\Delta t$ and approximate the derivative using a backward difference
@@
 We note that this only reflects our implementation; in principle any finite difference (or similar) approximation is possible.
-Finally, we determine $\mbb{x}_\mb{P}(t)$ and $\mbb{x}_\mb{Q}(t)$ by jointly solving \eqref{eq:itc_xbp} and \eqref{eq:itc_itiiib} using the learned $\mbt{x}_\mb{Q}(t)$, $\mbt{x}_\mb{P}(t)$ and the approximation of $\ddt \mbt{x}_\mb{P}(t)$.}
+Finally, we determine $\mbb{x}_\mb{P}(t)$ and $\mbb{x}_\mb{Q}(t)$ by jointly solving \eqref{eq:itc_xbp} and \eqref{eq:itc_itiiib} using the learned $\mbt{x}_\mb{Q}(t)$, $\mbt{x}_\mb{P}(t)$ and the approximation of $\ddt \mbt{x}_\mb{P}(t)$.
 
@@
 This becomes an even bigger advantage when knowledge about the solution is only required for specific points in time, since the algebraic equation may be solved pointwise.
-\textcolor{red}{As of now, we have no easy way to automatically determine which parameters appear in the ODE.
-But when combined with a sequential learning strategy, such as the one outlined in \autoref{sec:ne}, there may still be computational savings due to the learning method requiring less samples for the parameters not appearing in the ODE.}
+As of now, we have no easy way to automatically determine which parameters appear in the ODE.
+But when combined with a sequential learning strategy, such as the one outlined in \autoref{sec:ne}, there may still be computational savings due to the learning method requiring less samples for the parameters not appearing in the ODE.
 
@@
 The starting points for the strategy are given by all combinations of the boundary values for $L$ and $C$ together with $t = \SI{0}{\ms}$ and $t = \SI{10}{\ms}$ once for each combination.
-\textcolor{red}{We note that the choice to consider $L$ and $C$, and not for example $R$, as parameters here is arbitrary and only serves to illustrate the approach.
-One could still follow the same approach when considering e.g.~the initial conditions, or the diode model, as being parameterized.} Recalling \eqref{eq:itc_do1}, we see that $\varphi_3$ and $i_\mr{L}$ have to be learned as the differential variables.
+We note that the choice to consider $L$ and $C$, and not for example $R$, as parameters here is arbitrary and only serves to illustrate the approach.
+One could still follow the same approach when considering e.g.~the initial conditions, or the diode model, as being parameterized.
+Recalling \eqref{eq:itc_do1}, we see that $\varphi_3$ and $i_\mr{L}$ have to be learned as the differential variables.
 In addition to these two, we also consider the algebraic variable $\varphi_2$ and we then compare the accuracy of learning $\varphi_2$ directly to recovering it from the two differential variables.
@@
 The 7 unique values for $L$, that were requested for learning $\varphi_2$, did not require full simulations however, but rather could be reconstructed from \eqref{eq:itc_do2}.
-\textcolor{red}{This results from the fact that the ODE in \eqref{eq:itc_do2} does not depend on $L$, such that the solution of $\varphi_3$ also does not depend on that parameter.
-The same goes for the algebraic variable $i_\mr{L}$, thus $\varphi_2$ and $\varphi_1$ can be reconstructed only based on the knowledge of $\varphi_3$.} We again emphasize that time is also included as a parameter within the sample selection strategy, such that the reconstructions of the algebraic variables only need to be evaluated at the particular points in time that are requested by the strategy, rather than at all time points of the solution.
+This results from the fact that the ODE in \eqref{eq:itc_do2} does not depend on $L$, such that the solution of $\varphi_3$ also does not depend on that parameter.
+The same goes for the algebraic variable $i_\mr{L}$, thus $\varphi_2$ and $\varphi_1$ can be reconstructed only based on the knowledge of $\varphi_3$.
+We again emphasize that time is also included as a parameter within the sample selection strategy, such that the reconstructions of the algebraic variables only need to be evaluated at the particular points in time that are requested by the strategy, rather than at all time points of the solution.
 \begin{figure}[t]
@@
 \end{figure}
-\textcolor{red}{As a third and larger example we consider the rectifier circuit from \autoref{fig:ne_fwr2}.
+As a third and larger example we consider the rectifier circuit from \autoref{fig:ne_fwr2}.
 Aside from requiring the full index two implementation from \autoref{sec:ial}, this example also showcases another potential application in the context of electrothermal simulations.
@@
 We also note that rectifiers play an important role in many power electronics applications; they are key when converting high AC to lower DC voltages.
-In the example we still use a current source, since this leads to an index two circuit in this case, compare \autoref{th:ic_imna}, which requires the full index two approach from \autoref{sec:ial}.}
+In the example we still use a current source, since this leads to an index two circuit in this case, compare \autoref{th:ic_imna}, which requires the full index two approach from \autoref{sec:ial}.
 \begin{figure}[t]
@@
 \end{figure}
-\textcolor{red}{When decoupling the DAE arising from the circuit of \autoref{fig:ne_fwr2} using the dissection index, one finds
+When decoupling the DAE arising from the circuit of \autoref{fig:ne_fwr2} using the dissection index, one finds
 \begin{align*}
@@
 For the training data we consider temperatures $\SI{20}{\celsius} \leq T_1, T_2 \leq \SI{90}{\celsius}$, and the starting values for the sample selection strategy are again obtained by considering all combinations of the boundary values together with $t = \SI{0}{\ms}$ and $t = \SI{50}{\ms}$.
-We then execute the learning strategy up to a tolerance of $5 \cdot 10^{-3}$ and choose the prediction point $T_1 = \SI{65}{\celsius}$, $T_2 = \SI{85}{\celsius}$ that is not included in the training data.}
-
-\textcolor{red}{The solutions of the differential and algebraic variables, for $T_1 = 65\, \si{\celsius}$ and $T_2 = 85\, \si{\celsius}$, are shown in \autoref{fig:ne_fwr2_solution_d} and \autoref{fig:ne_fwr2_solution_a}, while \autoref{fig:ne_fwr2_error} shows the differences between the mean predictions and simulations for the same temperature values.
+We then execute the learning strategy up to a tolerance of $5 \cdot 10^{-3}$ and choose the prediction point $T_1 = \SI{65}{\celsius}$, $T_2 = \SI{85}{\celsius}$ that is not included in the training data.
+
+The solutions of the differential and algebraic variables, for $T_1 = 65\, \si{\celsius}$ and $T_2 = 85\, \si{\celsius}$, are shown in \autoref{fig:ne_fwr2_solution_d} and \autoref{fig:ne_fwr2_solution_a}, while \autoref{fig:ne_fwr2_error} shows the differences between the mean predictions and simulations for the same temperature values.
 The comparatively larger differences for $i_{\mr{L}_1}$ in \autoref{fig:ne_fwr2_error} stem from the prediction error being defined as a relative quantity and the absolute values of $i_{\mr{L}_1}$ being much larger than the others, compare \autoref{fig:ne_fwr2_solution_d}.
@@
 For $T_1 = 65\, \si{\celsius}$ and $T_2 = 85\, \si{\celsius}$ the consistency errors can be seen in \autoref{fig:ne_fwr2_consistency}.
-We observe that the reconstructed solution again outperforms the directly learned solution by several orders of magnitude.}
+We observe that the reconstructed solution again outperforms the directly learned solution by several orders of magnitude.
 \begin{figure}
@@
 
-% \begin{table}
-%     \caption{}
-%     \begin{center}
-%         \begin{tabular}{}
-%             \toprule
-%             & &\\
-%             \midrule
-%             & &\\
-%             \bottomrule
-%         \end{tabular}
-%     \end{center}
-%     \begin{tablenotes}
-%         \item notes
-%     \end{tablenotes}
-% \end{table}
-
-% \begin{algorithm}
-%     \begin{algorithmic}
-%     \end{algorithmic}
-% \end{algorithm}
-
 \section*{Acknowledgments}
@@
 
-\textcolor{red}{\begin{remark}
+\begin{remark}
     \label{re:dim_xtp}
@@
     thus the basis functions $\mbt{P}$ and $\mbt{Q}$ of $\mbb{W}^\T \mbb{K}_\mb{P}(\mb{x})$ are constant as well, and since $\mbb{W}^\T \mbb{f}(t)$ only depends on $t$, we find a unique solution for $\mbt{x}_\mb{P}$ that also only depends on $t$.
-\end{remark}}
-
-\bibliography{../../../ai-twilight/biblio}
-% \bibliography{biblio}
-
-% \input{../fig/vcr}
-% \input{../fig/do1}
-% \input{../fig/do2}
-% \input{../fig/ial}
-% \input{../fig/solution}
-% \input{../fig/convergence}
-% \input{../fig/full_wave_rectifier}
+\end{remark}
+
+\bibliography{biblio}
 

### 2024-01-25 code/fwr2_sc.m
graphical abstract

@@ -48,7 +48,6 @@ Pt = sym([0; 1; 0]);
 
 % DAE well determined
 Wb.' * Kb_P * Pt
-return
 
 Wt = null((Mt * Qt).');
 Wt = [0; -L_2; L_12];

### 2024-01-26 code/tc_sc.m
cleanup

@@ -1,137 +1,0 @@
-pkg load symbolic
-
-syms v_1 v_2 v_3 v_4 v_5 v_6 v_7 v_8 v_9 v_10 v_11 v_12 v_13 v_14 v_15 v_16
-syms i_L_1 i_L_2 i_L_3 i_L_4 i_L_5 i_L_6 i_L_7
-syms i_V_1 i_V_2
-
-syms R_1 R_2 R_3 R_4 R_5 R_6 R_7
-syms L_1 L_2 L_12 L_3 L_4 L_34 L_5 L_6 L_56 L_7
-syms C_1 C_2 C_3 C_4 C_5 C_6 C_7 C_8 C_9
-syms v_s_1 v_s_2
-
-n_v   = 16;
-n_R   = 7;
-n_L   = 7;
-n_C   = 9;
-n_V   = 2;
-
-A_R       = zeros(n_v,n_R);
-A_R(1,1)  = 1;
-A_R(2,1)  = -1;
-A_R(8,2)  = 1;
-A_R(9,2)  = -1;
-A_R(3,3)  = 1;
-A_R(4,3)  = -1;
-A_R(10,4) = 1;
-A_R(11,4) = -1;
-A_R(5,5)  = 1;
-A_R(6,5)  = -1;
-A_R(12,6) = 1;
-A_R(13,6) = -1;
-A_R(12,7) = 1;
-A_R(15,7) = -1;
-
-A_L       = zeros(n_v,n_L);
-A_L(2,1)  = 1;
-A_L(3,1)  = -1;
-A_L(9,2)  = 1;
-A_L(10,2) = -1;
-A_L(4,3)  = 1;
-A_L(5,3)  = -1;
-A_L(11,4) = 1;
-A_L(12,4) = -1;
-A_L(6,5)  = 1;
-A_L(7,5)  = -1;
-A_L(13,6) = 1;
-A_L(14,6) = -1;
-A_L(15,7) = 1;
-A_L(16,7) = -1;
-
-A_C       = zeros(n_v,n_C);
-A_C(3,1)  = 1;
-A_C(10,2) = 1;
-A_C(3,2)  = -1;
-A_C(5,3)  = 1;
-A_C(12,4) = 1;
-A_C(5,4)  = -1;
-A_C(7,5)  = 1;
-A_C(14,6) = 1;
-A_C(7,6)  = -1;
-A_C(10,7) = -1;
-A_C(12,8) = -1;
[…]

### 2024-01-26 code/test_circuit.m

@@ -1,148 +1,0 @@
-T_0    = 0;
-T      = 3;
-Deltat = 0.1;
-
-G_1 = G_2 = G_3 = G_4 = G_5 = G_6 = G_7 = 1;
-L_1 = L_2 = L_3 = L_4 = L_5 = L_6 = L_7 = 1;
-L_12 = L_34 = L_56 = 0.1;
-C_1 = C_2 = C_3 = C_4 = C_5 = C_6 = C_7 = C_8 = C_9 = 1;
-
-v_s_1 = @(t) sin(4*t) + cos(2*t);
-v_s_2 = @(t) sin(3*t) + cos(t);
-
-n_v   = 16;
-n_R   = 7;
-n_L   = 7;
-n_C   = 9;
-n_V   = 2;
-
-A_R       = zeros(n_v,n_R);
-A_R(1,1)  = 1;
-A_R(2,1)  = -1;
-A_R(8,2)  = 1;
-A_R(9,2)  = -1;
-A_R(3,3)  = 1;
-A_R(4,3)  = -1;
-A_R(10,4) = 1;
-A_R(11,4) = -1;
-A_R(5,5)  = 1;
-A_R(6,5)  = -1;
-A_R(12,6) = 1;
-A_R(13,6) = -1;
-A_R(12,7) = 1;
-A_R(15,7) = -1;
-
-A_L       = zeros(n_v,n_L);
-A_L(2,1)  = 1;
-A_L(3,1)  = -1;
-A_L(9,2)  = 1;
-A_L(10,2) = -1;
-A_L(4,3)  = 1;
-A_L(5,3)  = -1;
-A_L(11,4) = 1;
-A_L(12,4) = -1;
-A_L(6,5)  = 1;
-A_L(7,5)  = -1;
-A_L(13,6) = 1;
-A_L(14,6) = -1;
-A_L(15,7) = 1;
-A_L(16,7) = -1;
-
-A_C       = zeros(n_v,n_C);
-A_C(3,1)  = 1;
-A_C(10,2) = 1;
-A_C(3,2)  = -1;
-A_C(5,3)  = 1;
-A_C(12,4) = 1;
-A_C(5,4)  = -1;
-A_C(7,5)  = 1;
-A_C(14,6) = 1;
-A_C(7,6)  = -1;
-A_C(10,7) = -1;
-A_C(12,8) = -1;
-A_C(16,9) = -1;
-
[…]

### 2024-02-26 doc/article/arxiv/article.tex
alternative definition of G

@@
 
-In order to obtain a version of \eqref{eq:md_mnai} that is better suited to analysis and implementation, we consider the device function Jacobians
+\textcolor{red}{In order to obtain a version of \eqref{eq:md_mnai} that is better suited to analysis and implementation, we consider the device function Jacobians\footnote{\textcolor{red}{Note that, for nonzero denominators, $\mb{G}$ may also be defined componentwise as
+\begin{align*}
+    \big[ \mb{G} \big( \AT{R} \vphi \big) \big]_{ij} = \frac{\big[ \gR(\AT{R} \vphi) \big]_i}{\big[ \AT{R} \vphi \big]_j},
+\end{align*}
+such that \eqref{eq:md_mnai} and \eqref{eq:md_mnaii} become truly equivalent.
+(The components with zero denominators are set to zero.) This alternative definition has no impact on our approach, however practical implementations often work with $\mb{G}$ as defined in \eqref{eq:md_dfj}.}}}
 \begin{align}
@@
 \end{align}
-where we use the same notation for the Jacobians as in \autoref{sec:int}. Inserting \eqref{eq:md_dfj} into the original system, while linearizing $\gR$ and writing everything in matrix form, yields
+where we use the same notation for the Jacobians as in \autoref{sec:int}. Inserting \eqref{eq:md_dfj} into the original system and writing everything in matrix form yields
 \begin{align}

### 2024-04-05 doc/article/arxiv/article.tex
final revision

--- 2024-01-25 doc/article/arxiv/article.tex
@@
 \newcommand{\vphi}{\boldsymbol{\varphi}}
+% \iL, \iV, \is, \vs
 \renewcommand{\i}[1]{\mb{i}_\mr{#1}}
@@
 
-In order to obtain a version of \eqref{eq:md_mnai} that is better suited to analysis and implementation, we consider the device function Jacobians
+In order to obtain a version of \eqref{eq:md_mnai} that is better suited to analysis and implementation, we consider the device function Jacobians\footnote{Note that the definition of $\mb{G}$ via the Jacobian only serves to obtain a matrix form in \eqref{eq:md_mnaii} and has no impact on the general approach.
+As such, one may also work directly with the original system from \eqref{eq:md_mnai}, however software implementations of MNA often work with $\mb{G}$ as defined in \eqref{eq:md_dfj} (e.g.~when using Newton's method for solving nonlinear systems).}
 \begin{align}

### 2024-04-05 doc/article/arxiv/article_highlighted.tex
graphical abstract
final revision

--- 2024-01-23 doc/article/article.tex
@@
 
-\textcolor{red}{In order to obtain a version of \eqref{eq:md_mnai} that is better suited to analysis and implementation, we consider the device function Jacobians
+\textcolor{red}{In order to obtain a version of \eqref{eq:md_mnai} that is better suited to analysis and implementation, we consider the device function Jacobians\footnote{\textcolor{red}{Note that the definition of $\mb{G}$ via the Jacobian only serves to obtain a matrix form in \eqref{eq:md_mnaii} and has no impact on the general approach.
+As such, one may also work directly with the original system from \eqref{eq:md_mnai}, however software implementations of MNA often work with $\mb{G}$ as defined in \eqref{eq:md_dfj} (e.g.~when using Newton's method for solving nonlinear systems).}}}
 \begin{align}
@@
 \end{align}
-where we use the same notation for the Jacobians as in \autoref{sec:int}.} Inserting \eqref{eq:md_dfj} into the original system, while linearizing $\gR$ and writing everything in matrix form, yields
+where we use the same notation for the Jacobians as in \autoref{sec:int}. Inserting \eqref{eq:md_dfj} into the original system and writing everything in matrix form yields
 \begin{align}

### 2024-04-05 doc/article/arxiv/letter.tex
final revision

--- 2024-01-23 doc/article/letter.tex
@@
     \begin{itemize}
-        \item[1)] \textcolor{TUDa-0c}{Both examples are well selected from a didactic point of view.
-They contain constant $C$ and $L$ and present a rather trivial DAE-structure, since the basis considered in the dissection concept consist of standard unit vectors $e_i$.
-However, they do clarify all challenges that may appear in general.
-It should be explained that due to the properties of the incidence matrices, in general the ODEs have to be formulated for linear combinations of components.
-Moreover, for nonlinear $C$ and $G$ corresponding to multiport devices, it should be discussed if such a decoupling is still possible.
-On the one hand the theory should be more precise and on the other hand at least one additional larger nonlinear index-2 example with a more complex structure should be included, e.g.~a CV-loop involving several nonlinear capacitors.}
+        \item[1)] \textcolor{TUDa-0c}{Why do you have to linearize $g_R$ in equation (4)?
+This is not necessary for the well-known index-results for the MNA.
+Is this linearization important for your approach with the dissection index?}
         \begin{itemize}
-            \item We added a paragraph in section 2 to illustrate that the decoupling also works for nonlinear multiport devices.\\
-            ``\textcolor{red}{We note that \dots\ readily extends to multiport devices.
-A general inductive $n$-port for example, can also be modeled by a function $\phiL(\i{L}) = \big[ \phi_1(\i{L}), \dotsc, \phi_{n-1}(\i{L}) \big]^\T$, however the Jacobian is not necessarily diagonal in this case as the component functions of $\phiL$ can each depend on all the currents $\i{L} = [i_1, \cdots, i_{n-1}]^\T$.
-Regarding the incidence matrices, one chooses a reference node $\varphi_0$ for the multiport device and then considers $n-1$ branches from the remaining $n-1$ nodes to the reference.
-Interpreting the multiport as a single node and using Kirchhoff's current law then gives $i_0 = -\sum_{k=1}^{n-1} i_k$ for the reference current, when orienting all currents to point toward the device.
-Following this approach, there is then no difference in treating multiports compared to one-ports.}'' (p.2)
-            \item We made the mathematical presentation of the dissection index more precise, reformulating section 3 in terms of nonlinear DAEs in standard form while providing the assumptions that are necessary for our approach.
-We also added an appendix with remarks showing that MNA fulfills all assumptions.\\
-            ``\textcolor{red}{We consider a DAE in standard form,
-            \begin{align}
-                \mb{M}(\mb{x}) \ddt \mb{x} + \mb{K}(\mb{x}) \mb{x} + \mb{f}(t) = \mb{0}, \quad \ker \mb{M} \supsetneq \{ \mb{0} \},
-            \end{align}
-            where the matrix-valued functions $\mb{M}$ and $\mb{K}$ are derived from \dots\ by defining
-            \begin{align*}
-                \mb{M}(\mb{x}) = \mb{J}_{\mb{x}'} \big( \mb{F}(\mb{x}', \mb{x}, t) \big), \quad \mb{K}(\mb{x}) = \mb{J}_{\mb{x}} \big( \mb{F}(\mb{x}', \mb{x}, t) \big).
-            \end{align*}
-            Note that the description of MNA in \dots\ is precisely of this form.
-In the following, we will demonstrate the first two steps of the dissection index when applied to systems of the form of \dots, while stating the assumptions of our approach.
-Appendix \dots\ contains additional remarks showing that these assumptions are fulfilled by \dots.}'' (p.4)\\
-            See also Assumption 1 (p.4), Assumption 2 and Assumption 3 (p.5) as well as Remark 2 (p.19).\\
-            ``\textcolor{red}{Again recalling \dots\ we observe
-            \begin{align*}
-                \mb{K}(\mb{x}) = \begin{bmatrix}
-                    \A{R}^{\phantom{\T}} \mb{G} \big( \AT{R} \vphi \big) \AT{R} & \A{L} & \A{V}\\
-                    -\AT{L} & &\\
-                    -\AT{V} & &
-                \end{bmatrix}.
-            \end{align*}
-            Noting that $\mb{W} = \mb{Q}$ for MNA, due to the symmetry of $\mb{M}(\mb{x})$ in \dots, we find
-            \begin{align*}
-                \mbb{K}_\mb{Q}(\mb{x}) &= \mb{Q}^\T \mb{K}(\mb{x}) \mb{Q} = \begin{bmatrix}
-                    \mb{Q}_\mr{C}^\T \A{R}^{\phantom{\T}} \mb{G} \big( \AT{R} \vphi \big) \AT{R} \mb{Q}_\mr{C} & \mb{Q}_\mr{C}^\T \A{V}\\
-                    -\AT{V} \mb{Q}_\mr{C} &
-                \end{bmatrix}\\
-                \mbb{K}_\mb{P}(\mb{x}) &= \mb{Q}^\T \mb{K}(\mb{x}) \mb{P} = \begin{bmatrix}
-                    \mb{Q}_\mr{C}^\T \A{R}^{\phantom{\T}} \mb{G} \big( \AT{R} \vphi \big) \AT{R} \mb{P}_\mr{C} & \mb{Q}_\mr{C}^\T \A{L}\\
-                    -\AT{V} \mb{P}_\mr{C} &
-                \end{bmatrix}.
-            \end{align*}
-            Determining $\mbb{W}(\mb{x})$ using $\mbb{K}_\mb{Q}(\mb{x})$ gives
-            \begin{align*}
-                \mbb{W} = \begin{bmatrix}
-                    \mb{Q}_\mr{V} \mb{Q}_\mr{R} &\\
-                    & \mb{W}_\mr{V}
-                \end{bmatrix},
-            \end{align*}
-            where $\mr{im\, } \mb{Q}_\mr{V} = \ker \AT{V} \mb{Q}_\mr{C}$, $\mr{im\, } \mb{Q}_\mr{R} = \ker \AT{R} \mb{Q}_\mr{C} \mb{Q}_\mr{V}$ and $\mr{im\, } \mb{W}_\mr{V} = \ker \AT{V} \mb{Q}_\mr{C}$ are all constant such that $\mbb{W}$ is constant as well.
-This finally yields
-            \begin{align*}
-                \mbb{W}^\T \mbb{K}_\mb{P}(\mb{x}) = \begin{bmatrix}
-                    \mb{Q}_\mr{R}^\T \mb{Q}_\mr{V}^\T \mb{Q}_\mr{C}^\T \A{R}^{\phantom{\T}} \mb{G} \big( \AT{R} \vphi \big) \AT{R} \mb{P}_\mr{C} & \mb{Q}_\mr{R}^\T \mb{Q}_\mr{V}^\T \mb{Q}_\mr{C}^\T \A{L}\\
-                    -\mb{W}_\mr{V}^\T \AT{V} \mb{P}_\mr{C} &
-                \end{bmatrix} = \begin{bmatrix}
-                    \mb{0} & \mb{Q}_\mr{R}^\T \mb{Q}_\mr{V}^\T \mb{Q}_\mr{C}^\T \A{L}\\
-                    -\mb{W}_\mr{V}^\T \AT{V} \mb{P}_\mr{C} &
-                \end{bmatrix},
-            \end{align*}
-            thus the basis functions $\mbt{P}$ and $\mbt{Q}$ of $\mbb{W}^\T \mbb{K}_\mb{P}(\mb{x})$ are constant as well, and since $\mbb{W}^\T \mbb{f}(t)$ only depends on $t$, we find a unique solution for $\mbt{x}_\mb{P}$ that also only depends on $t$.}'' (p.19)
-            \item We added a larger index two example containing a nonlinear multiport device to illustrate the approach for a more complex example. (p.15-17)
-            \item This example also demonstrates that the differential variables may be linear combinations of the original variables.\\
-            ``\textcolor{red}{When decoupling the DAE arising from the circuit of \dots\ using the dissection index, one finds
-            \begin{align*}
-                \mbt{x}_\mb{Q} = \begin{bmatrix}
-                    v_{34}\\
-                    i_{\mr{L}_2}
-                \end{bmatrix}, \quad \mbt{x}_\mb{P} = i_{\mr{L}_1}, \quad \mbb{x}_\mb{P} = \begin{bmatrix}
-                    \varphi_2\\
-                    \varphi_4
-                \end{bmatrix}, \quad \mbb{x}_\mb{Q} = \varphi_1
-            \end{align*}
-            for the differential and algebraic variables using the notation from \dots.
-We first observe that there are only two differential variables $\mbt{x}_\mb{Q}$, again leading to a significant reduction in the learning effort.
-We also see that the first differential variable $v_{34} \coloneqq \varphi_3 - \varphi_4$ is a linear combination of the original variables of the DAE.
-This is a general phenomenon and does not interfere with our approach, as the original variables may always be reconstructed from the sets of differential and algebraic variables by reversing the splitting using \dots
-            \begin{align*}
-                \mb{x} = \mb{P} \mbt{x} + \mb{Q} \mbb{x} = \mb{P} \big( \mbt{P} \mbt{x}_\mb{P} + \mbt{Q} \mbt{x}_\mb{Q} \big) + \mb{Q} \big( \mbb{P} \mbb{x}_\mb{P} + \mbb{Q} \mbb{x}_\mb{Q} \big).
-            \end{align*}}'' (p.16)
-        \end{itemize}
-
-        \item[2)] \textcolor{TUDa-0c}{In the introduction, many general applications for machine learning in complex engineering applications are mentioned.
-However, I did not understand exactly the application that is considered in the article.
-In both examples, the parameter $R$ appearing in the explicit constraint is known.
-Please explain if this is realistic if e.g.~an application after model order reduction was the goal.
-If not, explain the applications you have in mind.}
-        \begin{itemize}
-        \item We added a remark that the choice of parameters is arbitrary and only serves to illustrate the approach.\\
-        ``\textcolor{red}{We note that the choice to consider $L$ and $C$, and not for example $R$, as parameters here is arbitrary and only serves to illustrate the approach.
-One could still follow the same approach when considering e.g.~the initial conditions, or the diode model, as being parameterized.}'' (p.11)
-        \item We outlined a potential application in the context of electrothermal simulations using the new example. (p.15)
-        \item Other potential applications include constrained optimization (with constraints on e.g.~currents through or voltages across certain devices) and validating the behavior of a circuit over its entire range of operating conditions (input currents/voltages, environment temperatures, device parameters) at the end of its development.
-        \end{itemize}
-
-        \item[3)] \textcolor{TUDa-0c}{Page 8, line 34-35: Can we always decouple the parameters into those that (may) appear in the algebraic equations and those that (may) appear in the ODE?
-Are there topological criteria for this?
-And is your approach reasonable if such a classification does not exist?}
-        \begin{itemize}
-            \item For the topological basis functions mentioned in section 3 this might be possible, but we have not explored this in more detail.
-            \item We pointed this out more explicitly in the manuscript and remarked that for sequential learning approaches, such as the one used in the paper, one can still benefit from parameters not appearing in the ODE without any explicit knowledge.\\
-            ``\textcolor{red}{As of now however, we have no knowledge of an easy way to automatically determine which parameters appear in the ODE.
-But when combined with a sequential learning strategy, such as the one outlined in \dots, there may still be computational savings due to the learning method requiring less samples for the parameters not appearing in the ODE.}'' (p.9)
-            \item The approach is also reasonable if the number of parameters cannot be reduced by the decoupling, since two significant advantages remain: better accuracy in the algebraic constraints and a reduced number of (differential) variables to learn.
-        \end{itemize}
-
-        \item[4)] \textcolor{TUDa-0c}{Page 12, line 32: Please explain how you reconstruct the values for $L$.}
-        \begin{itemize}
-            \item We added a paragraph that explains the reconstruction in more detail.\\
-            ``\textcolor{red}{This results from the fact that the ODE in \dots\ does not depend on $L$, such that the solution of $\varphi_3$ also does not depend on that parameter.
-The same goes for the algebraic variable $i_\mr{L}$, thus $\varphi_2$ and $\varphi_1$ can be reconstructed only based on the knowledge of $\varphi_3$.}'' (p.13)
-        \end{itemize}
-
-        \item[5)] \textcolor{TUDa-0c}{The bases of the dissection index are not unique, I guess. Are the results comparable for all bases? Is it e.g.~favorable to choose orthogonal bases?}
-        \begin{itemize}
-            \item The bases are indeed not unique.
-In the context of MNA there is a clear benefit to using the topological basis functions mentioned in section 3, and we added a remark explaining this.\\
-            ``\textcolor{red}{When considering large circuits, this topological decoupling along with its topological basis functions is to be preferred over other basis function choices, as it avoids the numerical computation of the basis functions, which becomes prohibitively expensive for large systems.}'' (p.8)
-        \end{itemize}
-
-        \item[6)] \textcolor{TUDa-0c}{You emphasize that the approach is independent of the machine learning method used.
-Is it also applicable for any other decoupling of DAEs into an inherent ODE and the algebraic constraints, also if it is based on other index concepts?}
-        \begin{itemize}
-            \item In principle it is possible to use any index concept that provides such a decoupling.
-There are however clear advantages when using the dissection index, in particular in the context of MNA, see also the answer to 5).
-            \item We added a remark about this at the beginning of section 3 and also motivated the dissection index more strongly.\\
-            ``\textcolor{red}{We focus on the dissection index, as it enables the decoupling of a DAE into an ODE and a set of purely algebraic equations.
-This is conceptually different from the perturbation index, however other index concepts such as the tractability and differentiation indices also use decoupling strategies.
-Still, the dissection index maintains some advantages over these concepts, as it provides a simple algorithmic procedure for the decoupling that is similar to the tractability index, but poses less strict smoothness assumptions.
-In the case of MNA without controlled sources \dots, it is even possible to find a purely topological decoupling based on the dissection index.
-We will not make use of this topological decoupling in the derivation however, but rather consider the dissection index for a more general class of DAEs, to formulate the assumptions that are necessary for our method to work also for DAEs other than \dots.}'' (p.4)
+            \item We added a footnote in section 2 to emphasize that this linearization is not necessary.
+It only serves to achieve a clear presentation of the approach by consistently writing MNA in a matrix form.\\
+            ``\textcolor{red}{Note that the definition of $\mb{G}$ via the Jacobian only serves to obtain a matrix form in \dots\ and has no impact on the general approach.
+As such, one may also work directly with the original system from \dots, however software implementations of MNA often work with $\mb{G}$ as defined in \dots\ (e.g.~when using Newton's method for solving nonlinear systems).}'' (p.2)
         \end{itemize}
@@
 
-    \paragraph{Reviewing: 2}
-    \begin{enumerate}
-        \item \textcolor{TUDa-0c}{The manuscript seems to present new and significant results, however the presentation quality is low.
-The reader is not properly guided through the mathematical derivations and the explanation of the proposed algorithms is weak and sometimes unclear.
-The proposed algorithm is mixed with derivations of formulas and it is difficult to deduce the proposed procedure starting from equations.
-For these reasons, although it is interesting, my recommendation is to reject the paper.}
-        \begin{itemize}
-            \item We reworked section 3, that introduces the dissection index, to more clearly link the ideas presented in the introduction and section 2 with our approach in section 4.
-Compare also the answer to 6) above, that motivates the dissection index more strongly.
-
-            \item We also used additional assumption (p.4-5) and definition (p.5-6) environments to better highlight the key points of section 3 and to give the section more structure, see also the answer to 1) above.
-
-            \item Furthermore, we shifted math heavy remarks to the appendix, compare the reply to 1) above, in order to streamline the explanation.
-
-            \item To more clearly distinguish the derivation of the dissection index from the introduction of our approach, we also expanded section 4.
-It now contains additional paragraphs describing the implementation for the index one and two cases.\\
-            ``\textcolor{red}{In order to recover the algebraic variables at time $t$ in the index one case, we only need to solve \dots\ for $\mbb{x}(t)$ using the learned $\mbt{x}(t)$.}\\
-            \textcolor{red}{In the index two case, we start by solving \dots\ for $\mbt{x}_\mb{P}(t)$ using the learned $\mbt{x}_\mb{Q}(t)$.
-Since we also require the derivative $\ddt \mbt{x}_\mb{P}(t)$ in \dots, we consider a small time increment $\Delta t$ and approximate the derivative using a backward difference
-            \begin{align*}
-                \ddt \mbt{x}_\mb{P}(t) \approx \frac{\mbt{x}_\mb{P}(t + \Delta t) - \mbt{x}_\mb{P}(t)}{\Delta t}.
-            \end{align*}
-            We note that this only reflects our implementation; in principle any finite difference (or similar) approximation is possible.
-Finally, we determine $\mbb{x}_\mb{P}(t)$ and $\mbb{x}_\mb{Q}(t)$ by jointly solving \dots\ using the learned $\mbt{x}_\mb{Q}(t)$, $\mbt{x}_\mb{P}(t)$ and the approximation of $\ddt \mbt{x}_\mb{P}(t)$.}'' (p.8-9)
-
-            \item Lastly, we removed the alternative approach (previous section 5.1) to achieve a clearer presentation of our main message.
-        \end{itemize}
-
-        \item \textcolor{TUDa-0c}{Since the concept of ``Index'' is key in understanding the manuscript the authors should define it properly and not just cite existent literature.}
-        \begin{itemize}
-            \item To better introduce the idea of an index, we shortened the introduction and devoted the entire subsection 2.1 to explaining index concepts in more detail.
-This also includes the definition of a second index concept (the perturbation index) along with an example to better link the term index with electrical circuits, and to motivate the importance of the notion of index in this context.
-(p.3)
-            \item The additional definition environments in section 3 (p.5-6) also serve to emphasize the definition of the dissection index.
-        \end{itemize}
-
-        \item \textcolor{TUDa-0c}{Moreover, some symbols and variables are poorly explained or defined, see as an example matrix $G$ appearing in equation 3.}
-        \begin{itemize}
-            \item We amended the definition of $\mb{G}$ to be directly compatible with the new explanation of multiport devices and the definition of a DAE in standard form, compare 1) above.\\
-            ``\textcolor{red}{In order to obtain a version of \dots\ that is better suited to analysis and implementation, we consider the device function Jacobians
-            \begin{align}
-                \mb{G} \big( \AT{R} \vphi \big) \coloneqq \mb{J}_{\AT{R} \vphi}(\gR), \quad \mb{L}(\i{L}) \coloneqq \mb{J}_{\i{L}}(\phiL), \quad \mb{C} \big( \AT{C} \vphi \big) \coloneqq \mb{J}_{\AT{C} \vphi}(\qC),
-            \end{align}
-            where we use the same notation for the Jacobians as in \dots.}'' (p.2)
-            \item During the reworking of section 2 and section 3, we also took great care to define all of the used symbols/variables.
-        \end{itemize}
-    \end{enumerate}
-
     We are confident that, considering our changes and additions, the manuscript now meets all acceptance criteria.

### 2024-04-05 doc/fig/ial.tex
graphical abstract
final revision

--- 2023-11-30 doc/fig/ial.tex
@@
 % landscape
+% \begin{tikzpicture}
+%     \node[rectangle, rounded corners, draw=black, semithick] (dae) at (0,0) {DAE};
+%     \node[rectangle, rounded corners, draw=black, semithick] (ode) at (2,-2) {\textcolor{TUDa-1a}{ODE}};
+%     \node[rectangle, rounded corners, draw=black, semithick] (ae) at (2,2) {\textcolor{TUDa-9a}{AE}};
+%     \node[rectangle, rounded corners, draw=black, semithick] (dd) at (5.5,-2) {\textcolor{TUDa-1a}{Differential variables}};
+%     \node[rectangle, rounded corners, draw=black, semithick] (ad) at (9.5,2) {\textcolor{TUDa-9a}{Algebraic variables}};
+
+%     \node[circle, fill=TUDa-9a, minimum size=3pt, inner sep=0] (ddae) at (5.5,2) {};
+
+%     \draw[->, >=latex, semithick, color=black] (dae) -- (ode) node[below, midway, sloped] {decouple};
+%     \draw[->, >=latex, semithick, color=black] (dae) -- (ae) node[above, midway, sloped] {decouple};
+%     \draw[<-, >=latex, semithick, color=TUDa-1a, dashed] (dd) -- (ode) node[above, midway, sloped] {learn};
+
+%     \draw[-, semithick, color=TUDa-9a, dotted] (dd) -- (ddae);
+%     \draw[-, semithick, color=TUDa-9a, dotted] (ae) -- (ddae);
+%     \draw[<-, >=latex, semithick, color=TUDa-9a, dotted] (ad) -- (ddae) node[above, midway, sloped] {reconstruct};
+% \end{tikzpicture}
+
+% poster
 \begin{tikzpicture}
     \node[rectangle, rounded corners, draw=black, semithick] (dae) at (0,0) {DAE};
-    \node[rectangle, rounded corners, draw=black, semithick] (ode) at (2,-2) {\textcolor{TUDa-1a}{ODE}};
-    \node[rectangle, rounded corners, draw=black, semithick] (ae) at (2,2) {\textcolor{TUDa-9a}{AE}};
-    \node[rectangle, rounded corners, draw=black, semithick] (dd) at (5.5,-2) {\textcolor{TUDa-1a}{Differential DOFs}};
-    \node[rectangle, rounded corners, draw=black, semithick] (ad) at (9.5,2) {\textcolor{TUDa-9a}{Algebraic DOFs}};
+    \node[rectangle, rounded corners, draw=black, semithick] (ode) at (2,-2) {\textcolor{TUDa-1d}{ODE}};
+    \node[rectangle, rounded corners, draw=black, semithick] (ae) at (2,2) {\textcolor{TUDa-3d}{AE}};
+    \node[rectangle, rounded corners, draw=black, semithick] (dd) at (5.5,-2) {\textcolor{TUDa-1d}{Differential variables}};
+    \node[rectangle, rounded corners, draw=black, semithick] (ad) at (9.5,2) {\textcolor{TUDa-3d}{Algebraic variables}};
 
-    \node[circle, fill=TUDa-9a, minimum size=3pt, inner sep=0] (ddae) at (5.5,2) {};
+    \node[circle, fill=TUDa-3c, minimum size=3pt, inner sep=0] (ddae) at (5.5,2) {};
 
-    \draw[->, >=latex, semithick, color=black] (dae) -- (ode) node[above, midway, sloped] {decouple};
+    \draw[->, >=latex, semithick, color=black] (dae) -- (ode) node[below, midway, sloped] {decouple};
     \draw[->, >=latex, semithick, color=black] (dae) -- (ae) node[above, midway, sloped] {decouple};
-    \draw[<-, >=latex, semithick, color=TUDa-1a, dashed] (dd) -- (ode) node[above, midway, sloped] {learn};
+    \draw[<-, >=latex, semithick, color=TUDa-1d, dashed] (dd) -- (ode) node[above, midway, sloped] {learn};
 
-    \draw[-, semithick, color=TUDa-9a, dotted] (dd) -- (ddae);
-    \draw[-, semithick, color=TUDa-9a, dotted] (ae) -- (ddae);
-    \draw[<-, >=latex, semithick, color=TUDa-9a, dotted] (ad) -- (ddae) node[above, midway, sloped] {reconstruct};
+    \draw[-, semithick, color=TUDa-3d, dotted] (dd) -- (ddae);
+    \draw[-, semithick, color=TUDa-3d, dotted] (ae) -- (ddae);
+    \draw[<-, >=latex, semithick, color=TUDa-3d, dotted] (ad) -- (ddae) node[above, midway, sloped] {reconstruct};
 \end{tikzpicture}
